### 2020-01-07 paper1/doc/paper1.tex
Add initial documents

%
% FH Technikum Wien
% !TEX encoding = UTF-8 Unicode
%
% Erstellung von Master- und Bachelorarbeiten an der FH Technikum Wien mit Hilfe von LaTeX und der Klasse TWBOOK
%
% Um ein eigenes Dokument zu erstellen, müssen Sie folgendes ergänzen:
% 1) Mit \documentclass[..] einstellen: Master- oder Bachelorarbeit, Studiengang und Sprache
% 2) Mit \newcommand{\FHTWCitationType}.. Zitierstandard festlegen (wird in der Regel vom Studiengang vorgegeben - bitte erfragen)
% 3) Deckblatt, Kurzfassung, etc. ausfüllen
% 4) und die Arbeit schreiben (die verwendeten Literaturquellen in Literatur.bib eintragen)
%
% Getestet mit TeXstudio mit Zeichenkodierung ISO-8859-1 (=ansinew/latin1) und MikTex unter Windows
% Zu beachten ist, dass die Kodierung der Datei mit der Kodierung des paketes inputenc zusammen passt!
% Die Kodierung der Datei twbook.cls MUSS ANSI betragen!
% Bei der Verwendung von UTF8 muss dnicht nur die Kodierung des Dokuments auf UTF8 gestellt sein, sondern auch die des BibTex-Files!
%
% Bugreports und Feedback bitte per E-Mail an latex@technikum-wien.at
%
% Versionen
% *) V0.7: 9.1.2015, RO: Modeline angepasst und verschoben
% *) V0.6: 10.10.2014, RO: Weitere Anpassung an die UK
% *) V0.5: 8.8.2014, WK: Literaturquellen überarbeitet und angepasst
% *) V0.4: 4.8.2014, WK: Initalversion in SVN eingespielt
%
\documentclass[MIS,Projekt,english]{twbook}%\documentclass[Bachelor,BMR,ngerman]{twbook}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

%
% Bitte in der folgenden Zeile den Zitierstandard festlegen
\newcommand{\FHTWCitationType}{IEEE} % IEEE oder HARVARD möglich - wenn Sie zwischen IEEE und HARVARD wechseln, bitte die temorären Dateien (aux, bbl, ...) löschen
%
\ifthenelse{\equal{\FHTWCitationType}{HARVARD}}{\usepackage{harvard}}{\usepackage{bibgerm}}

% Definition Code-Listings Formatierung:
\usepackage[final]{listings}
\lstset{captionpos=b, numberbychapter=false,caption=\lstname,frame=single, numbers=left, stepnumber=1, numbersep=2pt, xleftmargin=15pt, framexleftmargin=15pt, numberstyle=\tiny, tabsize=3, columns=fixed, basicstyle={\fontfamily{pcr}\selectfont\footnotesize}, keywordstyle=\bfseries, commentstyle={\color[gray]{0.33}\itshape}, stringstyle=\color[gray]{0.25}, breaklines, breakatwhitespace, breakautoindent}
\lstloadlanguages{[ANSI]C, C++, [gnu]make, gnuplot, Matlab}

%Formatieren des Quellcodeverzeichnisses
\makeatletter
% Setzen der Bezeichnungen für das Quellcodeverzeichnis/Abkürzungsverzeichnis in Abhängigkeit von der eingestellten Sprache
\providecommand\listacroname{}
\@ifclasswith{twbook}{english}
{%
    \renewcommand\lstlistingname{Code}
    \renewcommand\lstlistlistingname{List of Code}
    \renewcommand\listacroname{List of Abbreviations}
}{%
    \renewcommand\lstlistingname{Quellcode}
    \renewcommand\lstlistlistingname{Quellcodeverzeichnis}
    \renewcommand\listacroname{Abkürzungsverzeichnis}
}
% Wenn die Option listof=entryprefix gewählt wurde, Definition des Entyprefixes für das Quellcodeverzeichnis. Definition des Macros listoflolentryname analog zu listoflofentryname und listoflotentryname der KOMA-Klasse
\@ifclasswith{scrbook}{listof=entryprefix}
{%
    \newcommand\listoflolentryname\lstlistingname
}{%
}
\makeatother
\newcommand{\listofcode}{\phantomsection\lstlistoflistings}

% Die nachfolgenden Pakete stellen sonst nicht benötigte Features zur Verfügung
\usepackage{blindtext}

%
% Einträge für Deckblatt, Kurzfassung, etc.
%
\title{Projekt 1\\Paper1}
\author{Max\and Bernhard\and Frederic\and Sebastian}
%\studentnumber{XXXXXXXXXXXXXXX}
%\author{Titel Vorname Name, Titel\and{}Titel Vorname Name, Titel}
%\studentnumber{XXXXXXXXXXXXXXX\and{}XXXXXXXXXXXXXXX}
\supervisor{Titel Vorname Name, Titel}
%\supervisor[Begutachter]{Titel Vorname Name, Titel}
%\supervisor[Begutachterin]{Titel Vorname Name, Titel}
%\secondsupervisor{Titel Vorname Name, Titel}
%\secondsupervisor[Begutachter]{Titel Vorname Name, Titel}
%\secondsupervisor[Begutachterinnen]{Titel Vorname Name, Titel}
\place{Wien}
\kurzfassung{\blindtext}
\schlagworte{Schlagwort1, Schlagwort2, Schlagwort3, Schlagwort4}
\outline{\blindtext}
\keywords{Keyword1, Keyword2, Keyword3, Keyword4}
%\acknowledgements{\blindtext}

\begin{document}

%Festlegungen für den HARVARD-Zitierstandard
\ifthenelse{\equal{\FHTWCitationType}{HARVARD}}{
\bibliographystyle{Harvard_FHTW_MR}%Zitierstandard FH Technikum Wien, Studiengang Mechatronik/Robotik, Version 1.2e
\citationstyle{dcu}%Correct citation-style (Harvardand, ";" between citations, "," between author and year)
\citationmode{abbr}%use "et al." with first citation
\iflanguage{ngerman}{
    %Deutsch Neue Rechtschreibung
    \newcommand{\citepic}[1]{(Quelle: \protect\cite{#1})}%Zitat: Bild
    \newcommand{\citefig}[2]{(Quelle: \protect\cite{#1}, S. #2)}%Zitat: Bild aus Dokument
    \newcommand{\citefigm}[2]{(Quelle: modifiziert "ubernommen aus \protect\cite{#1}, S. #2)}%Zitat: modifiziertes Bild aus Dokument
    \newcommand{\citep}{\citeasnoun}%In-Line Zitiat entweder mit \citep{} oder \citeasnoun{}
    \newcommand{\acessedthrough}{Verf{\"u}gbar unter:}%Für URL-Angabe
    \newcommand{\acessedthroughp}{Verf{\"u}gbar bei:}%Für URL-Angabe (Geschützte Datenbank, Zugriff durch FH)
    \newcommand{\acessedat}{Zugang am}%Für URL-Datum-Angabe
    \newcommand{\singlepage}{S.}%Für Seitenangabe (einzelne Seite)
    \newcommand{\multiplepages}{S.}%Für Seitenangabe (mehrere Seiten)
    \newcommand{\chapternr}{K.}%Für Kapitelangabe
    \renewcommand{\harvardand}{\&}%Harvardand in Zitaten
    \newcommand{\abstractonly}{ausschließlich Abstract}
    \newcommand{\edition}{. Auflage}%Angabe der Auflage
}{
\iflanguage{german}{
    %Deutsch
    \newcommand{\citepic}[1]{(Quelle: \protect\cite{#1})}%Zitat: Bild
    \newcommand{\citefig}[2]{(Quelle: \protect\cite{#1}, S. #2)}%Zitat: Bild aus Dokument
    \newcommand{\citefigm}[2]{(Quelle: modifiziert "ubernommen aus \protect\cite{#1}, S. #2)}%Zitat: modifiziertes Bild aus Dokument
    \newcommand{\citep}{\citeasnoun}%In-Line Zitiat entweder mit \citep{} oder \citeasnoun{}
    \newcommand{\acessedthrough}{Verf{\"u}gbar unter:}%Für URL-Angabe
    \newcommand{\acessedthroughp}{Verf{\"u}gbar bei:}%Für URL-Angabe (Geschützte Datenbank, Zugriff durch FH)
    \newcommand{\acessedat}{Zugang am}%Für URL-Datum-Angabe
    \newcommand{\singlepage}{S.}%Für Seitenangabe (einzelne Seite)
    \newcommand{\multiplepages}{S.}%Für Seitenangabe (mehrere Seiten)
    \newcommand{\chapternr}{K.}%Für Kapitelangabe
    \renewcommand{\harvardand}{\&}%Harvardand in Zitaten
    \newcommand{\abstractonly}{ausschließlich Abstract}
    \newcommand{\edition}{. Auflage}%Angabe der Auflage
}{
    %Englisch
    \newcommand{\citepic}[1]{(Source: \protect\cite{#1})}%Zitat: Bild
    \newcommand{\citefig}[2]{(Source: \protect\cite{#1}, p. #2)}%Zitat: Bild aus Dokument
    \newcommand{\citefigm}[2]{(Source: taken with modification from \protect\cite{#1}, p. #2)}%Zitat: modifiziertes Bild aus Dokument
    \newcommand{\citep}{\citeasnoun}%In-Line Zitiat entweder mit \citep{} oder \citeasnoun{}
    \newcommand{\acessedthrough}{Available at:}%Für URL-Angabe
    \newcommand{\acessedthroughp}{Available through:}%Für URL-Angabe (Geschützte Datenbank, Zugriff durch FH)
    \newcommand{\acessedat}{Accessed}%Für URL-Datum-Angabe
    \newcommand{\singlepage}{p.}%Für Seitenangabe (einzelne Seite)
    \newcommand{\multiplepages}{pp.}%Für Seitenangabe (mehrere Seiten)
    \newcommand{\chapternr}{Ch.}%Für Kapitelangabe
    \renewcommand{\harvardand}{\&}%Harvardand in Zitaten
    \newcommand{\abstractonly}{Abstract only}
    \newcommand{\edition}{~edition}%Edition -> note, that you have to write "edition = {2nd},"!
}}}

\maketitle

%
% .. und hier beginnt die eigentliche Arbeit. Viel Erfolg beim Verfassen!
%
\chapter{Erste Überschrift der Ebene 1 (chapter)}
\blinddocument

\blindmathpaper

\section{Erste Überschrift Tiefe 2 (section)}
\blindtext

\subsection{Erste Überschrift Tiefe 3 (subsection)}
\blindtext

\subsubsection{Erste Überschrift Tiefe 4 (subsubsection)}
\blindtext

\chapter{Zweite Überschrift der Tiefe 1 (chapter)}
\blindtext

\section{Zweite Überschrift Tiefe 2 (section)}
\blindtext

\section{Zweite Überschrift Tiefe 2 (section)}
\blindtext

\subsection{Zweite Überschrift Tiefe 3 (subsection)}
\blindtext

\subsection{Dritte Überschrift Tiefe 3 (subsection)}
\blindtext

\subsubsection{Zweite Überschrift Tiefe 4 (subsubsection)}
\blindtext

\noindent Querverweise werden in \LaTeX{} automatisch erzeugt und verwaltet, damit sie leicht aktualisiert werden können. Hier wird zum Beispiel auf Abbildung \ref{Abb1} verwiesen.

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.5\linewidth]{PICs/buchruecken}
\caption{Beispiel für die Beschriftung eines Buchrückens.}\label{Abb1}
\end{figure}
\begin{figure}[!htbp]
\centering
\includegraphics[width=0.5\linewidth]{PICs/buchruecken}
\caption{2. Beispiel für die Beschriftung eines Buchrückens.}\label{Abb2}
\end{figure}

Und hier ist ein Verweis auf Tabelle \ref{tab1}. Das gezeigte Tabellenformat ist nur ein Beispiel. Tabellen können individuell gestaltet werden.

\begin{table}[!htbp]
\centering
\caption{Semesterplan der Lehrveranstaltung \glqq Angewandte Mathematik\grqq.}\label{tab1}
\begin{tabular}{| p{0.3\linewidth} | p{0.3\linewidth} | p{0.3\linewidth} |}\hline
Datum & Thema & Raum\\\hline
20.08.2008 & Graphentheorie & HS 3.13\\
01.10.2008 & Biomathematik & HS 1.05\\\hline
\end{tabular}
\end{table}
\begin{table}[!htbp]
\centering
\caption{2. Semesterplan der Lehrveranstaltung \glqq Angewandte Mathematik\grqq.}\label{tab2}
\begin{tabular}{| p{0.3\linewidth} | p{0.3\linewidth} | p{0.3\linewidth} |}\hline
Datum & Thema & Raum\\\hline
20.08.2008 & Graphentheorie & HS 3.13\\
01.10.2008 & Biomathematik & HS 1.05\\\hline
\end{tabular}
\end{table}

Hier wird auf die Formel \ref{Gl1} verwiesen.

\begin{align}
x = -\frac{p}{2}\pm\sqrt{\frac{p^2}{4}-q}\label{Gl1}
\end{align}
\begin{align}
x = -\frac{p}{2}\pm\sqrt{\frac{p^2}{4}-q}\label{Gl2}
\end{align}

\begin{lstlisting}[language=C++,name={1. Beispiel},label={sc:bsp:1}]
#include <iostream>

void SayHello(void)
{
    // Kommentar
    cout << "Hello World!" << endl;
}

int main(int argc, char **argv)
{
    SayHello();
    return 0;
}
\end{lstlisting}

Literaturverweise sollten automatisch verwaltet werden, vor allem, wenn es viele Quellenverweise gibt. Beispiele sind  \cite{Ko05a}, \cite{Ko05b}, \cite{MiGo05}, \cite{TeGo14}, \cite{HuHa07}, \cite{HuZi10}, \cite{ZiKu07}, \cite{He07}, \cite{SIE11}, \cite{SIE14}, \cite{ISO98}, \cite{ATM11}, \cite{Hu11}, \cite{Po10}. Das verwendete Zitierformat (bzw.~das Format des Literaturverzeichnisses) ist entspechend der Vorgaben der Studiengänge zu wählen.
Es wird dringend empfohlen, BibTeX~zu verwenden (wie in diesem Beispiel).

\chapter{Dritte Überschrift der Tiefe 1 (chapter)}
\begin{figure}[!htbp]
\centering
\includegraphics[width=0.5\linewidth]{PICs/buchruecken}
\caption{3. Beispiel für die Beschriftung eines Buchrückens.}\label{Abb3}
\end{figure}
\begin{figure}[!htbp]
\centering
\includegraphics[width=0.5\linewidth]{PICs/buchruecken}
\caption{4. Beispiel für die Beschriftung eines Buchrückens.}\label{Abb4}
\end{figure}

\begin{table}[!htbp]
\centering
\caption{3. Semesterplan der Lehrveranstaltung \glqq Angewandte Mathematik\grqq.}\label{tab3}
\begin{tabular}{| p{0.3\linewidth} | p{0.3\linewidth} | p{0.3\linewidth} |}\hline
Datum & Thema & Raum\\\hline
20.08.2008 & Graphentheorie & HS 3.13\\
01.10.2008 & Biomathematik & HS 1.05\\\hline
\end{tabular}
\end{table}
\begin{table}[!htbp]
\centering
\caption{4. Semesterplan der Lehrveranstaltung \glqq Angewandte Mathematik\grqq.}\label{tab4}
\begin{tabular}{| p{0.3\linewidth} | p{0.3\linewidth} | p{0.3\linewidth} |}\hline
Datum & Thema & Raum\\\hline
20.08.2008 & Graphentheorie & HS 3.13\\
01.10.2008 & Biomathematik & HS 1.05\\\hline
\end{tabular}
\end{table}

\begin{align}
x = -\frac{p}{2}\pm\sqrt{\frac{p^2}{4}-q}\label{Gl3}
\end{align}
\begin{align}
x = -\frac{p}{2}\pm\sqrt{\frac{p^2}{4}-q}\label{Gl4}
\end{align}
\begin{lstlisting}[language=C++,name={2. Beispiel},label={sc:bsp:2}]
#include <iostream>

void SayHello(void)
{
    // Kommentar
    cout << "Hello World!" << endl;
}

int main(int argc, char **argv)
{
    SayHello();
    return 0;
}
\end{lstlisting}

%
% Hier beginnen die Verzeichnisse.
%
\clearpage
\ifthenelse{\equal{\FHTWCitationType}{HARVARD}}{}{\bibliographystyle{gerabbrv}}
\bibliography{Literatur}
\clearpage

% Das Abbildungsverzeichnis
\listoffigures
\clearpage

% Das Tabellenverzeichnis
\listoftables
\clearpage

% Das Quellcodeverzeichnis
\listofcode
\clearpage

\phantomsection
\addcontentsline{toc}{chapter}{\listacroname}
\chapter*{\listacroname}
\begin{acronym}[XXXXX]
    \acro{ABC}[ABC]{Alphabet}
    \acro{WWW}[WWW]{world wide web}
    \acro{ROFL}[ROFL]{Rolling on floor laughing}
\end{acronym}

%
% Hier beginnt der Anhang.
%
\clearpage
\appendix
\chapter{Anhang A}
\clearpage
\chapter{Anhang B}
\end{document}

### 2020-02-22 paper2/paper2.tex
Add first version of simulation chapter

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{url}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Network Data Collection for Artifical Intelligence
}

\author{\IEEEauthorblockN{Sebastian Lipp}
\IEEEauthorblockA{\textit{IT Security} \\
\textit{FH Technikum Wien}\\
Vienna, Austria \\
cs19m032@technikum-wien.at}
\and
\IEEEauthorblockN{Damir Marijanovic}
\IEEEauthorblockA{\textit{IT Security} \\
\textit{FH Technikum Wien}\\
Vienna, Austria \\
cs19m031@technikum-wien.at}
\and
\IEEEauthorblockN{Boris Stampf}
\IEEEauthorblockA{\textit{IT Security} \\
\textit{FH Technikum Wien}\\
Vienna, Austria \\
cs19m006@technikum-wien.at}
}

\maketitle

\begin{abstract}
Collecting real network data for further processing is an important topic for the training of neuronal networks which filter suspicious from ordinary traffic. This paper focus mainly on research in the areas of network simulation, network recording and data extraction which should serve as input for the following papers.
\end{abstract}

\begin{IEEEkeywords}
artifical intelligence, neuronal networks, computer networks, network security, data collection
\end{IEEEkeywords}

\section{Introduction}
Neuronal networks which filter suspicious network traffic need a training data set of high-quality. This data set could be produced from real network traffic but this can be difficult to implement. Another option considers network simulators which can be easily configured depending on the selected scenarios.

This paper is split in three chapters. The first chapter compares available network simulators and their software interfaces and explores possibilities for their automated setup based on the scenario and parameters. The second chapter focuses on tools to record the network traffic. In the third chapter methods of feature extraction and label definition are covered. This information serves as foundation for the further implementation of a system for automated data collection for artifical intelligence.

At this point all our considerations are based on theoretical research  without practical implementation which will come to life in further work. Therefore this paper sets a direction and foundation for the next steps and shows possible solutions on a high-level.

\section{Simulation}

Before network simulators can be compared the requirements have to be defined. Because the produced network traffic should be close to real traffic it would be advisable to choose a network simulator which works with the simulation of real network equipment and can also integrate real hardware or virtual machines into the simulation which is necessary to redirect the network traffic to the recorder and data extractor. This is important to consider since other simulators like NS3 provide discrete-event network simulation which is a needless complexity overhead for this kind of application. It would be desirable if they are affordable and provide an easy-to-use interface to create and control specific network topologies through a programming language like Python to automate their creation. Moreover they should be resource-efficient and easy to deploy. The provisioning of the simulation hosts has to be simple and also scalable to be prepared for simulating denial-of-service attacks.

The next sections cover possible network simulators, system environments and the provisioning of the infrastructure.

\subsection{Comparing common network simulators}
There are many solutions available on the market to simulate networks. Popular network simulators which possibly fullfill the requirements above are for example VIRL, GNS3 or EVE-NG. This section focuses on the comparision of these simulators and the selection of the best solution for our use case.  \cite{b1}

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.55]{virl.png}}
\caption{Sample network topology in VIRL \cite{b2}}
\label{virl}
\end{figure}

VIRL (Cisco Virtual Internet Routing Lab) is a network modeling and simulation environment provided by Cisco which offers the ability to use simulation models of real Cisco platforms, integrate virtual machines and connect real networks. Furthermore it provides a RESTful API to configure networks without a graphical interface, sufficient documentation and scalability as well as easy deployment through prepared images. It is not for free and only supports VMware or bare-metal installs. Simulation models of other vendors are not supported. Tests in virtual machines unveiled a good responsiveness but a long topology loading time with high CPU load. Fig.~\ref{virl} shows a sample network topology created via the graphical user interface of VIRL.  \cite{b1} \cite{b3}

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.13]{gns3.png}}
\caption{Sample network topology in GNS3 \cite{b4}}
\label{gns3}
\end{figure}

GNS3 (Graphical Network Simulator 3) is a very popular free open source software to emulate and test networks. It offers the integration of virtual machines and real networks and supports models of various vendors but does not offer pre-installed device images due to license issues. It also provides a RESTful API to setup network topologies, a big community and scalibility. The software can be easily installed through packages on all operating systems and supports all hypervisors as well Docker containers. Performance tests showed a long topology loading time with high CPU load but a good response through the user interface. Fig.~\ref{gns3} shows a sample network topology created with GNS3.  \cite{b1} \cite{b5}

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.19]{eve-ng.png}}
\caption{Sample network topology in EVE-NG \cite{b6}}
\label{eve-ng}
\end{figure}

EVE-NG (Emulated Virtual Environment - Next Gen) is a network emulator which comes in a free and a professional edition and offers the interaction with real networks. It supports multiple vendors but does not provide device images. Advanced features like Docker, configuration management or Wireshark integration only comes with the purchase version. Moreover a RESTful API and images for the installation in virtual machines are available as well as a reasonable documentation. The software is scalable and performs well in virtual machines with a short topology loading time and provides a intuitive interface as shown in Fig.~\ref{eve-ng}.  \cite{b1} \cite{b7}

All solutions are very similar in their functionalities but distinguish in their vendor support, performance, documentation and price. VIRL only supports Cisco devices but comes with device images like GNS3 and EVE-NG support various vendors but do not ship images. But many proprietary and open source images are available on the internet for free. VIRL and EVE-NG are propriatary like GNS3 is free and open source. EVE-NG performs better in virtual machines than VIRL and GNS3.

Regarding these factors GNS3 fullfills our requirements the most because it is free, supports all vendors and comes with a big community and documentation. Our second option would be EVE-NG in the professional edition which also supports all vendors.

\subsection{Selecting the deployment environment}

Next steps include the integration of the selected network simulator in an scalable environment which can be setup easily.

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.25]{docker_vm.png}}
\caption{Virtual machines vs. Containers \cite{b8}}
\label{docker-vm}
\end{figure}

Hypervisors like VirtualBox or VMware provide the possibility to setup full virtual computers on a physical host where Docker containers are encapsuled sandboxes on an existing operating system like illustrated in Fig.~\ref{docker-vm}. It is clear that containers are more resource efficent but in virtual machines the systems are better isolated. In our case we could use both solutions because the network simulators are also capable to run in Docker containers. If we focus on resource efficency and scalability also Kubernetes comes in our minds which is an orchestrator for deploying Docker containers on multiple nodes but it is a more complex configuration overhead and needs certain amount of hardware to unfold its potential also for simple scenarios. Because we want an affordable solutions which can be used in most of the cases also on our computers with some exceptions, we decided to use a tool which can create virtual machines as well as Docker containers to be ready for the future. Regarding the corner cases servers or a Kubernetes cluster could be rent temporary on cloud platforms to support high-load scenarios e.g. big denial-of-service attacks.

A popular solution to bootstrap development environments in virtual machines is called Vagrant which provides lightweight base images and an interface automate the configuration of the virtual machines. Moreover it can be connected to provisioning tools to further configure host systems. Vagrant supports VirtualBox or VMware but also Docker and can be extended to support cloud services like AWS or even to install a whole Kubernetes cluster. It gives us the power to start with simple configuration and scale them up to more sophisticated ones using containerization and complex orchestrators. \cite{b9}

This means we are ready to create system environments on our host or in the cloud and install base systems on them. The next section focuses on the automated setup of the network simulator, recorder and data extractor within virtual machines or Docker containers as well as the configuration of network topologies based on the selected scenarios and parameters to produce, record and extract network traffic for further processing in neuronal networks to improve their strike rate.

\subsection{Provisioning of the infrastructure}

Vagrant supports multiple provisioners like Ansible, Terraform or simple bash scripts to configure the host systems and can also be setup through them. For the beginning we will use Ansible through Vagrant to configure the host system. Vagrant can setup multiple virtual machines or Docker containers on one host. If we would need to setup machines on multiple hosts we will come back to tools like Terraform for example. \cite{b9}

If all the possible network scenarios are defined - this was the scope of another paper - this network topologies have to be deployed through our deployment system consisting of Vagrant and Ansible. This means we will define the basic host setup and the network configurations for all the scenarios in Ansible templates which could be later adjusted via command line parameters. The topology can then be deployed by executing a shell command. This includes the deployment of the network recorder and the data extractor. In other words one or multiple scenarios can be started and connected to one recorder and data extractor to combine simultanuous attacks and everything will be deployed through a shell command with static configuration in templates and dynamic configuration with command line options. It will be also possible to integrate real hardware into the simulation.

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.28]{design_flow.png}}
\caption{Data collection workflow}
\label{design-flow}
\end{figure}

Fig.~\ref{docker-vm} shows a schematic overview of the simulation workflow. The input parameters are the type of scenario(s) and the respective settings which deploys hosts including the base systems and the simulators applying the network topologies. Moreover the network recorder and data extractor will be deployed to gather the network traffic and extract the features to create data sets which could be collected in a central database.

A real test case would then consist of multiple scenarios of attacks and usual network traffic recorded, extracted and collected in a central place. The next chapter focus on recording the generated data.

\section{Recording}

\section{Data Extraction}

\begin{thebibliography}{00}
\bibitem{b1} CBT Nuggets Blog (2020). 5 Best Network Simulators for Cisco Exams: CCNA, CCNP, CCIE. [online] Available at: \newline\url{https://www.cbtnuggets.com/blog/career/career-progression/5-best-network-simulators-for-cisco-exams-ccna-ccnp-and-ccie} [Accessed 22 Feb. 2020].
\bibitem{b2} Network Computing. (2020). Cisco VIRL: More Than A Certification Study Lab. [online] Available at: \url{https://www.networkcomputing.com/networking/cisco-virl-more-certification-study-lab} [Accessed 22 Feb. 2020].
\bibitem{b3} Virl.cisco.com. (2020). VIRL Getting Started Tutorial. [online] Available at: \url{http://virl.cisco.com/virl.apis.php} [Accessed 22 Feb. 2020].
\bibitem{b4} Gns3.com. (2020). GNS3 | The software that empowers network professionals. [online] Available at: \url{https://gns3.com/discussions/how-to-install-gns3-on-centos-7-} [Accessed 22 Feb. 2020].
\bibitem{b5} Gns3-server.readthedocs.io. (2020). Welcome to API documentation! — GNS3 2.1.22dev1-67e70c46 documentation. [online] Available at: \url{https://gns3-server.readthedocs.io/en/latest/} [Accessed 22 Feb. 2020].
\bibitem{b6} Eve-ng.net. (2020). Screenshots. [online] Available at: https://www.eve-ng.net/index.php/screenshots/ [Accessed 22 Feb. 2020].
\bibitem{b7} Eve-ng.net. (2020). EVE-NG API. [online] Available at: \url{https://www.eve-ng.net/index.php/documentation/howtos/how-to-eve-ng-api/} [Accessed 22 Feb. 2020].
\bibitem{b8} SOSC-2018. (2020). Docker. [online] Available at: \url{https://dodas-ts.github.io/SOSC-2018/docker.html} [Accessed 22 Feb. 2020].
\bibitem{b9} Chan, M. (2020). 15 Infrastructure as Code Tools to Automate Deployments - Thorn Tech. [online] Thorn Technologies. Available at: \url{https://www.thorntech.com/2018/04/15-infrastructure-as-code-tools/} [Accessed 22 Feb. 2020].
\end{thebibliography}
\vspace{12pt}
\color{red}
\end{document}

### 2020-02-23 paper2/paper2.tex
First version of data extraction section

@@
 
+Datasets for intrusion detection are often based on network packet data captured by sniffers, which can be stored in pcap files (as done in the UNSW-NB15 \cite{Nb2015} and CSE-CIC-IDS2018 \cite{Ids2018} datasets).
+Before a machine learning algorithm can be employed, features have to be extracted from the captured data.
+A wide range of different features is described in literature.
+For the sake of clarity, these features are divided into groups which are described in the following subsections.
+
+There also exists a large number of techniques for dimensionality reduction.
+The 40 features of the NSL-KDD data set could be reduced to 30 or 16 features, with no or low degradation in detection power, while highly reducing computation costs, by combining the following methods \cite{vasquez2014}:
+\begin{itemize}
+	\item Weight by maximum relevance
+	\item Minimum redundancy maximum relevance
+	\item Significance analysis for microarrays
+	\item Least absolute selection and shrinkage operator
+	\item Stability selection
+	\item Stepwise regression
+\end{itemize}
+
+Another aspect which has been examined is the stability of features in terms of time (year 2000/2001 vs.
+2008) and location (Internet core vs.
+Internet edge).
+In this regard, the packet size can be considered a stable feature \cite{este2009}.
+
+\subsection{Flow statistical features}
+
+A common approach is to group packets into flows, and then calculate statistical features based on these flows.
+Generally, a flow is a group of packets with the same source and destination IP addresses and ports, as well as protocol \cite{brownlee1998}.
+However, a flow can also be terminated by a TCP FIN packet or a timeout \cite{Ids2018}.
+Statistical features of network flows include the following:
+\begin{itemize}
+	\item Flow duration \cite{Nb2015, Ids2018}
+	\item Time between two flows or packets \cite{Ids2018}
+	\item Time between SYN and SYN/ACK, and between SYN/ACK and ACK packets during TCP connection setup \cite{Nb2015}
+	\item Active time before a flow becomes idle and vice versa \cite{Ids2018}
+	\item Jitter \cite{Nb2015}
+	\item Number of packets in total, with at least 1 byte of payload, or with certain protocol flags \cite{Ids2018, waizumi2007, celik2015}
+	\item Number of fragmented packets \cite{waizumi2007}
+	\item Number of retransmitted or dropped packets \cite{Nb2015}
+	\item Size of packets, payload, headers or segments \cite{Ids2018, celik2015}
+	\item Number of bytes/packets transferred per second, in bulk, or in the initial TCP window \cite{Ids2018, celik2015}
+	\item Number of RTT samples \cite{celik2015}
+	\item Time to live value \cite{Nb2015}
+	\item TCP window advertisement value \cite{Nb2015}
+	\item TCP base sequence number \cite{Nb2015}
+	\item Ratio between download and upload \cite{Ids2018}
+	\item Ratio between maximum and minimum packet size \cite{celik2015}
+	\item Number of flows with the current flow's IP address or port number \cite{waizumi2007, Nb2015}
+\end{itemize}
+Most of these features can be derived by statistical calculations such as total, minimum, maximum, mean, median, standard deviation, variance, skewness and kurtosis \cite{Ids2018, celik2015, xu2015}.
+Also, the features can be extracted from bidirectional flows, or separately for forward and backward directions \cite{Ids2018}.
+
+\subsection{Timeslot features}
+
+Timeslot features are extracted by counting the number of certain events in fixed time intervals.
+For the detection of DoS and probing traffic, the following timeslot features are proposed in \cite{waizumi2007}:
+\begin{itemize}
+	\item Number of TCP, UDP and ICMP packets
+	\item Number of bytes sent and received through all TCP connections
+	\item Number of TCP ports
+	\item Number of occurences of TCP flags
+	\item Number of DNS packets
+	\item Number of fragmented packets
+	\item Number of values of the four fields of IP addresses
+\end{itemize}
+
+\subsection{Behavioral features}
+
+A different approach is to define features which measure specific behavioral differences between benign and malicious software.
+The UNSW-NB15 datasets contains mostly flow statistical features, but also a few features based on behavior \cite{Nb2015}:
+\begin{itemize}
+	\item A binary value denoting if the source and destination IP addresses and port numbers are equal
+	\item Service (HTTP, FTP, SMTP, SSH, DNS, or IRC)
+	\item Pipelined depth into the connection of HTTP request/response transaction
+	\item Actual uncompressed content size of the data transferred from the server's HTTP service
+	\item Number of flows containing HTTP GET and POST methods
+	\item A binary value denoting if a FTP session is accessed by user and password
+	\item Number of flows containing FTP commands
+\end{itemize}
+
+Reference \cite{li2013} lists the following features which can be used for detecting communication with known and unknown botnets:
+\begin{itemize}
+	\item Accessing the main and backup DNS server at the same time (benign software would only access the backup DNS server if the main DNS server is not available)
+	\item Resolving a specific domain name periodically
+	\item Accessing a fast-flux service network (FFSN)
+	\item Downloading malicious binary code (which can be checked using antivirus software)
+	\item Scanning for open ports
+	\item Periodically creating null TCP connections (zero TCP window or IRC PING/PONG connections which are used for contacting the command and control server)
+\end{itemize}
+
+Similarily, the character code distribution of payloads can be used to detect buffer overflow traffic (where some character codes have exceptionally high frequency) \cite{waizumi2007}.
+
+\subsection{Host-based features}
+
+The previous sections examined different features which can be extracted from network traffic, but there are also host-based data which can be used for intrusion detection.
+These include operating system event logs \cite{Ids2018} and other log files (e.g., firewall, mail and FTP logs) \cite{abad2003}.
+
+\subsection*{Label definition}
+
+In addition to the feature vectors, datasets to be used for supervised learning need to include labels.
+For the purpose of intrusion detection, each record (e.g., network flow) has to be labelled  binary as benign or malicious traffic, or nominally to distinguish between certain attack types.
+For example, the UNSW-NB15 dataset defines a binary label (0 for normal and 1 for attack) and a nominal label ("Normal" for benign traffic and "Fuzzers", "Analysis", "Backdoors", "DoS", "Exploits", "Generic", "Reconnaissance", "Shellcode" or "Worms" for the various attack types) \cite{Nb2015}.
+For some applications, there may also be a nominal label for benign traffic (e.g., browsing, chat, mail, P2P, streaming, VoIP) \cite{bagui2017}.
+
+For data generated by simulated attacks, the labels can simply be based on the schedule of simulated attacks in combination with the IP addresses, ports and protocols involved \cite{Ids2018}.
+
 \begin{thebibliography}{00}
@@
 2020].
+\bibitem{Nb2015} Moustafa, N.
+(2015).
+The UNSW-NB15 Dataset Description.
+[online] Available at: \url{https://www.unsw.adfa.edu.au/unsw-canberra-cyber/cybersecurity/ADFA-NB15-Datasets/} [Accessed 22 Feb.
+2020]
+\bibitem{Ids2018} Communications Security Establishment (CSE) and Canadian Institute for Cybersecurity (CIC) (2018).
+CSE-CIC-IDS2018 on AWS.
+[online] Available at: \url{https://www.unb.ca/cic/datasets/ids-2018.html} [Accessed 22 Feb.
+2020].
+\bibitem{vasquez2014} Vázquez, F. Iglesias and Zseby, T. (2014). Analysis of network traffic features for anomaly detection. Machine Learning, vol. 101, 12 2014.
+\bibitem{este2009} Este, A.
+and Gringoli, F.
+and Salgarelli, L.
+(2009).
+On the stability of the information carried by traffic flow features at the packet level.
+SIGCOMM Comput.
+Commun.
+Rev., vol.
+39, p.
+13–18, June 2009.
+\bibitem{brownlee1998} Brownlee, N. (1998). Network management and realtime traffic flowmeasurement. J. Netw. Syst. Manage., vol. 6, p. 223–228, June 1998.
+\bibitem{waizumi2007} Waizumi, Y. and Sato, Y. and Nemoto, Y. (2007). A network-based anomaly detection system using multiple network features. WEBIST, 2007.
+\bibitem{celik2015} Celik, Z.
+Berkay and Walls, R.
+J.
+and McDaniel, P.
+and Swami, A.
+(2015).
+Malware traffic detection using tamper resistant features.
+MILCOM 2015 - 2015 IEEE Military Communications Conference, pp.
+330–335, Oct 2015.
+\bibitem{xu2015} Xu, M. and Zhu, W. and Xu, J. and Zheng, N. (2015). Towards selecting optimal features for flow statistical based network traffic classification. APNOMS 2015 - pp. 479–482, 08 2015.
+\bibitem{li2013} Li, W.
+M and Xie, S.
+L.
+and Luo, J.
+and Zhu, X.
+D.
+(2013).
+A detection method for botnet based on behavior features.
+Advanced Information and Computer Technology in Engineering and Manufacturing, Environmental Engineering, vol.
+765 of Advanced Materials Research, pp.
+1512–1517, Trans Tech Publications Ltd, 10 2013.
+\bibitem{abad2003} Abad, C.
+and Taylor, J.
+and Sengul, C.
+and Yurcik, W.
+and Zhou, Y.
+and Rowe, K.
+(2003).
+Log correlation for intrusion detection: a proof of concept.
+19th Annual Computer Security Applications Conference, 2003.
+Proceedings., pp.
+255–264, Dec 2003.
+\bibitem{bagui2017} Bagui, S.
+and Fang, X.
+and Kalaimannan, E.
+and Bagui, S.
+C.
+and Sheehan, J.
+(2017).
+Comparison of machine-learning algorithms for classification of vpn network traffic flow using time-related features.
+Journal of Cyber Security Technology, vol.
+1, no.
+2, pp.
+108–126, 2017.
 \end{thebibliography}
+
 \vspace{12pt}

### 2020-02-23 paper2/paper2.tex
Add pagebreak

--- 2020-02-22 paper2/paper2.tex
@@
 
+\newpage
+
 \begin{figure}[htbp]
@@
 
-If all the possible network scenarios are defined - this was the scope of another paper - this network topologies have to be deployed through our deployment system consisting of Vagrant and Ansible.
+If all the possible network scenarios are defined these network topologies have to be deployed through our deployment system consisting of Vagrant and Ansible.
 This means we will define the basic host setup and the network configurations for all the scenarios in Ansible templates which could be later adjusted via command line parameters.

### 2020-02-24 paper2/paper2.tex
Added some more content

--- 2020-02-23 paper2/paper2.tex
@@
 
+Other methods which can be used for that purpose are based on association rule mining (ARM) \cite{moustafa2015}, sequential forward selection (SFS) \cite{dehghani2010, xu2015}, or the T-test \cite{sopuru2019},
+
+When selecting the features, their limitations should be considered as well.
+Some features can easily be spoofed by attackers, while others are unable to detect encrypted malware traffic or cause performance problems.
+The development of new evasion techniques (e.g., disguising malicious traffic in widely used protocols such as HTTP) make the detection even harder \cite{celik2015}.
+
 Another aspect which has been examined is the stability of features in terms of time (year 2000/2001 vs.
@@
 In this regard, the packet size can be considered a stable feature \cite{este2009}.
+The suitability of features also depends on the type of network which is analyzed.
+For example, industrial control system networks are relatively static regarding IP address and port allocation, protocols and settings used, and number of connections, therefore features based on these attributes may provide good results in that particular environment \cite{mantere2013}.
 
@@
 
+The open-source Java software CICFlowMeter can be used to extract features of network flows (which end either by TCP FIN packet, or a configurable timeout) from pcap files.
+It allows to select from a list of available features, but can also be extended to support new features \cite{cicflowmeter}, and has therefore been used in various applications \cite{Ids2018, lashkari2017, sopuru2019}.
+
+\begin{figure}[htpb]
+\centerline{\includegraphics[scale=0.35]{hast-ids.png}}
+\caption{Feature learning process of HAST-IDS \cite{wang2018}}
+\label{hast-ids}
+\end{figure}
+
+The hierarchical spatial-temporal feature-based intrusion detection system (HAST-IDS), which is shown in fig.~\ref{hast-ids}, also extracts flow-based features.
+However, it first transforms the bytes of network flows (HAST-I) or packets (HAST-II) into images using one-hot encoding (OHE), and then uses deep convolutional neural networks (CNNs) and long short-term memory (LSTM) networks to learn features based on these images \cite{wang2018}.
+
 \subsection{Timeslot features}
@@
 Timeslot features are extracted by counting the number of certain events in fixed time intervals.
-For the detection of DoS and probing traffic, the following timeslot features are proposed in \cite{waizumi2007}:
+For the detection of DoS and probing traffic, the following timeslot features are proposed \cite{waizumi2007}:
 \begin{itemize}
@@
 A different approach is to define features which measure specific behavioral differences between benign and malicious software.
+
+One system following that approach is called "traffic aggregation for malware detection".
+It defines three characteristics which can be combined to distinguish between benign and malware traffic \cite{yen2011}:
+\begin{itemize}
+	\item Common destinations (assuming that compromised systems contact different hosts than uncompromised systems)
+	\item Similar payload (assuming that command-and-control traffic uses common protocol syntax)
+	\item Common software platforms (assuming that malware usually relies upon a specific operating system and also specific applications, such as web browsers)
+\end{itemize}
+
 The UNSW-NB15 datasets contains mostly flow statistical features, but also a few features based on behavior \cite{Nb2015}:
@@
 For example, the UNSW-NB15 dataset defines a binary label (0 for normal and 1 for attack) and a nominal label ("Normal" for benign traffic and "Fuzzers", "Analysis", "Backdoors", "DoS", "Exploits", "Generic", "Reconnaissance", "Shellcode" or "Worms" for the various attack types) \cite{Nb2015}.
-For some applications, there may also be a nominal label for benign traffic (e.g., browsing, chat, mail, P2P, streaming, VoIP) \cite{bagui2017}.
+For some applications, there may also be a nominal label for benign traffic (e.g., browsing, chat, mail, P2P, streaming, VoIP) \cite{lashkari2017, bagui2017}.
 
-For data generated by simulated attacks, the labels can simply be based on the schedule of simulated attacks in combination with the IP addresses, ports and protocols involved \cite{Ids2018}.
+For data generated by simulating attacks, the labels can simply be based on the schedule of simulated attacks in combination with the IP addresses, ports and protocols involved \cite{Ids2018}.
 
@@
 \bibitem{vasquez2014} Vázquez, F. Iglesias and Zseby, T. (2014). Analysis of network traffic features for anomaly detection. Machine Learning, vol. 101, 12 2014.
+\bibitem{moustafa2015} Moustafa, N.
+and Slay, J.
+(2015).
+The significant features of the UNSW-NB15 and the KDD99 data sets for Network Intrusion Detection Systems.
+2015 4th International Workshop on Building Analysis Datasets and Gathering Experience Returns for Security (BADGERS), Kyoto, pp.
+25-31, 2015.
+\bibitem{dehghani2010} Dehghani, F.
+and Movahhedinia, N.
+and Khayyambashi, M.
+R.
+and Kianian, S.
+(2010).
+Real-time Traffic Classification Based on Statistical and Payload Content Features.
+2010 2nd International Workshop on Intelligent Systems and Applications, Wuhan, pp.
+1-4, 2010.
+\bibitem{xu2015} Xu, M. and Zhu, W. and Xu, J. and Zheng, N. (2015). Towards selecting optimal features for flow statistical based network traffic classification. APNOMS 2015 - pp. 479–482, 08 2015.
+\bibitem{sopuru2019} Sopuru, J.
+and Sari, A.
+and Akkaya, M.
+(2019).
+Modeling A Malware Detection And Categorization System Based On Seven Network Flow-Based Features.
+International Journal of Innovative Technology and Exploring Engineering (IJITEE), vol.
+8(7), pp.
+2982-2989, 05 2019
+\bibitem{celik2015} Celik, Z.
+Berkay and Walls, R.
+J.
+and McDaniel, P.
+and Swami, A.
+(2015).
+Malware traffic detection using tamper resistant features.
+MILCOM 2015 - 2015 IEEE Military Communications Conference, pp.
+330–335, Oct 2015.
 \bibitem{este2009} Este, A.
@@
 13–18, June 2009.
+\bibitem{mantere2013} Mantere, M.
+and Sailio, M.
+and Noponen, S.
+(2013).
+Network Traffic Features for Anomaly Detection in Specific Industrial Control System Network.
+Future Internet, vol.
+5, pp.
+460-473, 2013
 \bibitem{brownlee1998} Brownlee, N. (1998). Network management and realtime traffic flowmeasurement. J. Netw. Syst. Manage., vol. 6, p. 223–228, June 1998.
 \bibitem{waizumi2007} Waizumi, Y. and Sato, Y. and Nemoto, Y. (2007). A network-based anomaly detection system using multiple network features. WEBIST, 2007.
-\bibitem{celik2015} Celik, Z.
-Berkay and Walls, R.
-J.
-and McDaniel, P.
-and Swami, A.
-(2015).
-Malware traffic detection using tamper resistant features.
-MILCOM 2015 - 2015 IEEE Military Communications Conference, pp.
-330–335, Oct 2015.
-\bibitem{xu2015} Xu, M. and Zhu, W. and Xu, J. and Zheng, N. (2015). Towards selecting optimal features for flow statistical based network traffic classification. APNOMS 2015 - pp. 479–482, 08 2015.
+\bibitem{cicflowmeter} Canadian Institute for Cybersecurity (CIC) (2016).
+Applications | Research | Canadian Institute for Cybersecurity | UNB.
+[online] Available at: \url{https://www.unb.ca/cic/research/applications.html#CICFlowMeter} [Accessed 24 Feb.
+2020].
+\bibitem{lashkari2017} Lashkari, A.
+H.
+and Gil, G.
+D.
+and Mamun, M.
+S.
+I.
+and Ghorbani, A.
+A.
+(2017).
+Characterization of Tor Traffic using Time based Features.
+Proceedings of the 3rd International Conference on Information Systems Security and Privacy (ICISSP 2017), pp.
+253-262, 2017.
+\bibitem{wang2018} Wang, W.
+and Sheng, Y.
+and Wang, J.
+and and Zeng, X.
+and Ye, X.
+and Huang, Y.
+and Zhu, M.
+(2018).
+HAST-IDS: Learning Hierarchical Spatial-Temporal Features Using Deep Neural Networks to Improve Intrusion Detection.
+IEEE Access, vol.
+6, pp.
+1792-1806, 2018.
+\bibitem{yen2011} Yen, T. (2011). Detecting stealthy malware using behavioralfeatures in network traffic. PhD thesis, Carnegie Mellon University, 2011.
 \bibitem{li2013} Li, W.

### 2020-02-24 paper2/paper2.tex
Fixed grammar
Fix grammar issues

--- 2020-02-23 paper2/paper2.tex
@@
 Collecting real network data for further processing is an important topic for the training of neuronal networks which filter suspicious from ordinary traffic.
-This paper focus mainly on research in the areas of network simulation, network recording and data extraction which should serve as input for the following papers.
+This paper focuses mainly on research in the areas of network simulation, network recording and data extraction which should serve as input for the following papers.
 \end{abstract}
@@
 \section{Introduction}
-Neuronal networks which filter suspicious network traffic need a training data set of high-quality.
+Neuronal networks which filter suspicious network traffic need a training data set of high quality.
 This data set could be produced from real network traffic but this can be difficult to implement.
@@
 At this point all our considerations are based on theoretical research  without practical implementation which will come to life in further work.
-Therefore this paper sets a direction and foundation for the next steps and shows possible solutions on a high-level.
+Therefore this paper sets a direction and foundation for the next steps and shows possible solutions on a high level.
 
@@
 There are many solutions available on the market to simulate networks.
-Popular network simulators which possibly fullfill the requirements above are for example VIRL, GNS3 or EVE-NG.
-This section focuses on the comparision of these simulators and the selection of the best solution for our use case.
+Popular network simulators which possibly fulfill the requirements above are for example VIRL, GNS3 or EVE-NG.
+This section focuses on the comparison of these simulators and the selection of the best solution for our use case.
 \cite{b1}
@@
 It offers the integration of virtual machines and real networks and supports models of various vendors but does not offer pre-installed device images due to license issues.
-It also provides a RESTful API to setup network topologies, a big community and scalibility.
+It also provides a RESTful API to setup network topologies, a big community and scalability.
 The software can be easily installed through packages on all operating systems and supports all hypervisors as well Docker containers.
@@
 It supports multiple vendors but does not provide device images.
-Advanced features like Docker, configuration management or Wireshark integration only comes with the purchase version.
+Advanced features like Docker, configuration management or Wireshark integration only come with the purchase version.
 Moreover a RESTful API and images for the installation in virtual machines are available as well as a reasonable documentation.
-The software is scalable and performs well in virtual machines with a short topology loading time and provides a intuitive interface as shown in Fig.~\ref{eve-ng}.
+The software is scalable and performs well in virtual machines with a short topology loading time and provides an intuitive interface as shown in Fig.~\ref{eve-ng}.
 \cite{b1} \cite{b7}
@@
 
-Next steps include the integration of the selected network simulator in an scalable environment which can be setup easily.
+Next steps include the integration of the selected network simulator in a scalable environment which can be set up easily.
 

### 2020-02-24 paper2/paper2.tex
Separate multiple citations by space instead of comma

--- 2020-02-24 paper2/paper2.tex
@@
 
-Other methods which can be used for that purpose are based on association rule mining (ARM) \cite{moustafa2015}, sequential forward selection (SFS) \cite{dehghani2010, xu2015}, or the T-test \cite{sopuru2019},
+Other methods which can be used for that purpose are based on association rule mining (ARM) \cite{moustafa2015}, sequential forward selection (SFS) \cite{dehghani2010} \cite{xu2015}, or the T-test \cite{sopuru2019},
 
@@
 Internet edge).
-In this regard, the packet size can be considered a stable feature \cite{este2009}.
+In this regard, the packet size can be considered as stable feature \cite{este2009}.
 The suitability of features also depends on the type of network which is analyzed.
@@
 \begin{itemize}
-	\item Flow duration \cite{Nb2015, Ids2018}
+	\item Flow duration \cite{Nb2015} \cite{Ids2018}
 	\item Time between two flows or packets \cite{Ids2018}
@@
 	\item Jitter \cite{Nb2015}
-	\item Number of packets in total, with at least 1 byte of payload, or with certain protocol flags \cite{Ids2018, waizumi2007, celik2015}
+	\item Number of packets in total, with at least 1 byte of payload, or with certain protocol flags \cite{Ids2018} \cite{waizumi2007} \cite{celik2015}
 	\item Number of fragmented packets \cite{waizumi2007}
 	\item Number of retransmitted or dropped packets \cite{Nb2015}
-	\item Size of packets, payload, headers or segments \cite{Ids2018, celik2015}
-	\item Number of bytes/packets transferred per second, in bulk, or in the initial TCP window \cite{Ids2018, celik2015}
+	\item Size of packets, payload, headers or segments \cite{Ids2018} \cite{celik2015}
+	\item Number of bytes/packets transferred per second, in bulk, or in the initial TCP window \cite{Ids2018} \cite{celik2015}
 	\item Number of RTT samples \cite{celik2015}
@@
 	\item Ratio between maximum and minimum packet size \cite{celik2015}
-	\item Number of flows with the current flow's IP address or port number \cite{waizumi2007, Nb2015}
+	\item Number of flows with the current flow's IP address or port number \cite{waizumi2007} \cite{Nb2015}
 \end{itemize}
-Most of these features can be derived by statistical calculations such as total, minimum, maximum, mean, median, standard deviation, variance, skewness and kurtosis \cite{Ids2018, celik2015, xu2015}.
+Most of these features can be derived by statistical calculations such as total, minimum, maximum, mean, median, standard deviation, variance, skewness and kurtosis \cite{Ids2018} \cite{celik2015, xu2015}.
 Also, the features can be extracted from bidirectional flows, or separately for forward and backward directions \cite{Ids2018}.
@@
 The open-source Java software CICFlowMeter can be used to extract features of network flows (which end either by TCP FIN packet, or a configurable timeout) from pcap files.
-It allows to select from a list of available features, but can also be extended to support new features \cite{cicflowmeter}, and has therefore been used in various applications \cite{Ids2018, lashkari2017, sopuru2019}.
+It allows to select from a list of available features, but can also be extended to support new features \cite{cicflowmeter}, and has therefore been used in various applications \cite{Ids2018} \cite{lashkari2017} \cite{sopuru2019}.
 
@@
 For example, the UNSW-NB15 dataset defines a binary label (0 for normal and 1 for attack) and a nominal label ("Normal" for benign traffic and "Fuzzers", "Analysis", "Backdoors", "DoS", "Exploits", "Generic", "Reconnaissance", "Shellcode" or "Worms" for the various attack types) \cite{Nb2015}.
-For some applications, there may also be a nominal label for benign traffic (e.g., browsing, chat, mail, P2P, streaming, VoIP) \cite{lashkari2017, bagui2017}.
+For some applications, there may also be a nominal label for benign traffic (e.g., browsing, chat, mail, P2P, streaming, VoIP) \cite{lashkari2017} \cite{bagui2017}.
 

### 2020-02-24 paper2/paper2.tex

--- 2020-02-24 paper2/paper2.tex
@@
 Because we want an affordable solutions which can be used in most of the cases also on our computers with some exceptions, we decided to use a tool which can create virtual machines as well as Docker containers to be ready for the future.
-Regarding the corner cases servers or a Kubernetes cluster could be rent temporary on cloud platforms to support high-load scenarios e.g.
+Regarding the corner cases servers or a Kubernetes cluster could be rented temporarily on cloud platforms to support high-load scenarios e.g.
 big denial-of-service attacks.
 
-A popular solution to bootstrap development environments in virtual machines is called Vagrant which provides lightweight base images and an interface automate the configuration of the virtual machines.
+A popular solution to bootstrap development environments in virtual machines is called Vagrant which provides lightweight base images and an interface to automate the configuration of the virtual machines.
 Moreover it can be connected to provisioning tools to further configure host systems.
 Vagrant supports VirtualBox or VMware but also Docker and can be extended to support cloud services like AWS or even to install a whole Kubernetes cluster.
-It gives us the power to start with simple configuration and scale them up to more sophisticated ones using containerization and complex orchestrators.
+It gives us the power to start with simple configurations and scale them up to more sophisticated ones using containerization and complex orchestrators.
 \cite{b9}
@@
 
-Vagrant supports multiple provisioners like Ansible, Terraform or simple bash scripts to configure the host systems and can also be setup through them.
+Vagrant supports multiple provisioners like Ansible, Terraform or simple bash scripts to configure the host systems and can also be set up through them.
 For the beginning we will use Ansible through Vagrant to configure the host system.
 Vagrant can setup multiple virtual machines or Docker containers on one host.
-If we would need to setup machines on multiple hosts we will come back to tools like Terraform for example.
+If we would need to set up machines on multiple hosts we will come back to tools like Terraform for example.
 \cite{b9}
@@
 This includes the deployment of the network recorder and the data extractor.
-In other words one or multiple scenarios can be started and connected to one recorder and data extractor to combine simultanuous attacks and everything will be deployed through a shell command with static configuration in templates and dynamic configuration with command line options.
+In other words one or multiple scenarios can be started and connected to one recorder and data extractor to combine simultaneous attacks and everything will be deployed through a shell command with static configuration in templates and dynamic configuration with command line options.
 It will be also possible to integrate real hardware into the simulation.
@@
 A real test case would then consist of multiple scenarios of attacks and usual network traffic recorded, extracted and collected in a central place.
-The next chapter focus on recording the generated data.
+The next chapter focuses on recording the generated data.
 

### 2020-02-25 paper2/paper2.tex
Improved some sentences

--- 2020-02-24 paper2/paper2.tex
@@
 There also exists a large number of techniques for dimensionality reduction.
-The 40 features of the NSL-KDD data set could be reduced to 30 or 16 features, with no or low degradation in detection power, while highly reducing computation costs, by combining the following methods \cite{vasquez2014}:
+The 40 features of the NSL-KDD data set could be reduced to 30 or 16 features, with no or low degradation in detection power.
+The benefits of a reduced feature set are faster processing, less memory consumption, and in some cases even higher detection precision, because some classifiers have problems with irrelevant or correlated features.
+This could be achieved by combining the following methods \cite{vasquez2014}:
 \begin{itemize}
@@
 The suitability of features also depends on the type of network which is analyzed.
-For example, industrial control system networks are relatively static regarding IP address and port allocation, protocols and settings used, and number of connections, therefore features based on these attributes may provide good results in that particular environment \cite{mantere2013}.
+For example, some industrial control system networks are relatively static regarding IP address allocation, the number of open connections and the ports, protocols and settings they use.
+Therefore, features based on these attributes may provide good results in that particular environment \cite{mantere2013}.
 
@@
 A common approach is to group packets into flows, and then calculate statistical features based on these flows.
-Generally, a flow is a group of packets with the same source and destination IP addresses and ports, as well as protocol \cite{brownlee1998}.
+Generally, a flow is a group of packets with the same source and destination IP addresses and ports, and the same protocol \cite{brownlee1998}.
 However, a flow can also be terminated by a TCP FIN packet or a timeout \cite{Ids2018}.

### 2020-02-25 paper2/paper2.tex

@@
 
-Vagrant supports multiple provisioners like Ansible, Terraform or simple bash scripts to configure the host systems and can also be setup through them.
+Vagrant supports multiple provisioners like Ansible, Terraform or simple bash scripts to configure the host systems and can also be set up through them.
 For the beginning we will use Ansible through Vagrant to configure the host system.
 Vagrant can setup multiple virtual machines or Docker containers on one host.
-If we would need to setup machines on multiple hosts we will come back to tools like Terraform for example.
+If we would need to set up machines on multiple hosts we will come back to tools like Terraform for example.
 \cite{b9}
 
-If all the possible network scenarios are defined - this was the scope of another paper - this network topologies have to be deployed through our deployment system consisting of Vagrant and Ansible.
+If all the possible network scenarios are defined these network topologies have to be deployed through our deployment system consisting of Vagrant and Ansible.
 This means we will define the basic host setup and the network configurations for all the scenarios in Ansible templates which could be later adjusted via command line parameters.

### 2020-02-25 paper2/paper2.tex

@@
 The benefits of a reduced feature set are faster processing, less memory consumption, and in some cases even higher detection precision, because some classifiers have problems with irrelevant or correlated features.
-This could be achieved by combining the following methods \cite{vasquez2014}:
+This could be achieved by combining the following methods \cite{vasquez2015}:
 \begin{itemize}
@@
 2020].
-\bibitem{vasquez2014} Vázquez, F. Iglesias and Zseby, T. (2014). Analysis of network traffic features for anomaly detection. Machine Learning, vol. 101, 12 2014.
+\bibitem{vasquez2015} Vázquez Iglesias, F. and Zseby, T. (2015). Analysis of network traffic features for anomaly detection. Machine Learning, vol. 101, 2015.
 \bibitem{moustafa2015} Moustafa, N.
@@
 (2010).
-Real-time Traffic Classification Based on Statistical and Payload Content Features.
+Real-time traffic classification based on statistical and payload content features.
 2010 2nd International Workshop on Intelligent Systems and Applications, Wuhan, pp.
@@
 (2019).
-Modeling A Malware Detection And Categorization System Based On Seven Network Flow-Based Features.
+Modeling a malware detection and categorization system based on seven network flow-based features.
 International Journal of Innovative Technology and Exploring Engineering (IJITEE), vol.
@@
 (2013).
-Network Traffic Features for Anomaly Detection in Specific Industrial Control System Network.
+Network traffic features for anomaly detection in specific industrial control system network.
 Future Internet, vol.
@@
 460-473, 2013
-\bibitem{brownlee1998} Brownlee, N. (1998). Network management and realtime traffic flowmeasurement. J. Netw. Syst. Manage., vol. 6, p. 223–228, June 1998.
-\bibitem{waizumi2007} Waizumi, Y. and Sato, Y. and Nemoto, Y. (2007). A network-based anomaly detection system using multiple network features. WEBIST, 2007.
+\bibitem{brownlee1998} Brownlee, N. (1998). Network management and realtime traffic flowmeasurement. J. Netw. Syst. Manage., vol. 6, pp. 223–228, June 1998.
+\bibitem{waizumi2007} Waizumi, Y.
+and Sato, Y.
+and Nemoto, Y.
+(2007).
+A network-based anomaly detection system using multiple network features.
+Proceedings of the Third International Conference on Web Information Systems and Technologies - Internet Technology, pp.
+410-413, 2007.
 \bibitem{cicflowmeter} Canadian Institute for Cybersecurity (CIC) (2016).
@@
 (2017).
-Characterization of Tor Traffic using Time based Features.
+Characterization of tor traffic using time based features.
 Proceedings of the 3rd International Conference on Information Systems Security and Privacy (ICISSP 2017), pp.
@@
 (2018).
-HAST-IDS: Learning Hierarchical Spatial-Temporal Features Using Deep Neural Networks to Improve Intrusion Detection.
+HAST-IDS: Learning hierarchical spatial-temporal features Using deep neural networks to improve intrusion detection.
 IEEE Access, vol.
@@
 1792-1806, 2018.
-\bibitem{yen2011} Yen, T. (2011). Detecting stealthy malware using behavioralfeatures in network traffic. PhD thesis, Carnegie Mellon University, 2011.
+\bibitem{yen2011} Yen, T. (2011). Detecting stealthy malware using behavioral features in network traffic. PhD thesis, Carnegie Mellon University, 2011.
 \bibitem{li2013} Li, W.

### 2020-02-25 paper2/paper2.tex
Modify expressions and cites

@@
 It offers the integration of virtual machines and real networks and supports models of various vendors but does not offer pre-installed device images due to license issues.
-It also provides a RESTful API to setup network topologies, a big community and scalability.
+It also provides a RESTful API to set up network topologies, a big community and scalability.
 The software can be easily installed through packages on all operating systems and supports all hypervisors as well Docker containers.
@@
 Regarding these factors GNS3 fullfills our requirements the most because it is free, supports all vendors and comes with a big community and documentation.
-Our second option would be EVE-NG in the professional edition which also supports all vendors.
+Our second option could be EVE-NG in the professional edition which also supports all vendors.
 
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[scale=0.25]{docker_vm.png}}
+\centerline{\includegraphics[scale=0.5]{docker_vm.png}}
 \caption{Virtual machines vs. Containers \cite{b8}}
@@
 
-Hypervisors like VirtualBox or VMware provide the possibility to setup full virtual computers on a physical host where Docker containers are encapsuled sandboxes on an existing operating system like illustrated in Fig.~\ref{docker-vm}.
+Hypervisors like VirtualBox or VMware provide the possibility to set up full virtual computers on a physical host where Docker containers are encapsuled sandboxes on an existing operating system like illustrated in Fig.~\ref{docker-vm}.
 It is clear that containers are more resource efficent but in virtual machines the systems are better isolated.
@@
 If we focus on resource efficency and scalability also Kubernetes comes in our minds which is an orchestrator for deploying Docker containers on multiple nodes but it is a more complex configuration overhead and needs certain amount of hardware to unfold its potential also for simple scenarios.
-Because we want an affordable solutions which can be used in most of the cases also on our computers with some exceptions, we decided to use a tool which can create virtual machines as well as Docker containers to be ready for the future.
-Regarding the corner cases servers or a Kubernetes cluster could be rented temporarily on cloud platforms to support high-load scenarios e.g.
+Because we want an affordable solutions which can be used in most of the cases also on our computers with some exceptions, we could use a tool which can create virtual machines as well as Docker containers to be ready for the future.
+Regarding the corner cases servers or a Kubernetes cluster could be rented temporarily on cloud platforms like AWS to support high-load scenarios e.g.
 big denial-of-service attacks.
+\cite{b8}
 
@@
 It gives us the power to start with simple configurations and scale them up to more sophisticated ones using containerization and complex orchestrators.
-\cite{b9}
+\cite{b9} \cite{b10}
 
-This means we are ready to create system environments on our host or in the cloud and install base systems on them.
+This means we would be ready to create system environments on our host or in the cloud and install base systems on them.
 The next section focuses on the automated setup of the network simulator, recorder and data extractor within virtual machines or Docker containers as well as the configuration of network topologies based on the selected scenarios and parameters to produce, record and extract network traffic for further processing in neuronal networks to improve their strike rate.
@@
 Vagrant supports multiple provisioners like Ansible, Terraform or simple bash scripts to configure the host systems and can also be set up through them.
-For the beginning we will use Ansible through Vagrant to configure the host system.
-Vagrant can setup multiple virtual machines or Docker containers on one host.
-If we would need to set up machines on multiple hosts we will come back to tools like Terraform for example.
-\cite{b9}
+For the beginning we could use Ansible through Vagrant to configure the host system.
+Vagrant can set up multiple virtual machines or Docker containers on one host.
+If we would need to set up machines on multiple hosts we could come back to tools like Terraform for example.
+\cite{b9} \cite{b10} \cite{b11}
 
-If all the possible network scenarios are defined these network topologies have to be deployed through our deployment system consisting of Vagrant and Ansible.
-This means we will define the basic host setup and the network configurations for all the scenarios in Ansible templates which could be later adjusted via command line parameters.
+If all the possible network scenarios are defined these network topologies have to be deployed through our deployment system.
+This means we could define the basic host setup and the network configurations for all the scenarios in Ansible templates which could be later adjusted via command line parameters.
 The topology can then be deployed by executing a shell command.
@@
 It will be also possible to integrate real hardware into the simulation.
+\cite{b9}
 
@@
 
-Fig.~\ref{docker-vm} shows a schematic overview of the simulation workflow.
+Fig.~\ref{docker-vm} shows a possible simulation workflow.
 The input parameters are the type of scenario(s) and the respective settings which deploys hosts including the base systems and the simulators applying the network topologies.
-Moreover the network recorder and data extractor will be deployed to gather the network traffic and extract the features to create data sets which could be collected in a central database.
+Moreover the network recorder and data extractor could be deployed to gather the network traffic and extract the features to create data sets which could be collected in a central database.
 
@@
 
-\subsection*{Label definition}
+A different approach is the use of features derived from raw system call traces.
+The advantage of system call traces over log files is that they contain only raw information about the interaction between programs and the operating system kernel, whereas log files contain data interpreted by various programs, and potentially include large amounts of irrelevant data.
+There are multiple methods of analysing system calls.
+Some only look at the pattern of system calls, while others also look at the arguments passed to these system calls.
+Also, there is a distinction between syntactic features, which depend only on contiguous patterns of system calls, and semantic features, which additionally consider the meaning of system call arguments or groups of discontiguous system calls.
+The advantage of the semantic approach is that it is able to counter mimicry attacks, where the payload is modified by the attacker with the aim that the system calls it executes cannot be classified as being malicious \cite{creech2014}.
+
+\subsection{Label definition}
 
@@
 
+\section{Conclusion}
+
+The quality of a dataset is generally crucial for the performance of machine learning algorithms, and generating a high quality dataset for intrusion detection systems is not a simple task.
+While the generation of representative benign and malicious data in itself is fundamental, the goal of this paper was to give an overview about the equally important tasks of network simulation, recording and data extraction.
+For each of these tasks, a variety of different approaches has been examined.
+Because each approach has its own strengths and limitations, they have to be evaluated individually for each application, depending on its specific requirements.
+Also, the further development of the Internet and the ongoing introduction of new attack techniques has to be considered, which may affect the suitability of current methods in the future and therefore cause the necessity to alter them or to develop new ones.
+
 \begin{thebibliography}{00}
-\bibitem{b1} CBT Nuggets Blog (2020).
-5 Best Network Simulators for Cisco Exams: CCNA, CCNP, CCIE.
-[online] Available at: \newline\url{https://www.cbtnuggets.com/blog/career/career-progression/5-best-network-simulators-for-cisco-exams-ccna-ccnp-and-ccie} [Accessed 22 Feb.
-2020].
+\bibitem{b1} Seifert, C.
+and Reißmann, S.
+and Rieger, S.
+and Pape, C.
+Evaluation von VIRL, GNS3 und Mininet als Virtual Network Testbeds in der Hochschullehre.
+11.
+DFN-Forum Kommunikationstechnologien.
+Bonn: Gesellschaft für Informatik e.V.., pp.
+103-112.
 \bibitem{b2} Network Computing.
@@
 \bibitem{b7} Eve-ng.net. (2020). EVE-NG API. [online] Available at: \url{https://www.eve-ng.net/index.php/documentation/howtos/how-to-eve-ng-api/} [Accessed 22 Feb. 2020].
-\bibitem{b8} SOSC-2018. (2020). Docker. [online] Available at: \url{https://dodas-ts.github.io/SOSC-2018/docker.html} [Accessed 22 Feb. 2020].
-\bibitem{b9} Chan, M.
-(2020).
-15 Infrastructure as Code Tools to Automate Deployments - Thorn Tech.
-[online] Thorn Technologies.
-Available at: \url{https://www.thorntech.com/2018/04/15-infrastructure-as-code-tools/} [Accessed 22 Feb.
-2020].
+\bibitem{b8} Bernstein, D. Containers and Cloud: From LXC to Docker to Kubernetes, in IEEE Cloud Computing, vol. 1, no. 3, pp. 81-84, Sept. 2014.
+\bibitem{b9} Nishant Kumar Singh, S.
+Thakur, H.
+Chaurasiya and H.
+Nagdev, Automated provisioning of application in IAAS cloud using Ansible configuration management, 1st International Conference on Next Generation Computing Technologies (NGCT), Dehradun, 2015, pp.
+81-85.
+\bibitem{b10} Spanaki, P. Cloud Computing: Security Issues and Establishing Virtual Cloud Environment via Vagrant to Secure Cloud Hosts, 2018, pp. 539-553
+\bibitem{b11} Ivanova, D. and Borovska, P. and Zahov, S. Development of PaaS using AWS and Terraform for medical imaging analytics, AIP Conference Proceedings, 2018
 \bibitem{Nb2015} Moustafa, N.
@@
 255–264, Dec 2003.
+\bibitem{creech2014} Creech, G.
+and Hu, J.
+(2014).
+A semantic approach to host-based intrusion detection systems using contiguous and discontiguous system call patterns.
+IEEE Transactions on Computers, vol.
+63, no.
+4, pp.
+807-819, April 2014.
 \bibitem{bagui2017} Bagui, S.

### 2020-02-25 paper1/paper1.tex
added parts about tools and started with realistic workloads

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Network packet generation for Artificial Intelligence\\
}

\author{\IEEEauthorblockN{ Bernhard Gally}
\IEEEauthorblockA{\textit{IT Security} \\
\textit{FH Technikum Wien}\\
Vienna, Austria\\
cs19m023@technikum-wien.at}
\and
\IEEEauthorblockN{Maximilian Wech}
\IEEEauthorblockA{\textit{IT Security} \\
\textit{FH Technikum Wien}\\
Vienna, Austria\\
cs19m020@technikum-wien.at}
}

\maketitle

\begin{abstract}
Supervised learning in artificial intelligence requires a training dataset that is already labeled. Doing so with network packets is a challenge due to the fact that even Intrusion Detection System are not always correct with their guessing. To overcome this problem it is necessary to generate malicious and benign packets and label them accordingly. To increase the quality of the dataset the ip addresses from the machines should not remain the same for all of the data and the malicious and benign traffic should be mixed.

\end{abstract}

\begin{IEEEkeywords}
artificial intelligence, neuronal networks, computer networks, network security, packet generation
\end{IEEEkeywords}

\section{Introduction}
Generating network traffic that can be used to train a neuronal network brings some challenges: there are many different attack types that need to be considered, a vast variety of attack tools to generate malicious traffic are available in the Internet each with their own advantages and disadvantages and generating realistic network traffic that contains that does not get mislabeled by intrusion detection systems just to name a few.

This paper is split into three chapters. The first chapter explains the different kinds of attacks and their targets. The second chapter focuses on tools to generate malicious traffic. The third chapter addresses the problem of generating useful network traffic.

At this point all of our considerations are based on theoretical research without practical implementation which will come to life in further work. Therefore this paper sets a direction and foundation for the next steps and shows possible solutions at a high level.

\section{Cyber-attacks: types, targets, vulnerabilites}
\label{sec:cattv}

This chapter intends to provide an overview of various types of cyber-attacks, cyber-attack targets and exploitable vulnerabilities.

\subsection{General information}
In order to fulfil the task, it initially appears necessary to describe typical characteristics of attacks. Attacks aim to exploit weaknesses, such as poorly configured systems, or errors in software and to circumvent security measures. Attackers deliberately try to make services unavailable, spy on or manipulate data and disrupt normal traffic as much as possible \cite{b1}. Executing attacks generally consists of several steps: First, the attacker's motivation must be clarified (e.g. proof of your own abilities). Then it is necessary to select a target and collect information about it. As soon as this has happened, the attack method is selected and the attack is performed \cite{b2}.

\subsection{Classifying attacks}

\medskip Basically, there are too many different cyber-attack methods to mention all of them. Therefore, various approaches exist in the literature to classify attacks into different groups. One way is to split attacks into active and passive ones \cite{b3,b4}.  In the case of passive attacks, attempts are not made to manipulate data in network traffic, but rather to eavesdrop and collect sensitive information. The opposite is the case with active attacks, which attempt to manipulate network packets, for example \cite{b4}.  Further approaches exist such as classifying into internal and external attacks, or classifying attacks based on the underlying OSI-layer.

\medskip In the literature often cited is the classification model presented in \cite{b5}, which works on the basis of four dimensions and offers a holistic view of attacks. First, an attack class is assigned based on the attack vector. The attacks target is specified in the second dimension. The vulnerabilities that are exploited by the attack are then noted. Finally, the fourth dimension analyses whether an attack has a payload or an effect that exceeds the typical characteristics of the type of attack \cite{b5} . The rest of this chapter will focus on further introducing this model and consider each dimension individually.

\subsubsection{First dimension}

\medskip As already mentioned, in the first dimension the attack class is determined on the basis of the attack vector. The following generic classifications are listed in \cite{b5}:

\smallskip Virus: this is a malware that replicates itself and, for example, infects other script code. When the infected source code is executed, the virus is also executed \cite{b6}.  Basically there are many different types of viruses. An example would be File Infectors, in which a virus is embedded in a file. If this file is executed, the virus is also executed \cite{b2}.

\smallskip Worms: a software that spreads independently by exploiting vulnerabilities \cite{b6}. Examples include mass mailing worms, which spread via email, and network-aware worms, which only require four steps to infect a foreign host and spread themselves \cite{b2}.

\smallskip Trojans: a software that appears harmless at first, but also contains a hidden, malicious function \cite{b6}. One example are logic bombs, which only execute their malicious payload when a specific condition (e.g. a certain time of a day) has been reached \cite{b2}.

\smallskip Buffer overflows: This type of attack is based on programming errors and is very often used to attack computers or networks. Principally an attempt is made to overfill a temporary memory area. The amount of data that exceeds the size of the memory overwrites and corrupts other data. One possible goal of this attack would be to take control of a process. The main types of this attack are stack and heap buffer overflows \cite{b2}.

\smallskip Denial of service attacks: this means attacks on the availability of systems.  An attempt is made to exhaust the resources required to run a service \cite{b6}. This attack can be host-based, network-based, or distributed. The latter version is a DDoS attack, in which several attackers attack a mutual target at the same time \cite{b2}.

\smallskip Network attacks: the category is divided into several sub-areas. If an attacker pretends to be someone else by manipulating certain addresses, this is called spoofing. This can take place on the basis of the MAC address (MAC Address Spoofing) or on the basis of the IP address (IP Spoofing). An attack in which IP spoofing is used is session hijacking. Here, an attacker tries to link into a session (based, for example, on TCP / Telnet) between two hosts and virtually replace one host. In the case of wireless network attacks, weaknesses such as too short initialization vectors of the WEP standard can be exploited. In the event of attacks on web applications, attempts are made to exploit security gaps in them.  For example, if there is no strict separation between user input and program instructions, or there is insufficient input validation, an attacker can successfully send manipulated SQL statements to the underlying database. The consequences would be spying out of data or losing control of the underlying server. \cite{b2}.

\smallskip Physical attacks: this type of attack aims to cause physical damage to target systems (i.e. computers or networks). This includes, for example, deliberately cutting network cables or destroying computers \cite{b2}.
Password attacks: here an attempt is made to find out a specific password in order to gain control over a PC, for example. This can be done in three different ways: trying out all possible password combinations (brute force), trying out a part of all possible password combinations (e.g. testing simple passwords) and determining deficiencies in password storage / encryption \cite{b2,b5}.

\smallskip Information gathering attacks: this type of attack does not change data; it is only about spying on (sensitive) information, which can be used for follow-up attacks \cite{b5}. A possible variant of this is sniffing of packets, to find out about IP and MAC addresses or passwords \cite{b2}.

\medskip \subsubsection{Second dimension}The targets of attacks are described in the second dimension of the classification model. These should be described as specifically as possible, which means that it is not sufficient, for example, to specify which server was attacked, but rather which service or which operating system is affected with which version. It should also be noted that an attack can target more than one target. Basically three groups of attack targets are defined \cite{b5}:

\smallskip Hardware: this group is divided into three areas. First of all, computers and their components (e.g. processors, random access memory) should be mentioned. Network devices such as routers or switches also belong to this category. Ultimately, peripheral devices such as monitors and keyboards are also included.

\smallskip Software: here a distinction is made between targets on operating systems (Windows, Linux, etc.) and various applications (email client, text processing program) that are executed on an operating system.

\smallskip Network: in this group, the target is not software or hardware but the network itself, or an associated network protocol (for example, TCP).

\medskip \subsubsection{Third dimension}In the third dimension of the classification model \cite{b5}, vulnerabilities that can be exploited by attacks are considered. Reference is made to the industry standard CVE - Common Vulnerabilities and Exposures \cite{b7}, which is used for the uniform naming of security gaps and vulnerabilities in computer systems. General vulnerability types are also defined in\cite{b8}:

\smallskip Implementation: if a system is not implemented according to the design, security gaps arise.

\smallskip Design: an incorrect design leads to a security-threatening implementation.

\smallskip Configuration: If a system is not configured correctly, doors are opened to the attackers.

\medskip \subsubsection{Fourth dimension}The last dimension analyzes whether an attack has a payload or an effect that exceeds the typical characteristics of the type of attack. This should result in a detailed view of the payload. For example, a worm's payload might be a Trojan, or the worm might just be manipulating files \cite{b5}.

\medskip The classification model offers a comprehensive view of attacks. For example, for a worm that is distributed via email (dimension 1), the target Windows 95 (dimension 2) with an inadequate configuration (dimension 3) and a virus as payload (dimension 4) is determined. Further dimensions could be specified, for example what damage is caused by an attack, or how many costs arise from successfully carrying out the attack. Overall an overview is provided of how attacks can be classified, which attack targets are possible, or which vulnerabilities exist.

\section{Tools}

There is a large number of network tools available in the world wide web. Many of them are free to download and can be used for malicious activities but there are also many that are for defensive purposes. Generally speaking there are two major groups: tools for network attackers and tools for network defenders.\cite{b1} The tools for network attackers are the ones generating the packets thus being the ones we will focus on.

\begin{figure}[htbp]  
\centerline{\includegraphics[scale=.55]{network-tools.jpg}}
\caption{taxonomy of network security tools \cite{b1}}
\end{figure}

\subsection{Information gathering}
Attackers usually collect information before starting their attack. Understanding the environment, what operating systems are used, finding open ports and available services are only a small part of information that can be obtained.

\smallskip Table \ref{table:sniffing-tools} contains a list of sniffing tools that are available online. Generally speaking sniffing tools do not generate packets and are therfore undetectable. However, on a switched ethernet network segment a sniffer does generate a small amount of packets such as ARP\cite{b9}. Other types of sniffers, such as Ettercap, can execute a Man-in-the-middle attack. This causes a lot of packets to be sent through the network as it needs to answer requests in order to spoof an identity.

\begin{table}[htbp]
\caption{Sniffing tools \cite{b1}} 
\centering
\begin{tabular}{ | c | c | }
\hline
Tool name & Purpose \\
\hline
Ethereal & Packet capturing \\
Tcpdump & Packet capturing  \\
Net2pcap & Packet capturing  \\
Snoop & Packet capturing \\
Angst & Sniffing  \\
Ngrep & Capturing packet  \\
Ettercap & Man-in-the-middle attack  \\
Dsniff & Password sniffing  \\
Cain \& able & Password recovery  \\
Aimsniff & Capturing packet \\
Tcptrace & Analysis of traffic \\
Tcptrack & TCP connection analysis  \\
Argus & Analysis of audit data \\
Karpski & Packet analyzer \\
IPgrab & Display packet header \\
Nast & Traffic analysis \\
Gulp & Packet capturing/visualization \\
Libpcap & Packet capturing \\
Nfsen & Flow capturing/ visualization \\
Nfdump & Flow capturing/visualization \\
\hline
\end{tabular}
\label{table:sniffing-tools}
\end{table}

Scanning tools are often more intrusive than sniffing tools and generate lots of packets that can be detected. Their aim is to show up open ports, available services, used operating systems. With the gathered information an attacker can pick a suitable attack vector.
Table \ref{table:scanning-tools} shows a few common network scanning tools.

\begin{table}[htbp]
\caption{Scanning tools \cite{b1}} 
\centering
\begin{tabular}{ | c | c | }
\hline
Tool name & Purpose \\
\hline
Nmap & Scanning	 \\
Amap & Scanning \\
Vmap & Version mapping \\
Unicornscan & Scanning \\
Ttlscan & Scanning \\
Ike-scan & Host discovery \\
Paketto & Scanning \\
\hline
\end{tabular}
\label{table:scanning-tools}
\end{table}

\subsection{Attack launching}
The most packets however are generated by the attacks themselves.
Especially some kinds of DoS attacks are designed to overload the network with packets such that the victim cannot respond to legitimate requests.

\subsubsection{Denial of service}
The Internet provides many tools that are capable of executing DoS attacks and can be easily used to crash hosts or even networks.
Two examples of such tools are Low Orbit Ion Cannon (LOIC) or High Orbit Ion Cannon (HOIC). Both of these tools are capable of launching DoS/DDoS attacks within a short amount of time.
LOIC supports TCP, UDP and HTTP whereas HOIC only supports the HTTP protocol\cite{b1}. A list of DoS and DDoS tools is shown in Table \ref{table:ddos-tools}.

\begin{table}[htbp]
\caption{DoS/DDoS tools \cite{b1}} 
\centering
\begin{tabular}{ | c | c | }
\hline
Tool name & Purpose \\
\hline
Jolt & DoS \\
Burbonic & DoS \\
Targa & DoS \\
Blas20 & DoS \\
Crazy Pinger & DoS \\
UDPFlood & DoS \\
FSMax & DoS \\
Nemsey & DoS \\
Panther & DoS \\
Land \& LaTierra & DoS \\
Slowloris & DoS \\
Blackenergy & DDoS \\
HOIC & DDoS \\
Shaft & DDoS \\
Knight & DDoS \\
Kaiten & DDoS \\
RefRef & DDoS \\
Hgod & DDoS \\
LOIC & DDoS \\
Trinoo & DDoS \\
TFN & DDoS \\
TFN2K & DDoS \\
Stachaldraht & DDoS \\
Mstream & DDoS \\
Trinity & DDoS \\
\hline
\end{tabular}
\label{table:ddos-tools}
\end{table}

\smallskip
\subsubsection{Packet forging}
Packet forging tools are necessary to manipulate packet information. With the help of such tools it is possible to generate traffic with non-existant IP addresses. They also allow to individually set the flags of network packets to a desired state. An extreme example of such a packet is the "Christmas tree packet", which has all available flags set to 1. A list of packet forging tools is shown in Table \ref{table:packetforging-tools}.

\begin{table}[htbp]
\caption{Packet forging tools \cite{b1}} 
\centering
\begin{tabular}{ | c | c | c |}
\hline
Tool name & Purpose \\
\hline
Packeth & Packet generator	 \\
Packit & Packet analysis and injection	 \\
Nemesis & Packet crafting/injection	 \\
Tcpinject & Packet generator	 \\
Libnet & Packet injection	 \\
SendIP & Packet generator	 \\
IPsocery & Packet generator	 \\
Pacgen & Packet generator	 \\
Arp-sk & ARP packet generator	 \\
ARP-SPOOF & Packet intercept	 \\
Libpal & Packet intercept	 \\
Aicmspend & ICMP packet flooding	\\
\hline
\end{tabular}
\label{table:packetforging-tools}
\end{table}

\smallskip
\subsubsection{Application layer}
Application layer attacks use legitimate requests to use up all the resources of the victim. Unlike DDoS attacks, they are more subtle and more difficult to detect since they are using legitimate requests.

\smallskip
\section{Generating a realistic network load}
A problem that exists with generating a realistic network load is the unwanted burstiness of real networks. Misconfigurations can result in traffic that looks similar to flooding attacks \cite{b10} and are hard to reproduce in small sandbox networks. Existing datasets often do not include such noise which result in unrealistic datasets \cite{b10}.

\smallskip According to \cite{b10} it might be impossible to create a single representative background trace that contains benign and malicious traffic. Reason being that the expected traffic workload and characteristics are highly dependent on the network and its hosts.

Tools such as Trident provide multiple strategies to create benign traffic \cite{b10}.

\subsection{NIDS-based strategies}
Using a wellknown network intrusion detection system (NIDS) such as Snort they groom a packet trace from a local site. This approach causes problems since benign packets that are flagged as malicious (false postives) are removed.

\subsection{Synthetic generation strategies}
Software robots emulate user behaviour and thus generate synthetic traffic. The idea behind it is that the robot only creates connections with known good hosts.

\subsection{Trust-based strategies}
The third strategy they support is by grooming a packet trace taken at a local site using connection heuristics. Examples of these heuristics are failure rates or scanning characteristics. This approach makes use of different connection characteristics of malicious and benign sources based on a model of malicious connection behaviour \cite{b10}.
The advantage of using this strategy is that it is not influenced by a particular system (NIDS independence) and is based on transport level characteristics. The disadvantage is that a connection that is sufficiently similar to benign users will not be detected.

\section{Summary and conclusion}

\begin{thebibliography}{00}
\bibitem{b1} N. Hoque, M. H. Bhuyan, R. C. Baishya, D. K. Bhattacharyya und J. K. Kalita, ``Network attacks: Taxonomy, tools and systems'', Journal of Network and Computer Applications, 2014. Jg., Nr. 40, 2014.
\bibitem{b2} S. Hansman,  ``A Taxonomy of Network and Computer Attack Methodologies'', 2003.
\bibitem{b3} M. W. Pawar and J. Anuradha,  ``Network Security and Types of Attacks in Network'', Procedia Computer Science,  Nr. 48, 2015.
\bibitem{b4} A. Gagandeep and K. Pawan,  ``Analysis of Different Security Attacks in MANETs on Protocol Stack A-Review'', International Journal of Engineering and Advanced Technology (IJEAT), 2012.
\bibitem{b5} S. Hansman and R. Hunt,  ``A taxonomy of network and computer attacks'', Computers \& Security, Nr. 24, 2004.
\bibitem{b6} W. Stallings und L. Brown,  ``Computer Security: Principles and Practice'',  3rd Ed., 2015.
\bibitem{b7} Common Vulnerabilities and Exposures. [Online] Available at: https://cve.mitre.org/.
\bibitem{b8} J. D. Howard,  ``An Analysis Of Security Incidents On The Internet'', 1995.
\bibitem{b9} R. Spangler, ``Packet Sniffer Detection with AntiSniff'', 2003
\bibitem{b10}  J. Sommers, V. Yegneswaran and P. Barford. ``Toward Comprehensive Traffic Generation for Online IDS Evaluation''. Technical report, Dept. of Computer Science, University of Wisconsin, August 2005.
\end{thebibliography}

\end{document}

### 2020-02-26 paper1/paper1.tex
Extended version

@@
 Supervised learning in artificial intelligence requires a training dataset that is already labeled.
-Doing so with network packets is a challenge due to the fact that even Intrusion Detection System are not always correct with their guessing.
+Doing so with network packets is a challenge due to the fact that even Intrusion Detection Systems are not always correct with their guessing.
 To overcome this problem it is necessary to generate malicious and benign packets and label them accordingly.
-To increase the quality of the dataset the ip addresses from the machines should not remain the same for all of the data and the malicious and benign traffic should be mixed.
+To increase the quality of the dataset the IP addresses from the machines should not remain the same for all of the data and the malicious and benign traffic should be mixed.
 
@@
 \section{Introduction}
-Generating network traffic that can be used to train a neuronal network brings some challenges: there are many different attack types that need to be considered, a vast variety of attack tools to generate malicious traffic are available in the Internet each with their own advantages and disadvantages and generating realistic network traffic that contains that does not get mislabeled by intrusion detection systems just to name a few.
+Generating network traffic that can be used to train a neuronal network brings some challenges: there are many different attack types that need to be considered, a vast variety of attack tools to generate malicious traffic are available in the Internet each with their own advantages and disadvantages and generating realistic network traffic that does not get mislabeled by intrusion detection systems just to name a few.
 
@@
 The opposite is the case with active attacks, which attempt to manipulate network packets, for example \cite{b4}.
-Further approaches exist such as classifying into internal and external attacks, or classifying attacks based on the underlying OSI-layer.
+
+At this point it also appears necessary to address the distinction between computer and network attacks mentioned in \cite{b2}.
+Computer attacks are spoken of when a computer system is attacked in any way.
+For example, attempts are made to slow down the performance of PCs or to manipulate data on hard disks.
+Network attacks are divided into two variants: The first option is to use the network to attack a computer system.
+Although these are network attacks, they can also be seen as a subset of computer attacks.
+The second option is when network attacks are actually directed only at the network and not at computer systems, for example by congesting a network with packets.
+The following figure is included to provide clarity and shows the relationship between computers and network attacks.
+
+\begin{figure}[htbp]  
+\centerline{\includegraphics[scale=.90]{computer-network-attacks.jpg}}
+\caption{Ratio of computer and network attacks \cite{b2}}
+\end{figure}
+
+\medskip Further approaches exist such as classifying into internal and external attacks, or classifying attacks based on the underlying OSI-layer.
 
@@
 \centerline{\includegraphics[scale=.55]{network-tools.jpg}}
-\caption{taxonomy of network security tools \cite{b1}}
+\caption{Taxonomy of network security tools \cite{b1}}
 \end{figure}

### 2020-02-26 paper2/paper2.tex
Added a paragraph about using system call traces for host-based features (previous commit was missing .tex file)

--- 2020-02-25 paper2/paper2.tex
@@
 
+A different approach is the use of features derived from raw system call traces.
+The advantage of system call traces over log files is that they contain only raw information about the interaction between programs and the operating system kernel, whereas log files contain data interpreted by various programs, and potentially include large amounts of irrelevant data.
+There are multiple methods of analysing system calls.
+Some only look at the pattern of system calls, while others also look at the arguments passed to these system calls.
+Also, there is a distinction between syntactic features, which depend only on contiguous patterns of system calls, and semantic features, which additionally consider the meaning of system call arguments or groups of discontiguous system calls.
+The advantage of the semantic approach is that it is able to counter mimicry attacks, where the payload is modified by the attacker with the aim that the system calls it executes cannot be classified as being malicious \cite{creech2014}.
+
 \subsection*{Label definition}

### 2020-02-27 paper2/paper2.tex
Added conclusion

@@
 
-\subsection*{Label definition}
+\subsection{Label definition}
 
@@
 For data generated by simulating attacks, the labels can simply be based on the schedule of simulated attacks in combination with the IP addresses, ports and protocols involved \cite{Ids2018}.
+
+\section{Conclusion}
+
+The quality of a dataset is generally crucial for the performance of machine learning algorithms, and generating a high quality dataset for intrusion detection systems is not a simple task.
+While the generation of representative benign and malicious data in itself is fundamental, the goal of this paper was to give an overview about the equally important tasks of network simulation, recording and data extraction.
+For each of these tasks, a variety of different approaches has been examined.
+Because each approach has its own strengths and limitations, they have to be evaluated individually for each application, depending on its specific requirements.
+Also, the further development of the Internet and the ongoing introduction of new attack techniques has to be considered, which may affect the suitability of current methods in the future and therefore cause the necessity to alter them or to develop new ones.
 

### 2020-02-27 paper1/paper1.tex
Added summary and conclusion

@@
 \begin{IEEEkeywords}
-artificial intelligence, neuronal networks, computer networks, network security, packet generation
+artificial intelligence, neural networks, computer networks, network security, packet generation
 \end{IEEEkeywords}
@@
 \section{Introduction}
-Generating network traffic that can be used to train a neuronal network brings some challenges: there are many different attack types that need to be considered, a vast variety of attack tools to generate malicious traffic are available in the Internet each with their own advantages and disadvantages and generating realistic network traffic that does not get mislabeled by intrusion detection systems just to name a few.
+Generating network traffic that can be used to train a neural network brings some challenges: there are many different attack types that need to be considered, a vast variety of attack tools to generate malicious traffic are available in the Internet each with their own advantages and disadvantages and generating realistic network traffic that does not get mislabeled by intrusion detection systems just to name a few.
 
@@
 \begin{figure}[htbp]  
-\centerline{\includegraphics[scale=.90]{computer-network-attacks.jpg}}
+\centerline{\includegraphics{computer-network-attacks.jpg}}
 \caption{Ratio of computer and network attacks \cite{b2}}
@@
 This includes, for example, deliberately cutting network cables or destroying computers \cite{b2}.
-Password attacks: here an attempt is made to find out a specific password in order to gain control over a PC, for example.
+
+\smallskip Password attacks: here an attempt is made to find out a specific password in order to gain control over a PC, for example.
 This can be done in three different ways: trying out all possible password combinations (brute force), trying out a part of all possible password combinations (e.g.
@@
 \begin{figure}[htbp]  
-\centerline{\includegraphics[scale=.55]{network-tools.jpg}}
+\centerline{\includegraphics[scale=.65]{network-tools.jpg}}
 \caption{Taxonomy of network security tools \cite{b1}}
@@
 \section{Summary and conclusion}
+Various aspects must be taken into account for the generation of realistic network traffic for the production of a dataset, which is subsequently used for supervised machine learning.
+In the course of this paper, cyber-attacks were explained in principle, various tools for carrying out the attacks were shown, and how realistic network traffic based on benign and malicious data packets can be generated.
+The topics covered are summarized at this point in order to emphasize the most important content again.
+
+Cyber -attacks generally try to exploit various vulnerabilities and carry out malicious actions.
+In order to provide an overview of the multitude of different attack methods, there are various approaches in the literature to classify the attacks into a few groups.
+The focus of this work is the classification model found in  \cite{b5}, which aims to enable a holistic view of attacks.
+Basically, nine different attack classes are described, including network attacks, denial of service attacks, information gathering attacks and various types of malware-based attacks.
+Subsequently, various attack targets (software, hardware and network), as well as vulnerabilities (in design, implementation, configuration) are explained in more detail.
+Tools for carrying out-cyber attacks are basically divided into two different groups, namely information gathering tools and attack launching tools.
+The former tries to find out various information about the system to be attacked (open ports, operating system used, etc.).
+Sniffing and scanning tools can be used for this.
+Attack launching tools for performing denial of service attacks and for packet forging were identified.
+Furthermore, it is also possible to proceed on the basis of legitimate requests on the application layer.
+The creation of realistic network traffic based on benign and malicious data packets is a complex undertaking.
+By using tools, various strategies exist to generate harmless traffic, for example by using software robots.
+
+\section{Further research}
+Future work on this topic could try to look at attack methods on a more detailed level in order to provide specific information.
+Furthermore, it makes sense to identify additional tools for carrying out cyber-attacks.
+There is also room for further research in the search for other ways of generating realistic network traffic including benign and malicious behavior.
 

### 2020-02-27 paper1/paper1.tex
Implemented feedback from Boris

@@
 \section{Introduction}
-Generating network traffic that can be used to train a neural network brings some challenges: there are many different attack types that need to be considered, a vast variety of attack tools to generate malicious traffic are available in the Internet each with their own advantages and disadvantages and generating realistic network traffic that does not get mislabeled by intrusion detection systems just to name a few.
+Generating network traffic that can be used to train a neural network brings some challenges: there are many different attack types that need to be considered, a vast variety of attack tools to generate malicious traffic, each with their own advantages and disadvantages, and realistic benign network traffic has to be generated, just to name a few.
 
@@
 The second chapter focuses on tools to generate malicious traffic.
-The third chapter addresses the problem of generating useful network traffic.
+The third chapter addresses the problem of generating realistic network traffic.
 
@@
 Executing attacks generally consists of several steps: First, the attacker's motivation must be clarified (e.g.
-proof of your own abilities).
+proof of the attacker's own abilities).
 Then it is necessary to select a target and collect information about it.
@@
 \centerline{\includegraphics{computer-network-attacks.jpg}}
-\caption{Ratio of computer and network attacks \cite{b2}}
+\caption{Relationship of computer and network attacks \cite{b2}}
 \end{figure}
@@
 
-\medskip In the literature often cited is the classification model presented in \cite{b5}, which works on the basis of four dimensions and offers a holistic view of attacks.
+\medskip The classification model presented in \cite{b5}, which works on the basis of four dimensions and offers a holistic view of attacks, is often cited in the literature.
 First, an attack class is assigned based on the attack vector.
-The attacks target is specified in the second dimension.
+The attack's target is specified in the second dimension.
 The vulnerabilities that are exploited by the attack are then noted.
@@
 \subsection{Attack launching}
-The most packets however are generated by the attacks themselves.
-Especially some kinds of DoS attacks are designed to overload the network with packets such that the victim cannot respond to legitimate requests.
+The most packets, however, are generated by the attacks themselves.
+Especially some kinds of DoS attacks are designed to overload the network with packets so that the victim cannot respond to legitimate requests.
 
@@
 \smallskip According to \cite{b10} it might be impossible to create a single representative background trace that contains benign and malicious traffic.
-Reason being that the expected traffic workload and characteristics are highly dependent on the network and its hosts.
+The reason is that the expected traffic workload and characteristics are highly dependent on the network and its hosts.
 
@@
 \subsection{NIDS-based strategies}
-Using a wellknown network intrusion detection system (NIDS) such as Snort they groom a packet trace from a local site.
+Using a wellknown network intrusion detection system (NIDS) such as Snort, these tools can groom a packet trace from a local site.
 This approach causes problems since benign packets that are flagged as malicious (false postives) are removed.

### 2020-02-27 paper1/paper1.tex
Extended abstract and introduction

@@
 To increase the quality of the dataset the IP addresses from the machines should not remain the same for all of the data and the malicious and benign traffic should be mixed.
-
+Therefore, a literature search is carried out in the relevant specialist literature, scientific work and publications in order to research the necessary basics.
+Different types of attacks, tools for carrying out attacks and how realistic network traffic can be generated are to be determined.
+The results show that some approaches exist to divide the multitude of different attack methods into a few, individual groups.
+Nine generic attack classes as well as typical attack targets and vulnerabilities are identified.
+In addition, various tools for carrying out attacks are determined.
+These can be divided into the areas of information gathering and attack launching.
+Furthermore, different strategies are considered to generate realistic network traffic consisting of benign and malicious data packets.
+In conclusion, the information determined in the context of this work helps to generate an appropriate dataset for supervised machine learning.
 \end{abstract}
@@
 Therefore this paper sets a direction and foundation for the next steps and shows possible solutions at a high level.
+The results are elaborated on the basis of an extensive literature search in relevant specialist literature, scientific works and publications.
+The list of individual sources can be found at the end of the work.
 
@@
 \section{Further research}
-Future work on this topic could try to look at attack methods on a more detailed level in order to provide specific information.
+Future work on this topic could try to look at attack methods on a more detailed level in order to provide more specific information.
 Furthermore, it makes sense to identify additional tools for carrying out cyber-attacks.

### 2020-02-27 paper1/paper1.tex
Continued implementing feedback from Boris

@@
 \smallskip Worms: a software that spreads independently by exploiting vulnerabilities \cite{b6}.
-Examples include mass mailing worms, which spread via email, and network-aware worms, which only require four steps to infect a foreign host and spread themselves \cite{b2}.
+Examples include mass mailing worms, which spread via email, and network-aware worms, which require only a few steps to infect a foreign host and spread themselves \cite{b2}.
 
@@
 \medskip The classification model offers a comprehensive view of attacks.
-For example, for a worm that is distributed via email (dimension 1), the target Windows 95 (dimension 2) with an inadequate configuration (dimension 3) and a virus as payload (dimension 4) is determined.
+For example, for a worm that is distributed via email (dimension 1), the target Windows 95 (dimension 2) with an inadequate configuration (dimension 3) and a virus as payload (dimension 4) could be determined.
 Further dimensions could be specified, for example what damage is caused by an attack, or how many costs arise from successfully carrying out the attack.
@@
 There is a large number of network tools available in the world wide web.
-Many of them are free to download and can be used for malicious activities but there are also many that are for defensive purposes.
+Many of them are free to download and can be used for malicious activities but there are also many that can be used for defensive purposes.
 Generally speaking there are two major groups: tools for network attackers and tools for network defenders.\cite{b1} The tools for network attackers are the ones generating the packets thus being the ones we will focus on.
@@
 \smallskip Table \ref{table:sniffing-tools} contains a list of sniffing tools that are available online.
-Generally speaking sniffing tools do not generate packets and are therfore undetectable.
+Generally speaking, sniffing tools do not generate packets and are therfore undetectable.
 However, on a switched ethernet network segment a sniffer does generate a small amount of packets such as ARP\cite{b9}.

### 2020-02-27 paper2/paper2.tex
Update paper2.tex

--- 2020-02-25 paper2/paper2.tex
@@
 
-Hypervisors like VirtualBox or VMware provide the possibility to set up full virtual computers on a physical host where Docker containers are encapsuled sandboxes on an existing operating system like illustrated in Fig.~\ref{docker-vm}.
+Hypervisors like VirtualBox or VMware provide the possibility to set up full virtual computers on a physical host (see Fig.~\ref{docker-vm}a) while Docker containers are encapsuled sandboxes on an existing operating system (see Fig.~\ref{docker-vm}b).
 It is clear that containers are more resource efficent but in virtual machines the systems are better isolated.
@@
 If we focus on resource efficency and scalability also Kubernetes comes in our minds which is an orchestrator for deploying Docker containers on multiple nodes but it is a more complex configuration overhead and needs certain amount of hardware to unfold its potential also for simple scenarios.
-Because we want an affordable solutions which can be used in most of the cases also on our computers with some exceptions, we could use a tool which can create virtual machines as well as Docker containers to be ready for the future.
+Because we want an affordable solution which can be used in most of the cases also on our computers with some exceptions, we could use a tool which can create virtual machines as well as Docker containers to be ready for the future.
 Regarding the corner cases servers or a Kubernetes cluster could be rented temporarily on cloud platforms like AWS to support high-load scenarios e.g.
@@
 
-Fig.~\ref{docker-vm} shows a possible simulation workflow.
+Fig.~\ref{design-flow} shows a possible simulation workflow.
 The input parameters are the type of scenario(s) and the respective settings which deploys hosts including the base systems and the simulators applying the network topologies.
@@
 \section{Recording}
+
+Before the network traffic can be analysed, it has to be recorded first.
+Therefore, this chapter will give an overview of the available third-party software which can be used to record all the incoming network traffic.
+The functional requirements for a tool are chiefly the pace of handling the network traffic.
+All of these tools are client-based tools.
+Means that the software needs a client to be set up and used.
+This means that either the network target for the traffic has to be a client or the e.g.
+switch has to clone a port with all the incoming and outgoing traffic and direct all the data to a client running one or more of these tools.
+In some cases, analysing of network traffic can take much longer than the data is send or received by the network which means, that all the data need to be buffered or stored.
+In this case a packet-to-disk packet recorder is needed.
+This device can capture large amount of network traffic with a data rate of up to 40 Gbit/s and write it into files.
+
+\subsection{pcap Library}
+
+Every tools needs APIs\footnote{Application Programming Interface} to capture all the traffic sent through the network interface.
+Therefore, libraries were made for both Unix-like and Windows based operating systems.
+While \textit{libpcap} is used for Unix-like operating systems, a Windows system need the library called \textit{WinPcap}.
+With an pcap library software has the ability to tap all the network traffic from the selected interface.
+Based on the possibility to include the library in \textbf{C} and \textbf{C++}, \textit{pcap} can be used very easily into software.
+With the release of \textit{libpcap} 1.0.0, the zero-copy mechanism was implemented which allows the software to read the network traffic directly without copying the data from kernel space to user space.
+
+\subsection{Network Recording Tools}
+
+In this section we will itemise five well known and useful network recording tools.
+Some of the tools are console based tools, the other do have a GUI or both.
+All of these tools are client-based tools.
+Means that the software needs a Client to be set up and used.
+This means that either the network target for the traffic has to be a client or the e.g.
+switch has to clone a port with all the incoming and outgoing traffic and direct all the data to a client running one or more of these tools.
+
+\subsection*{tcpdump}
+
+Especially in the world of Unix based operating systems, \textit{tcpdump} is one of the best known tools for monitoring and sniffing network traffic.
+\textit{tcpdump} is exclusively command line based tool.
+This property makes \textit{tcpdump} a tool for advanced console line users only.
+But once you get familiar with the commands and options of \textit{tcpdump}, it will get a  a very efficient and powerful tool for many applications.
+\textit{tcpdump} uses the \textit{libpcap} network library of Unix based operating systems for capturing network traffic.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[scale=0.4]{tcpdump.png}}
+\caption{tcpdump Network Capture}
+\label{tcpdump}
+\end{figure}
+
+With the wide set of options, \textit{tcpdump} is a pretty agile tool for dumping any analysing network traffic. A few of the options are listed below \cite{wikiUbuntu}.
+
+\begin{itemize}
+\item \textbf{-i eth0} the traffic from the interface \textit{eth0} is captured
+\item \textbf{-w example.dump} all the output is written into a declared file
+\item \textbf{-e} the ethernet header will recorded too
+\item \textbf{-r example.dump} parsing the file \textit{example.dump} into tcpdump
+\end{itemize}
+
+For the Windows enthusiasts a Windows derivative of \textit{tcpdump} is available.
+It is called \textit{WinDump}.
+\textit{WinDump} uses so called WinPcap library for Windows operating systems.
+Which has the same APIs as libpcap for Unix-like systems.
+This means, network traffic captured, e.g.
+on a Ubuntu system can be analysed on a Windows 10 Client afterwards.
+
+\subsection*{Wireshark}
+
+\textit{Wireshark} is for good reason the best known cross-platform networking tool available.
+In 2006 \textit{Wireshark} was started as a fork of the software called \textit{Ethereal} by the Wireshark-Community due to trademark issues \cite{asrodiapatel2012}.
+The performance of this tool is powerful and the set of functions is neatly arranged.
+By using GTK toolkit, \textit{Wireshark} has implemented a GUI \footnote{Graphical User Interface}\cite{suribatra2012}.
+The GUI is separated typically into three segmentations, traffic overview, details of selected packet and the raw data of each packet.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[scale=0.48]{wireshark.png}}
+\caption{Wireshark Network Capture}
+\label{wireshark}
+\end{figure}
+
+A good asset of \textit{Wireshark} is its available set of filters for capturing and displaying network data.
+With the help of plugins, \textit{Wireshark} can even handle with new or dedicated protocols.
+Another big asset of \textit{Wireshark} is its capibility of capturing raw USB traffic from an USB-Port.
+By highlighting the traffic with different colours as shown in Fig.~\ref{wireshark} it's easier to scan for required traffic.
+
+\subsection*{Netsniff-ng}
+
+\textit{Netsniff} is a another Unix based tool for network analysis.
+\textit{Netsniff} was the first tool which had the ability to analyse the whole traffic with a zero-copy procedure, means that \textit{Netsniff} uses the same data as the kernel do.
+Based on this functional principle, \textit{Netsniff} is one of the best performing tools for network sniffing.
+\textit{Netsniff} is like \textit{tcpdump} exclusively an console based software.
+The set of options contains 38 capabilities on the latest version of \textit{Netsniff}: 0.6.4 (effective 26.02.2020).
+
+\subsection*{Capsa}
+
+The fourth tool is called \textit{Capsa} from \textit{Colasoft} founded 2001 in United States.
+\textit{Capsa} is suitable for real-time network data capturing both LAN and Wireless Interfaces.
+A big advantage of \textit{Capsa} is the ability to evaluate the network traffic into graphical output like diagrams.
+This feature can help identifying perpetrator of inadvertent network traffic quickly by a human.
+The free version of \textit{Capsa} is heavily restricted regarding available functions.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[scale=0.30]{capsa.png}}
+\caption{Graphical Evaluation of Network Traffic with Capsa}
+\label{capsa}
+\end{figure}
+
+\subsection*{EtherApe}
+
+Yet another tool for network monitoring is called \textit{EtherApe}.
+\textit{EtherApe} is an Unix based software with a meaningful graphic interface and graphical data output of network traffic.
+With the help graphical illustration anomalies like a DDoS attack can be detected easily as shown in Fig.~\ref{etherape}.
+Every host is shown as a node in circle, every connection is shown as link between two nodes and every protocol has an own colour for the link.
+With higher network traffic between two hosts the link between nodes is shown larger.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[scale=0.37]{etherape.png}}
+\caption{Graphic Representation of Network Traffic}
+\label{etherape}
+\end{figure}
 
@@
 108–126, 2017.
+\bibitem{wikiUbuntu} ubuntuusers.de
+\bibitem{asrodiapatel2012} P. Asrodia, H. Patel (2012). Analysis  of  Various  Packet  Sniffing  Tools  for Network  Monitoring  and  Analysis.
+\bibitem{suribatra2012} S. Suri, V. Batra, Comparative Study of Network Monitoring Tools.
 \end{thebibliography}

### 2020-02-27 paper1/paper1.tex
Implemented feedback from Boris III

@@
 
-Cyber -attacks generally try to exploit various vulnerabilities and carry out malicious actions.
+Cyber-attacks generally try to exploit various vulnerabilities and carry out malicious actions.
 In order to provide an overview of the multitude of different attack methods, there are various approaches in the literature to classify the attacks into a few groups.
@@
 Subsequently, various attack targets (software, hardware and network), as well as vulnerabilities (in design, implementation, configuration) are explained in more detail.
-Tools for carrying out-cyber attacks are basically divided into two different groups, namely information gathering tools and attack launching tools.
-The former tries to find out various information about the system to be attacked (open ports, operating system used, etc.).
+Tools for carrying out cyber-attacks are basically divided into two different groups, namely information gathering tools and attack launching tools.
+The former try to find out various information about the system to be attacked (open ports, operating system used, etc.).
 Sniffing and scanning tools can be used for this.

### 2020-02-27 paper1/paper1.tex
Added a table for better clarity

@@
 A possible variant of this is sniffing of packets, to find out about IP and MAC addresses or passwords \cite{b2}.
+
+\smallskip To provide better clarity, the individual attack classes, including some examples, are listed again in the following table:
+
+\begin{table}[htbp]
+\caption{Different attack classes and examples \cite{b2}} 
+\centering
+\begin{tabular}{ | c | c | }
+\hline
+Type of attack & Example(s) \\
+\hline
+Virus & File Infector, Macro Virus \\
+Worm & Mass-Mailing Worm, Network-Aware Worm \\
+Trojan & Logic Bomb \\
+Buffer overflow & Stack Buffer Overflow, Heap Overflow \\
+Denial of Service Attack & Host Based, Network Based, Distributed \\
+Network-Based Attack & Spoofing,  Web Application Attacks \\
+Physical Attack & Basic attacks \\
+Password Attack & Dictionary Attack, Brute Force \\
+Information Gathering &	Sniffing, Mapping, Security Scanning \\
+\hline
+\end{tabular}
+\label{table:attack-types}
+\end{table}
 

### 2020-02-28 paper2/paper2.tex
Fix errors

--- 2020-02-27 paper2/paper2.tex
@@
 
-Hypervisors like VirtualBox or VMware provide the possibility to setup full virtual computers on a physical host where Docker containers are encapsuled sandboxes on an existing operating system like illustrated in Fig.~\ref{docker-vm}.
+Hypervisors like VirtualBox or VMware provide the possibility to set up full virtual computers on a physical host (see Fig.~\ref{docker-vm}a) while Docker containers are encapsuled sandboxes on an existing operating system (see Fig.~\ref{docker-vm}b).
 It is clear that containers are more resource efficent but in virtual machines the systems are better isolated.
@@
 If we focus on resource efficency and scalability also Kubernetes comes in our minds which is an orchestrator for deploying Docker containers on multiple nodes but it is a more complex configuration overhead and needs certain amount of hardware to unfold its potential also for simple scenarios.
-Because we want an affordable solutions which can be used in most of the cases also on our computers with some exceptions, we decided to use a tool which can create virtual machines as well as Docker containers to be ready for the future.
-Regarding the corner cases servers or a Kubernetes cluster could be rented temporarily on cloud platforms to support high-load scenarios e.g.
+Because we want an affordable solution which can be used in most of the cases also on our computers with some exceptions, we could use a tool which can create virtual machines as well as Docker containers to be ready for the future.
+Regarding the corner cases servers or a Kubernetes cluster could be rented temporarily on cloud platforms like AWS to support high-load scenarios e.g.
 big denial-of-service attacks.
+\cite{b8}
 
@@
 
-Fig.~\ref{docker-vm} shows a schematic overview of the simulation workflow.
+Fig.~\ref{design-flow} shows a possible simulation workflow.
 The input parameters are the type of scenario(s) and the respective settings which deploys hosts including the base systems and the simulators applying the network topologies.
-Moreover the network recorder and data extractor will be deployed to gather the network traffic and extract the features to create data sets which could be collected in a central database.
+Moreover the network recorder and data extractor could be deployed to gather the network traffic and extract the features to create data sets which could be collected in a central database.
 
@@
 \begin{thebibliography}{00}
-\bibitem{b1} CBT Nuggets Blog (2020).
-5 Best Network Simulators for Cisco Exams: CCNA, CCNP, CCIE.
-[online] Available at: \newline\url{https://www.cbtnuggets.com/blog/career/career-progression/5-best-network-simulators-for-cisco-exams-ccna-ccnp-and-ccie} [Accessed 22 Feb.
-2020].
+\bibitem{b1} Seifert, C.
+and Reißmann, S.
+and Rieger, S.
+and Pape, C.
+Evaluation von VIRL, GNS3 und Mininet als Virtual Network Testbeds in der Hochschullehre.
+11.
+DFN-Forum Kommunikationstechnologien.
+Bonn: Gesellschaft für Informatik e.V.., pp.
+103-112.
 \bibitem{b2} Network Computing.

### 2020-02-28 paper2/paper2.tex
Update paper2.tex

--- 2020-02-27 paper2/paper2.tex
@@
 
-With the wide set of options, \textit{tcpdump} is a pretty agile tool for dumping any analysing network traffic. A few of the options are listed below \cite{wikiUbuntu}.
+With the wide set of options, \textit{tcpdump} is a pretty agile tool for dumping any analysing network traffic. A few of the options are listed below \cite{tcpdumporg}.
 
@@
 
+If \textit{tcpdump} is compiled with cryptography, it is able to decrypt IPsec ESP packets in real time \cite{goyal2017}.
 For the Windows enthusiasts a Windows derivative of \textit{tcpdump} is available.
@@
 \centerline{\includegraphics[scale=0.30]{capsa.png}}
-\caption{Graphical Evaluation of Network Traffic with Capsa}
+\caption{Graphical Evaluation of Network Traffic with Capsa \cite{firewallcx}}
 \label{capsa}
@@
 108–126, 2017.
-\bibitem{wikiUbuntu} ubuntuusers.de
-\bibitem{asrodiapatel2012} P. Asrodia, H. Patel (2012). Analysis  of  Various  Packet  Sniffing  Tools  for Network  Monitoring  and  Analysis.
-\bibitem{suribatra2012} S. Suri, V. Batra, Comparative Study of Network Monitoring Tools.
+\bibitem{tcpdumporg} Van Jacobson, C. Leres, S. McCanne,tcpdump.org (26.02.2020) https://www.tcpdump.org/manpages/tcpdump.1.html [Accessed 27.02.2020]
+\bibitem{asrodiapatel2012} P. Asrodia, H. Patel (2012). Analysis  of  Various  Packet  Sniffing  Tools  for Network  Monitoring  and  Analysis, ISSN  No.  (Online)  :  2277-2626
+\bibitem{suribatra2012} S.
+Suri, V.
+Batra, Comparative Study of Network Monitoring Tools, International Journal ofInnovative Technology and Exploring Engineering (IJITEE)ISSN: 2278-3075,Volume-1, Issue-3, August 2012
+\bibitem{goyal2017} P. Goyal, A. Goyal (2017), Comparative study of two most popular packet sniffing tools-Tcpdump and Wireshark
+\bibitem{firewallcx} http://www.firewall.cx/images/stories/review-capsa-enterprisev7-7.jpg [Accessed 26.02.2020]
 \end{thebibliography}

### 2020-02-28 paper2/paper2.tex
Replace and remove cite

@@
 All of these tools are client-based tools.
-Means that the software needs a client to be set up and used.
 This means that either the network target for the traffic has to be a client or the e.g.
 switch has to clone a port with all the incoming and outgoing traffic and direct all the data to a client running one or more of these tools.
-In some cases, analysing of network traffic can take much longer than the data is send or received by the network which means, that all the data need to be buffered or stored.
-In this case a packet-to-disk packet recorder is needed.
+In some cases, analysing of network traffic can take much longer than the data is sent or received by the network which means, that all the data need to be buffered or stored.
+In this case a packet-to-disk recorder is needed.
 This device can capture large amount of network traffic with a data rate of up to 40 Gbit/s and write it into files.
@@
 Therefore, libraries were made for both Unix-like and Windows based operating systems.
-While \textit{libpcap} is used for Unix-like operating systems, a Windows system need the library called \textit{WinPcap}.
-With an pcap library software has the ability to tap all the network traffic from the selected interface.
-Based on the possibility to include the library in \textbf{C} and \textbf{C++}, \textit{pcap} can be used very easily into software.
+While \textit{libpcap} is used for Unix-like operating systems, a Windows system needs the library called \textit{WinPcap}.
+In 2013 the last release (4.1.3) of \textit{WinPcap} was published.
+Since then \textit{Npcap} acts as a substitute\cite{winpcaporg}.
+With a pcap library, software has the ability to tap all the network traffic from the selected interface.
+Based on the possibility to include the library in \textbf{C} and \textbf{C++}, \textit{pcap} can be used very easily in software.
 With the release of \textit{libpcap} 1.0.0, the zero-copy mechanism was implemented which allows the software to read the network traffic directly without copying the data from kernel space to user space.
@@
 
-In this section we will itemise five well known and useful network recording tools.
-Some of the tools are console based tools, the other do have a GUI or both.
-All of these tools are client-based tools.
-Means that the software needs a Client to be set up and used.
-This means that either the network target for the traffic has to be a client or the e.g.
-switch has to clone a port with all the incoming and outgoing traffic and direct all the data to a client running one or more of these tools.
+In this section we will itemise five well known and useful network recording tools. Some of the tools are console based tools, others have a GUI, or both.
 
@@
 Especially in the world of Unix based operating systems, \textit{tcpdump} is one of the best known tools for monitoring and sniffing network traffic.
-\textit{tcpdump} is exclusively command line based tool.
-This property makes \textit{tcpdump} a tool for advanced console line users only.
+\textit{tcpdump} is an exclusively command line based tool.
+This property makes \textit{tcpdump} a tool for advanced command line users only.
 But once you get familiar with the commands and options of \textit{tcpdump}, it will get a  a very efficient and powerful tool for many applications.
@@
 
-With the wide set of options, \textit{tcpdump} is a pretty agile tool for dumping any analysing network traffic. A few of the options are listed below \cite{tcpdumporg}.
+With the wide set of options, \textit{tcpdump} is a pretty agile tool for dumping and analysing network traffic. A few of the options are listed below \cite{tcpdumporg}.
 
@@
 It is called \textit{WinDump}.
-\textit{WinDump} uses so called WinPcap library for Windows operating systems.
-Which has the same APIs as libpcap for Unix-like systems.
+\textit{WinDump} uses the \textit{WinPcap} library, which provides the same API for Windows operating systems as \textit{libpcap} for Unix-like systems.
 This means, network traffic captured, e.g.
@@
 The performance of this tool is powerful and the set of functions is neatly arranged.
-By using GTK toolkit, \textit{Wireshark} has implemented a GUI \footnote{Graphical User Interface}\cite{suribatra2012}.
-The GUI is separated typically into three segmentations, traffic overview, details of selected packet and the raw data of each packet.
+By using the GTK toolkit, \textit{Wireshark} has implemented a GUI \footnote{Graphical User Interface}\cite{suribatra2012}.
+The GUI is separated typically into three segments: the traffic overview, and the details and raw data of the selected packet.
 
@@
 With the help of plugins, \textit{Wireshark} can even handle with new or dedicated protocols.
-Another big asset of \textit{Wireshark} is its capibility of capturing raw USB traffic from an USB-Port.
+Another big asset of \textit{Wireshark} is its capability of capturing raw USB traffic from an USB port.
 By highlighting the traffic with different colours as shown in Fig.~\ref{wireshark} it's easier to scan for required traffic.
 
-\subsection*{Netsniff-ng}
+\subsection*{netsniff-ng}
 
-\textit{Netsniff} is a another Unix based tool for network analysis.
-\textit{Netsniff} was the first tool which had the ability to analyse the whole traffic with a zero-copy procedure, means that \textit{Netsniff} uses the same data as the kernel do.
-Based on this functional principle, \textit{Netsniff} is one of the best performing tools for network sniffing.
-\textit{Netsniff} is like \textit{tcpdump} exclusively an console based software.
-The set of options contains 38 capabilities on the latest version of \textit{Netsniff}: 0.6.4 (effective 26.02.2020).
+\textit{netsniff-ng} is a another Unix based tool for network analysis.
+\textit{netsniff-ng} was the first tool which had the ability to analyse the whole traffic with a zero-copy procedure, which means that \textit{netsniff-ng} uses the same data as the kernel does.
+Based on this functional principle, \textit{netsniff-ng} is one of the best performing tools for network sniffing.
+\textit{netsniff-ng} is like \textit{tcpdump} exclusively an console based software.
+The set of options contains 38 capabilities on the latest version 0.6.4 of \textit{netsniff-ng}.
 
@@
 A big advantage of \textit{Capsa} is the ability to evaluate the network traffic into graphical output like diagrams.
-This feature can help identifying perpetrator of inadvertent network traffic quickly by a human.
+This feature can help identifying the source of undesirable network traffic quickly by a human.
 The free version of \textit{Capsa} is heavily restricted regarding available functions.
@@
 \textit{EtherApe} is an Unix based software with a meaningful graphic interface and graphical data output of network traffic.
-With the help graphical illustration anomalies like a DDoS attack can be detected easily as shown in Fig.~\ref{etherape}.
+With the help of graphical illustration anomalies like a DDoS attack can be detected easily as shown in Fig.~\ref{etherape}.
 Every host is shown as a node in circle, every connection is shown as link between two nodes and every protocol has an own colour for the link.
@@
 \centerline{\includegraphics[scale=0.37]{etherape.png}}
-\caption{Graphic Representation of Network Traffic}
+\caption{Graphic Representation of Network Traffic \cite{etherapewiki}}
 \label{etherape}
@@
 A common approach is to group packets into flows, and then calculate statistical features based on these flows.
-Generally, a flow is a group of packets with the same source and destination IP addresses and ports, and the same protocol \cite{brownlee1998}.
+Generally, a flow is a group of packets with the same source and destination IP addresses and ports, and the same protocol.
 However, a flow can also be terminated by a TCP FIN packet or a timeout \cite{Ids2018}.
@@
 \bibitem{b8} Bernstein, D. Containers and Cloud: From LXC to Docker to Kubernetes, in IEEE Cloud Computing, vol. 1, no. 3, pp. 81-84, Sept. 2014.
-\bibitem{b9} Nishant Kumar Singh, S.
-Thakur, H.
-Chaurasiya and H.
-Nagdev, Automated provisioning of application in IAAS cloud using Ansible configuration management, 1st International Conference on Next Generation Computing Technologies (NGCT), Dehradun, 2015, pp.
-81-85.
+\bibitem{b9} Sandobalín, J.
+and Insfran, E.
+and Abrahão, S.
+(2020).
+On the Effectiveness of Tools to Support Infrastructure as Code: Model-Driven versus Code-Centric.
+IEEE Access.
+10.1109/ACCESS.2020.2966597.
 \bibitem{b10} Spanaki, P. Cloud Computing: Security Issues and Establishing Virtual Cloud Environment via Vagrant to Secure Cloud Hosts, 2018, pp. 539-553
 \bibitem{b11} Ivanova, D. and Borovska, P. and Zahov, S. Development of PaaS using AWS and Terraform for medical imaging analytics, AIP Conference Proceedings, 2018
+\bibitem{winpcaporg} WinPcap: News, Riverbed Technology (2010), https://www.winpcap.org/misc/copyright.htm [Accessed 26.02.2020]
+\bibitem{tcpdumporg} Van Jacobson, C. Leres, S. McCanne,tcpdump.org (26.02.2020) https://www.tcpdump.org/manpages/tcpdump.1.html [Accessed 27.02.2020]
+\bibitem{goyal2017} P. Goyal, A. Goyal (2017), Comparative study of two most popular packet sniffing tools-Tcpdump and Wireshark
+\bibitem{asrodiapatel2012} P. Asrodia, H. Patel (2012). Analysis  of  Various  Packet  Sniffing  Tools  for Network  Monitoring  and  Analysis, ISSN  No.  (Online)  :  2277-2626
+\bibitem{suribatra2012} S.
+Suri, V.
+Batra, Comparative Study of Network Monitoring Tools, International Journal ofInnovative Technology and Exploring Engineering (IJITEE)ISSN: 2278-3075,Volume-1, Issue-3, August 2012
+\bibitem{firewallcx} http://www.firewall.cx/images/stories/review-capsa-enterprisev7-7.jpg [Accessed 26.02.2020]
+\bibitem{etherapewiki} https://upload.wikimedia.org/wikipedia/commons/a/ae/Etherape-gui.png [Accessed 25.02.2020]
 \bibitem{Nb2015} Moustafa, N.
@@
 460-473, 2013
-\bibitem{brownlee1998} Brownlee, N. (1998). Network management and realtime traffic flowmeasurement. J. Netw. Syst. Manage., vol. 6, pp. 223–228, June 1998.
 \bibitem{waizumi2007} Waizumi, Y.
@@
 108–126, 2017.
-\bibitem{tcpdumporg} Van Jacobson, C. Leres, S. McCanne,tcpdump.org (26.02.2020) https://www.tcpdump.org/manpages/tcpdump.1.html [Accessed 27.02.2020]
-\bibitem{asrodiapatel2012} P. Asrodia, H. Patel (2012). Analysis  of  Various  Packet  Sniffing  Tools  for Network  Monitoring  and  Analysis, ISSN  No.  (Online)  :  2277-2626
-\bibitem{suribatra2012} S.
-Suri, V.
-Batra, Comparative Study of Network Monitoring Tools, International Journal ofInnovative Technology and Exploring Engineering (IJITEE)ISSN: 2278-3075,Volume-1, Issue-3, August 2012
-\bibitem{goyal2017} P. Goyal, A. Goyal (2017), Comparative study of two most popular packet sniffing tools-Tcpdump and Wireshark
-\bibitem{firewallcx} http://www.firewall.cx/images/stories/review-capsa-enterprisev7-7.jpg [Accessed 26.02.2020]
 \end{thebibliography}

### 2020-02-28 paper2/paper2.tex

@@
 A common approach is to group packets into flows, and then calculate statistical features based on these flows.
-Generally, a flow is a group of packets with the same source and destination IP addresses and ports, and the same protocol.
+Generally, a flow is a group of packets with the same source and destination IP addresses and ports, and the same protocol \cite{brownlee1998}.
 However, a flow can also be terminated by a TCP FIN packet or a timeout \cite{Ids2018}.
@@
 \bibitem{b8} Bernstein, D. Containers and Cloud: From LXC to Docker to Kubernetes, in IEEE Cloud Computing, vol. 1, no. 3, pp. 81-84, Sept. 2014.
-\bibitem{b9} Sandobalín, J.
-and Insfran, E.
-and Abrahão, S.
-(2020).
-On the Effectiveness of Tools to Support Infrastructure as Code: Model-Driven versus Code-Centric.
-IEEE Access.
-10.1109/ACCESS.2020.2966597.
+\bibitem{b9} Nishant Kumar Singh, S.
+Thakur, H.
+Chaurasiya and H.
+Nagdev, Automated provisioning of application in IAAS cloud using Ansible configuration management, 1st International Conference on Next Generation Computing Technologies (NGCT), Dehradun, 2015, pp.
+81-85.
 \bibitem{b10} Spanaki, P. Cloud Computing: Security Issues and Establishing Virtual Cloud Environment via Vagrant to Secure Cloud Hosts, 2018, pp. 539-553
@@
 460-473, 2013
+\bibitem{brownlee1998} Brownlee, N. (1998). Network management and realtime traffic flowmeasurement. J. Netw. Syst. Manage., vol. 6, pp. 223–228, June 1998.
 \bibitem{waizumi2007} Waizumi, Y.

### 2020-02-28 paper2/paper2.tex

--- 2020-02-28 paper2/paper2.tex
@@
 \begin{abstract}
-Collecting real network data for further processing is an important topic for the training of neuronal networks which filter suspicious from ordinary traffic.
+Collecting real network data for further processing is an important topic for the training of neural networks which filter suspicious from ordinary traffic.
 This paper focuses mainly on research in the areas of network simulation, network recording and data extraction which should serve as input for the following papers.
@@
 \begin{IEEEkeywords}
-artifical intelligence, neuronal networks, computer networks, network security, data collection
+artifical intelligence, neural networks, computer networks, network security, data collection
 \end{IEEEkeywords}
@@
 \section{Introduction}
-Neuronal networks which filter suspicious network traffic need a training data set of high quality.
+Neural networks which filter suspicious network traffic need a training data set of high quality.
 This data set could be produced from real network traffic but this can be difficult to implement.
@@
 This means we would be ready to create system environments on our host or in the cloud and install base systems on them.
-The next section focuses on the automated setup of the network simulator, recorder and data extractor within virtual machines or Docker containers as well as the configuration of network topologies based on the selected scenarios and parameters to produce, record and extract network traffic for further processing in neuronal networks to improve their strike rate.
+The next section focuses on the automated setup of the network simulator, recorder and data extractor within virtual machines or Docker containers as well as the configuration of network topologies based on the selected scenarios and parameters to produce, record and extract network traffic for further processing in neural networks to improve their strike rate.
 

### 2020-02-28 paper1/paper1.tex
Added a missing reference

@@
 
-\title{Network packet generation for Artificial Intelligence\\
-}
+\title{Network Packet Generation for Artificial Intelligence}
 
@@
 
-\smallskip To provide better clarity, the individual attack classes, including some examples, are listed again in the following table:
+\smallskip To provide better clarity the individual attack classes, including some examples, are listed in Table \ref{table:attack-types}.:
 

### 2020-04-03 research/paper2/paper2.tex
First gns3 server setup

--- 2020-02-28 paper2/paper2.tex
@@
 
-\title{Network Data Collection for Artifical Intelligence
-}
+\title{Network Data Collection for Artificial Intelligence}
 
@@
 \begin{IEEEkeywords}
-artifical intelligence, neural networks, computer networks, network security, data collection
+artificial intelligence, neural networks, computer networks, network security, data collection
 \end{IEEEkeywords}
@@
 In the third chapter methods of feature extraction and label definition are covered.
-This information serves as foundation for the further implementation of a system for automated data collection for artifical intelligence.
+This information serves as foundation for the further implementation of a system for automated data collection for artificial intelligence.
 

### 2020-04-03 simulation/connect.sh

@@ -1,0 +1,4 @@
+#!/bin/bash
+
+ssh -L 1194:localhost:1194 mcs20-paai@mcs20-paai-n1.cs.technikum-wien.at -f sleep infinity
+sudo openvpn client.ovpn

### 2020-06-27 paper1/paper1.tex
Add paper1 template

--- 2020-02-27 paper1/paper1.tex
@@
 
-\title{Network packet generation for Artificial Intelligence\\
-}
+\title{Automated Network Packet Generation for Evaluation of Neural Networks in Intrusion Prevention Systems}
 
-\author{\IEEEauthorblockN{ Bernhard Gally}
+\author{\IEEEauthorblockN{Bernhard Gally}
 \IEEEauthorblockA{\textit{IT Security} \\
@@
 \and
-\IEEEauthorblockN{Maximilian Wech}
+\IEEEauthorblockN{Sebastian Lipp}
 \IEEEauthorblockA{\textit{IT Security} \\
 \textit{FH Technikum Wien}\\
-Vienna, Austria\\
-cs19m020@technikum-wien.at}
+Vienna, Austria \\
+cs19m032@technikum-wien.at}
+\and
+\IEEEauthorblockN{Damir Marijanovic}
+\IEEEauthorblockA{\textit{IT Security} \\
+\textit{FH Technikum Wien}\\
+Vienna, Austria \\
+cs19m031@technikum-wien.at}
 }
@@
 \begin{abstract}
-Supervised learning in artificial intelligence requires a training dataset that is already labeled.
-Doing so with network packets is a challenge due to the fact that even Intrusion Detection Systems are not always correct with their guessing.
-To overcome this problem it is necessary to generate malicious and benign packets and label them accordingly.
-To increase the quality of the dataset the IP addresses from the machines should not remain the same for all of the data and the malicious and benign traffic should be mixed.
-Therefore, a literature search is carried out in the relevant specialist literature, scientific work and publications in order to research the necessary basics.
-Different types of attacks, tools for carrying out attacks and how realistic network traffic can be generated are to be determined.
-The results show that some approaches exist to divide the multitude of different attack methods into a few, individual groups.
-Nine generic attack classes as well as typical attack targets and vulnerabilities are identified.
-In addition, various tools for carrying out attacks are determined.
-These can be divided into the areas of information gathering and attack launching.
-Furthermore, different strategies are considered to generate realistic network traffic consisting of benign and malicious data packets.
-In conclusion, the information determined in the context of this work helps to generate an appropriate dataset for supervised machine learning.
+@Damir
 \end{abstract}
@@
 \section{Introduction}
-Generating network traffic that can be used to train a neural network brings some challenges: there are many different attack types that need to be considered, a vast variety of attack tools to generate malicious traffic, each with their own advantages and disadvantages, and realistic benign network traffic has to be generated, just to name a few.
+@Damir
 
-This paper is split into three chapters.
-The first chapter explains the different kinds of attacks and their targets.
-The second chapter focuses on tools to generate malicious traffic.
-The third chapter addresses the problem of generating realistic network traffic.
+\section{Infrastructure}
+@Sebastian
 
-At this point all of our considerations are based on theoretical research without practical implementation which will come to life in further work.
-Therefore this paper sets a direction and foundation for the next steps and shows possible solutions at a high level.
-The results are elaborated on the basis of an extensive literature search in relevant specialist literature, scientific works and publications.
-The list of individual sources can be found at the end of the work.
+\section{Simulation}
+@Bernhard
 
-\section{Cyber-attacks: types, targets, vulnerabilites}
-\label{sec:cattv}
-
-This chapter intends to provide an overview of various types of cyber-attacks, cyber-attack targets and exploitable vulnerabilities.
-
-\subsection{General information}
-In order to fulfil the task, it initially appears necessary to describe typical characteristics of attacks.
-Attacks aim to exploit weaknesses, such as poorly configured systems, or errors in software and to circumvent security measures.
-Attackers deliberately try to make services unavailable, spy on or manipulate data and disrupt normal traffic as much as possible \cite{b1}.
-Executing attacks generally consists of several steps: First, the attacker's motivation must be clarified (e.g.
-proof of the attacker's own abilities).
-Then it is necessary to select a target and collect information about it.
-As soon as this has happened, the attack method is selected and the attack is performed \cite{b2}.
-
-\subsection{Classifying attacks}
-
-\medskip Basically, there are too many different cyber-attack methods to mention all of them.
-Therefore, various approaches exist in the literature to classify attacks into different groups.
-One way is to split attacks into active and passive ones \cite{b3,b4}.
-In the case of passive attacks, attempts are not made to manipulate data in network traffic, but rather to eavesdrop and collect sensitive information.
-The opposite is the case with active attacks, which attempt to manipulate network packets, for example \cite{b4}.
-
-At this point it also appears necessary to address the distinction between computer and network attacks mentioned in \cite{b2}.
-Computer attacks are spoken of when a computer system is attacked in any way.
-For example, attempts are made to slow down the performance of PCs or to manipulate data on hard disks.
-Network attacks are divided into two variants: The first option is to use the network to attack a computer system.
-Although these are network attacks, they can also be seen as a subset of computer attacks.
-The second option is when network attacks are actually directed only at the network and not at computer systems, for example by congesting a network with packets.
-The following figure is included to provide clarity and shows the relationship between computers and network attacks.
-
-\begin{figure}[htbp]  
-\centerline{\includegraphics{computer-network-attacks.jpg}}
-\caption{Relationship of computer and network attacks \cite{b2}}
-\end{figure}
-
-\medskip Further approaches exist such as classifying into internal and external attacks, or classifying attacks based on the underlying OSI-layer.
-
-\medskip The classification model presented in \cite{b5}, which works on the basis of four dimensions and offers a holistic view of attacks, is often cited in the literature.
-First, an attack class is assigned based on the attack vector.
-The attack's target is specified in the second dimension.
-The vulnerabilities that are exploited by the attack are then noted.
-Finally, the fourth dimension analyses whether an attack has a payload or an effect that exceeds the typical characteristics of the type of attack \cite{b5} .
-The rest of this chapter will focus on further introducing this model and consider each dimension individually.
-
-\subsubsection{First dimension}
-
-\medskip As already mentioned, in the first dimension the attack class is determined on the basis of the attack vector. The following generic classifications are listed in \cite{b5}:
-
-\smallskip Virus: this is a malware that replicates itself and, for example, infects other script code.
-When the infected source code is executed, the virus is also executed \cite{b6}.
-Basically there are many different types of viruses.
-An example would be File Infectors, in which a virus is embedded in a file.
-If this file is executed, the virus is also executed \cite{b2}.
-
-\smallskip Worms: a software that spreads independently by exploiting vulnerabilities \cite{b6}.
-Examples include mass mailing worms, which spread via email, and network-aware worms, which require only a few steps to infect a foreign host and spread themselves \cite{b2}.
-
-\smallskip Trojans: a software that appears harmless at first, but also contains a hidden, malicious function \cite{b6}.
-One example are logic bombs, which only execute their malicious payload when a specific condition (e.g.
-a certain time of a day) has been reached \cite{b2}.
-
-\smallskip Buffer overflows: This type of attack is based on programming errors and is very often used to attack computers or networks.
-Principally an attempt is made to overfill a temporary memory area.
-The amount of data that exceeds the size of the memory overwrites and corrupts other data.
-One possible goal of this attack would be to take control of a process.
-The main types of this attack are stack and heap buffer overflows \cite{b2}.
-
-\smallskip Denial of service attacks: this means attacks on the availability of systems.
-An attempt is made to exhaust the resources required to run a service \cite{b6}.
-This attack can be host-based, network-based, or distributed.
-The latter version is a DDoS attack, in which several attackers attack a mutual target at the same time \cite{b2}.
-
-\smallskip Network attacks: the category is divided into several sub-areas.
-If an attacker pretends to be someone else by manipulating certain addresses, this is called spoofing.
-This can take place on the basis of the MAC address (MAC Address Spoofing) or on the basis of the IP address (IP Spoofing).
-An attack in which IP spoofing is used is session hijacking.
-Here, an attacker tries to link into a session (based, for example, on TCP / Telnet) between two hosts and virtually replace one host.
-In the case of wireless network attacks, weaknesses such as too short initialization vectors of the WEP standard can be exploited.
-In the event of attacks on web applications, attempts are made to exploit security gaps in them.
-For example, if there is no strict separation between user input and program instructions, or there is insufficient input validation, an attacker can successfully send manipulated SQL statements to the underlying database.
-The consequences would be spying out of data or losing control of the underlying server.
-\cite{b2}.
-
-\smallskip Physical attacks: this type of attack aims to cause physical damage to target systems (i.e.
-computers or networks).
-This includes, for example, deliberately cutting network cables or destroying computers \cite{b2}.
-
-\smallskip Password attacks: here an attempt is made to find out a specific password in order to gain control over a PC, for example.
-This can be done in three different ways: trying out all possible password combinations (brute force), trying out a part of all possible password combinations (e.g.
-testing simple passwords) and determining deficiencies in password storage / encryption \cite{b2,b5}.
-
-\smallskip Information gathering attacks: this type of attack does not change data; it is only about spying on (sensitive) information, which can be used for follow-up attacks \cite{b5}.
-A possible variant of this is sniffing of packets, to find out about IP and MAC addresses or passwords \cite{b2}.
-
-\medskip \subsubsection{Second dimension}The targets of attacks are described in the second dimension of the classification model.
-These should be described as specifically as possible, which means that it is not sufficient, for example, to specify which server was attacked, but rather which service or which operating system is affected with which version.
-It should also be noted that an attack can target more than one target.
-Basically three groups of attack targets are defined \cite{b5}:
-
-\smallskip Hardware: this group is divided into three areas.
-First of all, computers and their components (e.g.
-processors, random access memory) should be mentioned.
-Network devices such as routers or switches also belong to this category.
-Ultimately, peripheral devices such as monitors and keyboards are also included.
-
-\smallskip Software: here a distinction is made between targets on operating systems (Windows, Linux, etc.) and various applications (email client, text processing program) that are executed on an operating system.
-
-\smallskip Network: in this group, the target is not software or hardware but the network itself, or an associated network protocol (for example, TCP).
-
-\medskip \subsubsection{Third dimension}In the third dimension of the classification model \cite{b5}, vulnerabilities that can be exploited by attacks are considered.
-Reference is made to the industry standard CVE - Common Vulnerabilities and Exposures \cite{b7}, which is used for the uniform naming of security gaps and vulnerabilities in computer systems.
-General vulnerability types are also defined in\cite{b8}:
-
-\smallskip Implementation: if a system is not implemented according to the design, security gaps arise.
-
-\smallskip Design: an incorrect design leads to a security-threatening implementation.
-
-\smallskip Configuration: If a system is not configured correctly, doors are opened to the attackers.
-
-\medskip \subsubsection{Fourth dimension}The last dimension analyzes whether an attack has a payload or an effect that exceeds the typical characteristics of the type of attack.
-This should result in a detailed view of the payload.
-For example, a worm's payload might be a Trojan, or the worm might just be manipulating files \cite{b5}.
-
-\medskip The classification model offers a comprehensive view of attacks.
-For example, for a worm that is distributed via email (dimension 1), the target Windows 95 (dimension 2) with an inadequate configuration (dimension 3) and a virus as payload (dimension 4) could be determined.
-Further dimensions could be specified, for example what damage is caused by an attack, or how many costs arise from successfully carrying out the attack.
-Overall an overview is provided of how attacks can be classified, which attack targets are possible, or which vulnerabilities exist.
-
-\section{Tools}
-
-There is a large number of network tools available in the world wide web.
-Many of them are free to download and can be used for malicious activities but there are also many that can be used for defensive purposes.
-Generally speaking there are two major groups: tools for network attackers and tools for network defenders.\cite{b1} The tools for network attackers are the ones generating the packets thus being the ones we will focus on.
-
-\begin{figure}[htbp]  
-\centerline{\includegraphics[scale=.65]{network-tools.jpg}}
-\caption{Taxonomy of network security tools \cite{b1}}
-\end{figure}
-
-\subsection{Information gathering}
-Attackers usually collect information before starting their attack.
-Understanding the environment, what operating systems are used, finding open ports and available services are only a small part of information that can be obtained.
-
-\smallskip Table \ref{table:sniffing-tools} contains a list of sniffing tools that are available online.
-Generally speaking, sniffing tools do not generate packets and are therfore undetectable.
-However, on a switched ethernet network segment a sniffer does generate a small amount of packets such as ARP\cite{b9}.
-Other types of sniffers, such as Ettercap, can execute a Man-in-the-middle attack.
-This causes a lot of packets to be sent through the network as it needs to answer requests in order to spoof an identity.
-
-\begin{table}[htbp]
-\caption{Sniffing tools \cite{b1}} 
-\centering
-\begin{tabular}{ | c | c | }
-\hline
-Tool name & Purpose \\
-\hline
-Ethereal & Packet capturing \\
-Tcpdump & Packet capturing  \\
-Net2pcap & Packet capturing  \\
-Snoop & Packet capturing \\
-Angst & Sniffing  \\
-Ngrep & Capturing packet  \\
-Ettercap & Man-in-the-middle attack  \\
-Dsniff & Password sniffing  \\
-Cain \& able & Password recovery  \\
-Aimsniff & Capturing packet \\
-Tcptrace & Analysis of traffic \\
-Tcptrack & TCP connection analysis  \\
-Argus & Analysis of audit data \\
-Karpski & Packet analyzer \\
-IPgrab & Display packet header \\
-Nast & Traffic analysis \\
-Gulp & Packet capturing/visualization \\
-Libpcap & Packet capturing \\
-Nfsen & Flow capturing/ visualization \\
-Nfdump & Flow capturing/visualization \\
-\hline
-\end{tabular}
-\label{table:sniffing-tools}
-\end{table}
-
-Scanning tools are often more intrusive than sniffing tools and generate lots of packets that can be detected.
-Their aim is to show up open ports, available services, used operating systems.
-With the gathered information an attacker can pick a suitable attack vector.
-Table \ref{table:scanning-tools} shows a few common network scanning tools.
-
-\begin{table}[htbp]
-\caption{Scanning tools \cite{b1}} 
-\centering
-\begin{tabular}{ | c | c | }
-\hline
-Tool name & Purpose \\
-\hline
-Nmap & Scanning	 \\
-Amap & Scanning \\
-Vmap & Version mapping \\
-Unicornscan & Scanning \\
-Ttlscan & Scanning \\
-Ike-scan & Host discovery \\
-Paketto & Scanning \\
-\hline
-\end{tabular}
-\label{table:scanning-tools}
-\end{table}
-
-\subsection{Attack launching}
-The most packets, however, are generated by the attacks themselves.
-Especially some kinds of DoS attacks are designed to overload the network with packets so that the victim cannot respond to legitimate requests.
-
-\subsubsection{Denial of service}
-The Internet provides many tools that are capable of executing DoS attacks and can be easily used to crash hosts or even networks.
-Two examples of such tools are Low Orbit Ion Cannon (LOIC) or High Orbit Ion Cannon (HOIC). Both of these tools are capable of launching DoS/DDoS attacks within a short amount of time.
-LOIC supports TCP, UDP and HTTP whereas HOIC only supports the HTTP protocol\cite{b1}. A list of DoS and DDoS tools is shown in Table \ref{table:ddos-tools}.
-
-\begin{table}[htbp]
-\caption{DoS/DDoS tools \cite{b1}} 
-\centering
-\begin{tabular}{ | c | c | }
-\hline
-Tool name & Purpose \\
-\hline
-Jolt & DoS \\
-Burbonic & DoS \\
-Targa & DoS \\
-Blas20 & DoS \\
-Crazy Pinger & DoS \\
-UDPFlood & DoS \\
-FSMax & DoS \\
-Nemsey & DoS \\
-Panther & DoS \\
-Land \& LaTierra & DoS \\
-Slowloris & DoS \\
-Blackenergy & DDoS \\
-HOIC & DDoS \\
-Shaft & DDoS \\
-Knight & DDoS \\
-Kaiten & DDoS \\
-RefRef & DDoS \\
-Hgod & DDoS \\
-LOIC & DDoS \\
-Trinoo & DDoS \\
-TFN & DDoS \\
-TFN2K & DDoS \\
-Stachaldraht & DDoS \\
-Mstream & DDoS \\
-Trinity & DDoS \\
-\hline
-\end{tabular}
-\label{table:ddos-tools}
-\end{table}
-
-\smallskip
-\subsubsection{Packet forging}
-Packet forging tools are necessary to manipulate packet information.
-With the help of such tools it is possible to generate traffic with non-existant IP addresses.
-They also allow to individually set the flags of network packets to a desired state.
-An extreme example of such a packet is the "Christmas tree packet", which has all available flags set to 1.
-A list of packet forging tools is shown in Table \ref{table:packetforging-tools}.
-
-\begin{table}[htbp]
-\caption{Packet forging tools \cite{b1}} 
-\centering
-\begin{tabular}{ | c | c | c |}
-\hline
-Tool name & Purpose \\
-\hline
-Packeth & Packet generator	 \\
-Packit & Packet analysis and injection	 \\
-Nemesis & Packet crafting/injection	 \\
-Tcpinject & Packet generator	 \\
-Libnet & Packet injection	 \\
-SendIP & Packet generator	 \\
-IPsocery & Packet generator	 \\
-Pacgen & Packet generator	 \\
-Arp-sk & ARP packet generator	 \\
-ARP-SPOOF & Packet intercept	 \\
-Libpal & Packet intercept	 \\
-Aicmspend & ICMP packet flooding	\\
-\hline
-\end{tabular}
-\label{table:packetforging-tools}
-\end{table}
-
-\smallskip
-\subsubsection{Application layer}
-Application layer attacks use legitimate requests to use up all the resources of the victim.
-Unlike DDoS attacks, they are more subtle and more difficult to detect since they are using legitimate requests.
-
-\smallskip
-\section{Generating a realistic network load}
-A problem that exists with generating a realistic network load is the unwanted burstiness of real networks.
-Misconfigurations can result in traffic that looks similar to flooding attacks \cite{b10} and are hard to reproduce in small sandbox networks.
-Existing datasets often do not include such noise which result in unrealistic datasets \cite{b10}.
-
-\smallskip According to \cite{b10} it might be impossible to create a single representative background trace that contains benign and malicious traffic.
-The reason is that the expected traffic workload and characteristics are highly dependent on the network and its hosts.
-
-Tools such as Trident provide multiple strategies to create benign traffic \cite{b10}.
-
-\subsection{NIDS-based strategies}
-Using a wellknown network intrusion detection system (NIDS) such as Snort, these tools can groom a packet trace from a local site.
-This approach causes problems since benign packets that are flagged as malicious (false postives) are removed.
-
-\subsection{Synthetic generation strategies}
-Software robots emulate user behaviour and thus generate synthetic traffic. The idea behind it is that the robot only creates connections with known good hosts.
-
-\subsection{Trust-based strategies}
-The third strategy they support is by grooming a packet trace taken at a local site using connection heuristics.
-Examples of these heuristics are failure rates or scanning characteristics.
-This approach makes use of different connection characteristics of malicious and benign sources based on a model of malicious connection behaviour \cite{b10}.
-The advantage of using this strategy is that it is not influenced by a particular system (NIDS independence) and is based on transport level characteristics.
-The disadvantage is that a connection that is sufficiently similar to benign users will not be detected.
-
-\section{Summary and conclusion}
-Various aspects must be taken into account for the generation of realistic network traffic for the production of a dataset, which is subsequently used for supervised machine learning.
-In the course of this paper, cyber-attacks were explained in principle, various tools for carrying out the attacks were shown, and how realistic network traffic based on benign and malicious data packets can be generated.
-The topics covered are summarized at this point in order to emphasize the most important content again.
-
-Cyber-attacks generally try to exploit various vulnerabilities and carry out malicious actions.
-In order to provide an overview of the multitude of different attack methods, there are various approaches in the literature to classify the attacks into a few groups.
-The focus of this work is the classification model found in  \cite{b5}, which aims to enable a holistic view of attacks.
-Basically, nine different attack classes are described, including network attacks, denial of service attacks, information gathering attacks and various types of malware-based attacks.
-Subsequently, various attack targets (software, hardware and network), as well as vulnerabilities (in design, implementation, configuration) are explained in more detail.
-Tools for carrying out cyber-attacks are basically divided into two different groups, namely information gathering tools and attack launching tools.
-The former try to find out various information about the system to be attacked (open ports, operating system used, etc.).
-Sniffing and scanning tools can be used for this.
-Attack launching tools for performing denial of service attacks and for packet forging were identified.
-Furthermore, it is also possible to proceed on the basis of legitimate requests on the application layer.
-The creation of realistic network traffic based on benign and malicious data packets is a complex undertaking.
-By using tools, various strategies exist to generate harmless traffic, for example by using software robots.
-
-\section{Further research}
-Future work on this topic could try to look at attack methods on a more detailed level in order to provide more specific information.
-Furthermore, it makes sense to identify additional tools for carrying out cyber-attacks.
-There is also room for further research in the search for other ways of generating realistic network traffic including benign and malicious behavior.
+\section{Conclusion}
+@Sebastian
 
 \begin{thebibliography}{00}
-\bibitem{b1} N.
-Hoque, M.
-H.
-Bhuyan, R.
-C.
-Baishya, D.
-K.
-Bhattacharyya und J.
-K.
-Kalita, ``Network attacks: Taxonomy, tools and systems'', Journal of Network and Computer Applications, 2014.
-Jg., Nr.
-40, 2014.
-\bibitem{b2} S. Hansman,  ``A Taxonomy of Network and Computer Attack Methodologies'', 2003.
-\bibitem{b3} M. W. Pawar and J. Anuradha,  ``Network Security and Types of Attacks in Network'', Procedia Computer Science,  Nr. 48, 2015.
-\bibitem{b4} A. Gagandeep and K. Pawan,  ``Analysis of Different Security Attacks in MANETs on Protocol Stack A-Review'', International Journal of Engineering and Advanced Technology (IJEAT), 2012.
-\bibitem{b5} S. Hansman and R. Hunt,  ``A taxonomy of network and computer attacks'', Computers \& Security, Nr. 24, 2004.
-\bibitem{b6} W. Stallings und L. Brown,  ``Computer Security: Principles and Practice'',  3rd Ed., 2015.
-\bibitem{b7} Common Vulnerabilities and Exposures. [Online] Available at: https://cve.mitre.org/.
-\bibitem{b8} J. D. Howard,  ``An Analysis Of Security Incidents On The Internet'', 1995.
-\bibitem{b9} R. Spangler, ``Packet Sniffer Detection with AntiSniff'', 2003
-\bibitem{b10}  J.
-Sommers, V.
-Yegneswaran and P.
-Barford.
-``Toward Comprehensive Traffic Generation for Online IDS Evaluation''.
-Technical report, Dept.
-of Computer Science, University of Wisconsin, August 2005.
+\bibitem{b1} zzz
 \end{thebibliography}

### 2020-06-28 paper2/paper2.tex
Created paper2 and added first version of neural network chapter

@@
 \usepackage{cite}
-\usepackage{url}
 \usepackage{amsmath,amssymb,amsfonts}
@@
 
-\title{Network Data Collection for Artifical Intelligence
-}
+\title{Title}
 
-\author{\IEEEauthorblockN{Sebastian Lipp}
+\author{\IEEEauthorblockN{Boris Stampf}
+\IEEEauthorblockA{\textit{IT Security} \\
+\textit{FH Technikum Wien}\\
+Vienna, Austria\\
+cs19m006@technikum-wien.at}
+\and
+\IEEEauthorblockN{Maximilian Wech}
 \IEEEauthorblockA{\textit{IT Security} \\
@@
 Vienna, Austria \\
-cs19m032@technikum-wien.at}
-\and
-\IEEEauthorblockN{Damir Marijanovic}
-\IEEEauthorblockA{\textit{IT Security} \\
-\textit{FH Technikum Wien}\\
-Vienna, Austria \\
-cs19m031@technikum-wien.at}
-\and
-\IEEEauthorblockN{Boris Stampf}
-\IEEEauthorblockA{\textit{IT Security} \\
-\textit{FH Technikum Wien}\\
-Vienna, Austria \\
-cs19m006@technikum-wien.at}
+cs19m020@technikum-wien.at}
 }
@@
 \begin{abstract}
-Collecting real network data for further processing is an important topic for the training of neural networks which filter suspicious from ordinary traffic.
-This paper focuses mainly on research in the areas of network simulation, network recording and data extraction which should serve as input for the following papers.
+@TBA
 \end{abstract}
@@
 \begin{IEEEkeywords}
-artifical intelligence, neural networks, computer networks, network security, data collection
+artificial intelligence, neural networks, computer networks, network security
 \end{IEEEkeywords}
@@
 \section{Introduction}
-Neural networks which filter suspicious network traffic need a training data set of high quality.
-This data set could be produced from real network traffic but this can be difficult to implement.
-Another option considers network simulators which can be easily configured depending on the selected scenarios.
+@Boris
 
-This paper is split in three chapters.
-The first chapter compares available network simulators and their software interfaces and explores possibilities for their automated setup based on the scenario and parameters.
-The second chapter focuses on tools to record the network traffic.
-In the third chapter methods of feature extraction and label definition are covered.
-This information serves as foundation for the further implementation of a system for automated data collection for artifical intelligence.
+\section{Feature Extraction}
+@Boris
 
-At this point all our considerations are based on theoretical research  without practical implementation which will come to life in further work.
-Therefore this paper sets a direction and foundation for the next steps and shows possible solutions on a high level.
+\section{Neural Network}
+The created dataset will be used in the further course of this paper to build a neural network.
+With this it should be possible to detect anomalies in network traffic (e.
+g.
+DoS attacks) but also harmless data packets.
+This is a continuation of  \cite{max1}, in which an attempt was made to create a neural network for intrusion detection on the basis of the publicly available CSE-CIC-IDS2018 dataset  \cite{max2}.
+The knowledge gained at that time is used, the source codes are reused accordingly and slightly adapted to the new dataset.
+In principle, the performance of neural networks over different datasets is to be checked.
+The following scenarios will be considered:
+\smallskip
 
-\section{Simulation}
+Scenario 1) What performance can be achieved with a neural network when it is trained and tested with the dataset created in this work? 
+\smallskip
 
-Before network simulators can be compared the requirements have to be defined.
-Because the produced network traffic should be close to real traffic it would be advisable to choose a network simulator which works with the simulation of real network equipment and can also integrate real hardware or virtual machines into the simulation which is necessary to redirect the network traffic to the recorder and data extractor.
-This is important to consider since other simulators like NS3 provide discrete-event network simulation which is a needless complexity overhead for this kind of application.
-It would be desirable if they are affordable and provide an easy-to-use interface to create and control specific network topologies through a programming language like Python to automate their creation.
-Moreover they should be resource-efficient and easy to deploy.
-The provisioning of the simulation hosts has to be simple and also scalable to be prepared for simulating denial-of-service attacks.
+Scenario 2) How well does the neural network created in Scenario 1 work when tested with the CSE-CIC-IDS2018 dataset  \cite{max2}? 
+\smallskip
 
-The next sections cover possible network simulators, system environments and the provisioning of the infrastructure.
+Scenario 3) How well does the neural network trained in \cite{max1} work when tested with the dataset created in this work?
+\smallskip
 
-\subsection{Comparing common network simulators}
-There are many solutions available on the market to simulate networks.
-Popular network simulators which possibly fulfill the requirements above are for example VIRL, GNS3 or EVE-NG.
-This section focuses on the comparison of these simulators and the selection of the best solution for our use case.
-\cite{b1}
+It is suspected that there are differences in the intrusion detection datasets because, for example, cyber attacks are not always executed or recorded in the same way or the network infrastructure is structured differently.
+The aim is to determine how sensitive neural networks react to this and are thus dependent on the data with which they have been trained.
+This chapter is designed in such a way that the various steps for building up the neural network are first described and then the results obtained are examined in detail.
 
-\begin{figure}[htbp]
-\centerline{\includegraphics[scale=0.55]{virl.png}}
-\caption{Sample network topology in VIRL \cite{b2}}
-\label{virl}
-\end{figure}
+\subsection{Setup}
+For the practical implementation, a notebook with 16GB RAM and a 2.7 GHz dual-core processor was used.
+This hardware was sufficient for the fulfillment of the task.
+The training of the neural network took about 4 hours.
+PyCharm Community Edition 2020.1 was used as a Python development environment on Windows 10.
+Packages were basically the same as in \cite{max1}.
+The Seaborn package 0.10.1 was also used to create heatmaps.
 
-VIRL (Cisco Virtual Internet Routing Lab) is a network modeling and simulation environment provided by Cisco which offers the ability to use simulation models of real Cisco platforms, integrate virtual machines and connect real networks.
-Furthermore it provides a RESTful API to configure networks without a graphical interface, sufficient documentation and scalability as well as easy deployment through prepared images.
-It is not for free and only supports VMware or bare-metal installs.
-Simulation models of other vendors are not supported.
-Tests in virtual machines unveiled a good responsiveness but a long topology loading time with high CPU load.
-Fig.~\ref{virl} shows a sample network topology created via the graphical user interface of VIRL.
-\cite{b1} \cite{b3}
+\subsection{Preprocessing}
+This process is based on the one from\cite{max1}.
+This involves preparing the created dataset for training the neural network accordingly.
+A difference compared to \cite{max1} is that this time no headers were in the middle of the created CSV file.
+For this reason, the "skip\_rows" parameter did not have to be used in the "read\_csv" pandas function.
+In addition, identical entries in the dataset (duplicates) are also removed and a corresponding data type is specified for each feature.
+The conversion of the labels into a numeric format also takes place.
+At this point it should be noted that due to time constraints, not all types of attack from \cite{max2} could be reproduced in this work.
+This can be seen from the fact that some (numerical) labels are not represented in the created dataset.
+In the context of future work, the missing types of attack can be added.
+It is important that the labels in the created dataset must have the exact same numerical representation as in \cite{max1}.
+This means, for example, that benign network traffic must be labeled with "0" again.
+Furthermore, no types of attack were implemented in the context of this work that are not available in the CSE-CIC-IDS2018 dataset \cite{max2}.
+This is because the neural network from \cite{max1} could not recognize them and thus no performance comparison would be possible.
+For this reason, attacks from the CSE-CIC-IDS2018 dataset \cite{max2} were selectively reproduced.
+It should also be noted that in \cite{max1} some attacks of the publicly available dataset were not included.
+
+\smallskip
+The following Table \ref{table:overviewdatasets} provides an overview of both datasets:
+\begin{table}[htbp]
+\caption{Overview of the two datasets } 
+\centering
+\begin{tabular}{ | c | c | c | }
+\hline
+Label & \multicolumn{1}{|p{2cm}|}{\centering Number of data \\ points in the dataset  \\ created in this work}  & \multicolumn{1}{|p{2cm}|}{\centering Number of data \\ points in the \\  CSE-CIC-IDS2018 dataset \cite{max2}}\\
+\hline
+benign (0) & 63.264 & 10.180.908 \\
+\hline
+ddos attacks-loic-http (1) & Not available & 575.364 \\
+\hline
+ddos attack-hoic (2) & 193.729	& 198.861 \\
+\hline
+dos attacks-hulk (3) & Not available & 145.199 \\
+\hline
+bot (4) & Not available & 144.535 \\
+\hline
+ssh-bruteforce (5) & 29.182 & 94.048 \\
+\hline
+dos-goldeneye (6) & 7.368.370	 & 41.406 \\
+\hline
+dos attacks-slowloris (7) & Not available & 9.908 \\
+\hline
+ddos attack-loic-udp (8) & Not available & 1.730 \\
+\hline
+\end{tabular}
+\label{table:overviewdatasets}
+\end{table}
 
@@
 
+While the CSE-CIC-IDS2018 dataset \cite{max2} has a total of 10,182,638 data points (6.41GB distributed over multiple CSV files), the dataset created in this work consists of 7,654,545 data points (4.1GB, only one CSV file).
+When looking at Table 1, it is noticeable that both datasets are unbalanced and there is one label for which comparatively many data points exist (dataset created in this work: "dos-goldeneye", CSE-CIC-IDS2018 dataset: "benign").
+The preprocessing process of the final dataset takes about 10 minutes.
+This is comparatively faster than in \cite{max1}, which could be due to the smaller file size and the fact that the current dataset consists of only one CSV file (there is no need to merge several files).
+Finally, the dataset was divided into a training and a test set at a ratio of 80:20.
+While the former is used to train the neural network, the latter is used to determine the performance of the neural network (with data not used for training).
+
+\subsection{Training the neural network}
+After the preprocessing of the data has been completed, the next task is to train and test the neural network.
+The architecture recommended in \cite{max1} (three hidden layers with 69 neurons each) and the hyperparameters (Learn rate: 0.
+0001, Optimizer: adam, Weight initialization: normal, Activation function: relu, Number of neurons per layer: 69, Epochs: 850, Batch size: 256) are reused.
+In addition, "MinMaxScale" was used to normalize the data and a weight value of 500 was set for benign" traffic.
+At this point it is worth mentioning that attempts have been made to adjust individual parameters such as the learn rate or the weight value of benign traffic in order to achieve a better result.
+However, this has not had any positive effects.
+Future work could deal with this in more detail and try to find optimal hyperparameters by an extensive gridsearch.
+Of course, this would require appropriate hardware resources.
+
+\smallskip
+Training of the neural network with the training set took 4 hours and was stopped after 129 epochs.
+As in \cite{max1}, "EarlyStopping" with a value of 50 epochs was used.
+This means that after 50 epochs without improvement, the learning process is terminated to avoid overfitting.
+
+\subsection{Results}
+In the further course of this chapter, the aim is to present and evaluate the results achieved.
+To determine how good the classifications of machine learning algorithms are, an appropriate measure is required.
+The following metrics are relevant \cite{max3}:
+\smallskip
+
+Accuracy: How likely is the model to make a correct prediction?
+\smallskip
+
+Precision for class A: How likely is the classification to be correct if the model predicts class A?
+\smallskip
+
+Recall for class A: How much data of a class A can the model correctly predict?
+\smallskip
+
+Both the dataset created in this work and the CSE-CIC-IDS2018 dataset \cite{max2} are unbalanced (different amounts of data per class).
+For this reason, the use of Accuracy does not seem to be appropriate.
+If a very large amount of data were available for a class A and not for classes B and C, then the accuracy would be very dependent on class A.
+Instead, it is advisable to graphically display the classification results of the neural network in a confusion matrix and to calculate precision and recall respectively.
+The following text analyses the results of the three scenarios in the introduction:
+Figure  \ref{fig:cm1} shows the result of the first scenario in which a neural network was trained and tested with the dataset created in this work.
+The classification results are largely good.
+However, it is noticeable, that only 62.5\% of all data of the "ssh-bruteforce" attack type can be correctly classified (comparatively low recall).
+The precision value for "benign" traffic is also comparatively lower, which means that it is only 86\% correct if the neural network predicts this class.
+All other precision and recall values are well over 95\%.
+
 \begin{figure}[htbp]  
-\centerline{\includegraphics[scale=0.13]{gns3.png}}
-\caption{Sample network topology in GNS3 \cite{b4}}
-\label{gns3}
+\centerline{\includegraphics[scale=0.65]{NeuesModellNeueDaten.png}}
+\caption{Confusion Matrix for Szenario 1}
+\label{fig:cm1}
 \end{figure}
 
-GNS3 (Graphical Network Simulator 3) is a very popular free open source software to emulate and test networks.
-It offers the integration of virtual machines and real networks and supports models of various vendors but does not offer pre-installed device images due to license issues.
-It also provides a RESTful API to set up network topologies, a big community and scalability.
-The software can be easily installed through packages on all operating systems and supports all hypervisors as well Docker containers.
-Performance tests showed a long topology loading time with high CPU load but a good response through the user interface.
-Fig.~\ref{gns3} shows a sample network topology created with GNS3.
-\cite{b1} \cite{b5}
+Next, the neural network trained in Scenario 1 was reused and tested with the CSE-CIC-IDS2018 dataset \cite{max2}.
+There are many more classes (attack types) in this dataset than in the one used to train the neural network.
+Thus, the model would not be able to recognize these classes.
+For this reason, it is necessary to reduce the CSE-CIC-IDS2018 dataset \cite{max2} to the known four classes ("benign", "bot", "ssh-bruteforce" and "dos-goldeneye").
+The reduced dataset is then used to test the neural network.
+The Confusion Matrix in Figure  \ref{fig:cm2} shows that the classification result is anything but good.
+On the positive side, the precision for benign traffic is very high, so if the model predicts this class, the prediction is 99.4\% correct.
+However, only 24.2\% of all data in this class can be correctly predicted.
+All other precision and recall values are or move around 0\%.
+It is worth mentioning that the "ssh-bruteforce" attack cannot be recognized correctly once.
+Scenario 1 already indicated a worse classification result for this class.
+Both "bot"- and "dos-goldeneye" attacks could be detected at least a few times, but to a very small extent.
+Overall, this suggests that there may be differences in the two datasets that significantly affect the classification outcome of the neural network.
+Thus, there is a suspicion that the machine learning model is overfitting.
 
 \begin{figure}[htbp]  
-\centerline{\includegraphics[scale=0.19]{eve-ng.png}}
-\caption{Sample network topology in EVE-NG \cite{b6}}
-\label{eve-ng}
+\centerline{\includegraphics[scale=0.65]{NeuesModellAlteDaten.png}}
+\caption{Confusion Matrix for Szenario 2}
+\label{fig:cm2}
 \end{figure}
 
-EVE-NG (Emulated Virtual Environment - Next Gen) is a network emulator which comes in a free and a professional edition and offers the interaction with real networks.
-It supports multiple vendors but does not provide device images.
-Advanced features like Docker, configuration management or Wireshark integration only come with the purchase version.
-Moreover a RESTful API and images for the installation in virtual machines are available as well as a reasonable documentation.
-The software is scalable and performs well in virtual machines with a short topology loading time and provides an intuitive interface as shown in Fig.~\ref{eve-ng}.
-\cite{b1} \cite{b7}
-
-All solutions are very similar in their functionalities but distinguish in their vendor support, performance, documentation and price.
-VIRL only supports Cisco devices but comes with device images like GNS3 and EVE-NG support various vendors but do not ship images.
-But many proprietary and open source images are available on the internet for free.
-VIRL and EVE-NG are propriatary like GNS3 is free and open source.
-EVE-NG performs better in virtual machines than VIRL and GNS3.
-
-Regarding these factors GNS3 fullfills our requirements the most because it is free, supports all vendors and comes with a big community and documentation.
-Our second option could be EVE-NG in the professional edition which also supports all vendors.
-
-\subsection{Selecting the deployment environment}
-
-Next steps include the integration of the selected network simulator in a scalable environment which can be set up easily.
+In the third scenario, the neural network from \cite{max1}, which was trained with CSE-CIC-IDS2018 Dataset \cite{max2}, is now tested with the dataset of this work.
+This continues the low classification success (see Figure \ref{fig:cm3}).
+In the class "benign", 99.8\% of all data are correctly predicted (recall), but the precision is noticeable low (0.
+008\%).
+This means that the neural network recognizes many data points as "benign", but these are mostly other classes.
+Conversely, this is the case with "bot", where only few data of this class are correctly recognized, but whenever the neural network recognizes this class, it is a correct prediction.
+All other precision and recall values are or move around 0\%.
+It is particularly noticeable that "bot"-, "ssh-bruteforce"- and "dos-goldeneye" attacks are mostly classified as benign traffic.
+Furthermore, it can be seen that "ssh-bruteforce"- and "dos-goldeneye" attacks cannot be correctly detected once; bot attacks can be correctly predicted nine times in total.
+This makes it clear how weak the classifications of the neural network trained in \cite{max1} are with the dataset created in this work.
+Also worth mentioning is that for some data points the classes "ddos attacks-loic-http", "dos attacks-hulk" and "dos attacks-slowloris" are predicted.
+These are not available in the dataset created in this work, but in the CSE-CIC-IDS2018 dataset \cite{max2}, with which the neural network was trained.
+For this reason, this is not surprising.
 
 \begin{figure}[htbp]  
-\centerline{\includegraphics[scale=0.5]{docker_vm.png}}
-\caption{Virtual machines vs. Containers \cite{b8}}
-\label{docker-vm}
+\centerline{\includegraphics[scale=0.65]{AltesModellNeueDaten.png}}
+\caption{Confusion Matrix for Szenario 3}
+\label{fig:cm3}
 \end{figure}
 
-Hypervisors like VirtualBox or VMware provide the possibility to set up full virtual computers on a physical host (see Fig.~\ref{docker-vm}a) while Docker containers are encapsuled sandboxes on an existing operating system (see Fig.~\ref{docker-vm}b).
-It is clear that containers are more resource efficent but in virtual machines the systems are better isolated.
-In our case we could use both solutions because the network simulators are also capable to run in Docker containers.
-If we focus on resource efficency and scalability also Kubernetes comes in our minds which is an orchestrator for deploying Docker containers on multiple nodes but it is a more complex configuration overhead and needs certain amount of hardware to unfold its potential also for simple scenarios.
-Because we want an affordable solution which can be used in most of the cases also on our computers with some exceptions, we could use a tool which can create virtual machines as well as Docker containers to be ready for the future.
-Regarding the corner cases servers or a Kubernetes cluster could be rented temporarily on cloud platforms like AWS to support high-load scenarios e.g.
-big denial-of-service attacks.
-\cite{b8}
-
-A popular solution to bootstrap development environments in virtual machines is called Vagrant which provides lightweight base images and an interface to automate the configuration of the virtual machines.
-Moreover it can be connected to provisioning tools to further configure host systems.
-Vagrant supports VirtualBox or VMware but also Docker and can be extended to support cloud services like AWS or even to install a whole Kubernetes cluster.
-It gives us the power to start with simple configurations and scale them up to more sophisticated ones using containerization and complex orchestrators.
-\cite{b9} \cite{b10}
-
-This means we would be ready to create system environments on our host or in the cloud and install base systems on them.
-The next section focuses on the automated setup of the network simulator, recorder and data extractor within virtual machines or Docker containers as well as the configuration of network topologies based on the selected scenarios and parameters to produce, record and extract network traffic for further processing in neural networks to improve their strike rate.
-
-\subsection{Provisioning of the infrastructure}
-
-Vagrant supports multiple provisioners like Ansible, Terraform or simple bash scripts to configure the host systems and can also be set up through them.
-For the beginning we could use Ansible through Vagrant to configure the host system.
-Vagrant can set up multiple virtual machines or Docker containers on one host.
-If we would need to set up machines on multiple hosts we could come back to tools like Terraform for example.
-\cite{b9} \cite{b10} \cite{b11}
-
-If all the possible network scenarios are defined these network topologies have to be deployed through our deployment system.
-This means we could define the basic host setup and the network configurations for all the scenarios in Ansible templates which could be later adjusted via command line parameters.
-The topology can then be deployed by executing a shell command.
-This includes the deployment of the network recorder and the data extractor.
-In other words one or multiple scenarios can be started and connected to one recorder and data extractor to combine simultaneous attacks and everything will be deployed through a shell command with static configuration in templates and dynamic configuration with command line options.
-It will be also possible to integrate real hardware into the simulation.
-\cite{b9}
+Scenarios 2 and 3 show that if in this context a neural network is tested with a different dataset than the one it was trained with, then the classification performance is poor.
+There can be many reasons for this.
+The suspicion, however, is that there are differences in the datasets and that this affects the predictions of the neural networks.
+One approach is to perform a correlation analysis.
+This makes it possible to determine the statistical relationship between two features (variables) of a dataset.
+The correlation coefficient is to be calculated, which indicates the strength of the correlation in the interval [-1.1].
+A positive correlation means the larger the variable A, the larger the variable B.
+The opposite is the negative correlation, in which the larger the variable A is, the smaller the variable B.
+If the correlation coefficient is 0, this is called a neutral correlation and there is no relationship between the variables \cite{max4}.
+The idea is to determine the correlation between the features (variables) for each dataset individually.
+From this, graphical representations (heatmaps \cite{max5}) can be created in order to be able to compare whether correlation differences can be detected in the datasets.
+Since the number of features is high in both datasets, it makes sense to use only a subset of the features for the analysis.
+Thus, the correlation between the first 36 features is determined for both datasets.
+Figure \ref{fig:hm1} and Figure \ref{fig:hm2} each show the results for the two datasets.
+Clear differences are recognizable.
+It is worth mentioning that Figure \ref{fig:hm1} shows numerous negative correlations, which are not at all or hardly recognizable in Figure \ref{fig:hm2}.
+It is also noticeable that in Figure \ref{fig:hm2}, white lines occur, meaning that a feature in the dataset created in this work always has a value of zero.
+This shows that there are differences in the datasets.
 
 \begin{figure}[htbp]  
-\centerline{\includegraphics[scale=0.28]{design_flow.png}}
-\caption{Data collection workflow}
-\label{design-flow}
+\centerline{\includegraphics[scale=0.8]{Heatmap-AlteDaten-Filtered-0-35.png}}
+\caption{Heatmap of the first 36 features of the CSE-CIC-IDS2018 Dataset}
+\label{fig:hm1}
 \end{figure}
 
-Fig.~\ref{design-flow} shows a possible simulation workflow.
-The input parameters are the type of scenario(s) and the respective settings which deploys hosts including the base systems and the simulators applying the network topologies.
-Moreover the network recorder and data extractor could be deployed to gather the network traffic and extract the features to create data sets which could be collected in a central database.
-
-A real test case would then consist of multiple scenarios of attacks and usual network traffic recorded, extracted and collected in a central place.
-The next chapter focuses on recording the generated data.
-
-\section{Recording}
-
-Before the network traffic can be analysed, it has to be recorded first.
-Therefore, this chapter will give an overview of the available third-party software which can be used to record all the incoming network traffic.
-The functional requirements for a tool are chiefly the pace of handling the network traffic.
-All of these tools are client-based tools.
-This means that either the network target for the traffic has to be a client or the e.g.
-switch has to clone a port with all the incoming and outgoing traffic and direct all the data to a client running one or more of these tools.
-In some cases, analysing of network traffic can take much longer than the data is sent or received by the network which means, that all the data need to be buffered or stored.
-In this case a packet-to-disk recorder is needed.
-This device can capture large amount of network traffic with a data rate of up to 40 Gbit/s and write it into files.
-
-\subsection{pcap Library}
-
-Every tools needs APIs\footnote{Application Programming Interface} to capture all the traffic sent through the network interface.
-Therefore, libraries were made for both Unix-like and Windows based operating systems.
-While \textit{libpcap} is used for Unix-like operating systems, a Windows system needs the library called \textit{WinPcap}.
-In 2013 the last release (4.1.3) of \textit{WinPcap} was published.
-Since then \textit{Npcap} acts as a substitute\cite{winpcaporg}.
-With a pcap library, software has the ability to tap all the network traffic from the selected interface.
-Based on the possibility to include the library in \textbf{C} and \textbf{C++}, \textit{pcap} can be used very easily in software.
-With the release of \textit{libpcap} 1.0.0, the zero-copy mechanism was implemented which allows the software to read the network traffic directly without copying the data from kernel space to user space.
-
-\subsection{Network Recording Tools}
-
-In this section we will itemise five well known and useful network recording tools. Some of the tools are console based tools, others have a GUI, or both.
-
-\subsection*{tcpdump}
-
-Especially in the world of Unix based operating systems, \textit{tcpdump} is one of the best known tools for monitoring and sniffing network traffic.
-\textit{tcpdump} is an exclusively command line based tool.
-This property makes \textit{tcpdump} a tool for advanced command line users only.
-But once you get familiar with the commands and options of \textit{tcpdump}, it will get a  a very efficient and powerful tool for many applications.
-\textit{tcpdump} uses the \textit{libpcap} network library of Unix based operating systems for capturing network traffic.
-
 \begin{figure}[htbp]  
-\centerline{\includegraphics[scale=0.4]{tcpdump.png}}
-\caption{tcpdump Network Capture}
-\label{tcpdump}
+\centerline{\includegraphics[scale=0.8]{Heatmap-NeueDaten-0-35.png}}
+\caption{Heatmap of the first 36 features of the dataset created in this work}
+\label{fig:hm2}
 \end{figure}
 
-With the wide set of options, \textit{tcpdump} is a pretty agile tool for dumping and analysing network traffic. A few of the options are listed below \cite{tcpdumporg}.
-
-\begin{itemize}
-\item \textbf{-i eth0} the traffic from the interface \textit{eth0} is captured
-\item \textbf{-w example.dump} all the output is written into a declared file
-\item \textbf{-e} the ethernet header will recorded too
-\item \textbf{-r example.dump} parsing the file \textit{example.dump} into tcpdump
-\end{itemize}
-
-If \textit{tcpdump} is compiled with cryptography, it is able to decrypt IPsec ESP packets in real time \cite{goyal2017}.
-For the Windows enthusiasts a Windows derivative of \textit{tcpdump} is available.
-It is called \textit{WinDump}.
-\textit{WinDump} uses the \textit{WinPcap} library, which provides the same API for Windows operating systems as \textit{libpcap} for Unix-like systems.
-This means, network traffic captured, e.g.
-on a Ubuntu system can be analysed on a Windows 10 Client afterwards.
-
-\subsection*{Wireshark}
-
-\textit{Wireshark} is for good reason the best known cross-platform networking tool available.
-In 2006 \textit{Wireshark} was started as a fork of the software called \textit{Ethereal} by the Wireshark-Community due to trademark issues \cite{asrodiapatel2012}.
-The performance of this tool is powerful and the set of functions is neatly arranged.
-By using the GTK toolkit, \textit{Wireshark} has implemented a GUI \footnote{Graphical User Interface}\cite{suribatra2012}.
-The GUI is separated typically into three segments: the traffic overview, and the details and raw data of the selected packet.
-
-\begin{figure}[htbp]
-\centerline{\includegraphics[scale=0.48]{wireshark.png}}
-\caption{Wireshark Network Capture}
-\label{wireshark}
-\end{figure}
-
-A good asset of \textit{Wireshark} is its available set of filters for capturing and displaying network data.
-With the help of plugins, \textit{Wireshark} can even handle with new or dedicated protocols.
-Another big asset of \textit{Wireshark} is its capability of capturing raw USB traffic from an USB port.
-By highlighting the traffic with different colours as shown in Fig.~\ref{wireshark} it's easier to scan for required traffic.
-
-\subsection*{netsniff-ng}
-
-\textit{netsniff-ng} is a another Unix based tool for network analysis.
-\textit{netsniff-ng} was the first tool which had the ability to analyse the whole traffic with a zero-copy procedure, which means that \textit{netsniff-ng} uses the same data as the kernel does.
-Based on this functional principle, \textit{netsniff-ng} is one of the best performing tools for network sniffing.
-\textit{netsniff-ng} is like \textit{tcpdump} exclusively an console based software.
-The set of options contains 38 capabilities on the latest version 0.6.4 of \textit{netsniff-ng}.
-
-\subsection*{Capsa}
-
-The fourth tool is called \textit{Capsa} from \textit{Colasoft} founded 2001 in United States.
-\textit{Capsa} is suitable for real-time network data capturing both LAN and Wireless Interfaces.
-A big advantage of \textit{Capsa} is the ability to evaluate the network traffic into graphical output like diagrams.
-This feature can help identifying the source of undesirable network traffic quickly by a human.
-The free version of \textit{Capsa} is heavily restricted regarding available functions.
-
-\begin{figure}[htbp]
-\centerline{\includegraphics[scale=0.30]{capsa.png}}
-\caption{Graphical Evaluation of Network Traffic with Capsa \cite{firewallcx}}
-\label{capsa}
-\end{figure}
-
-\subsection*{EtherApe}
-
-Yet another tool for network monitoring is called \textit{EtherApe}.
-\textit{EtherApe} is an Unix based software with a meaningful graphic interface and graphical data output of network traffic.
-With the help of graphical illustration anomalies like a DDoS attack can be detected easily as shown in Fig.~\ref{etherape}.
-Every host is shown as a node in circle, every connection is shown as link between two nodes and every protocol has an own colour for the link.
-With higher network traffic between two hosts the link between nodes is shown larger.
-
-\begin{figure}[htbp]
-\centerline{\includegraphics[scale=0.37]{etherape.png}}
-\caption{Graphic Representation of Network Traffic \cite{etherapewiki}}
-\label{etherape}
-\end{figure}
-
-\section{Data Extraction}
-
-Datasets for intrusion detection are often based on network packet data captured by sniffers, which can be stored in pcap files (as done in the UNSW-NB15 \cite{Nb2015} and CSE-CIC-IDS2018 \cite{Ids2018} datasets).
-Before a machine learning algorithm can be employed, features have to be extracted from the captured data.
-A wide range of different features is described in literature.
-For the sake of clarity, these features are divided into groups which are described in the following subsections.
-
-There also exists a large number of techniques for dimensionality reduction.
-The 40 features of the NSL-KDD data set could be reduced to 30 or 16 features, with no or low degradation in detection power.
-The benefits of a reduced feature set are faster processing, less memory consumption, and in some cases even higher detection precision, because some classifiers have problems with irrelevant or correlated features.
-This could be achieved by combining the following methods \cite{vasquez2015}:
-\begin{itemize}
-	\item Weight by maximum relevance
-	\item Minimum redundancy maximum relevance
-	\item Significance analysis for microarrays
-	\item Least absolute selection and shrinkage operator
-	\item Stability selection
-	\item Stepwise regression
-\end{itemize}
-
-Other methods which can be used for that purpose are based on association rule mining (ARM) \cite{moustafa2015}, sequential forward selection (SFS) \cite{dehghani2010} \cite{xu2015}, or the T-test \cite{sopuru2019},
-
-When selecting the features, their limitations should be considered as well.
-Some features can easily be spoofed by attackers, while others are unable to detect encrypted malware traffic or cause performance problems.
-The development of new evasion techniques (e.g., disguising malicious traffic in widely used protocols such as HTTP) make the detection even harder \cite{celik2015}.
-
-Another aspect which has been examined is the stability of features in terms of time (year 2000/2001 vs.
-2008) and location (Internet core vs.
-Internet edge).
-In this regard, the packet size can be considered as stable feature \cite{este2009}.
-The suitability of features also depends on the type of network which is analyzed.
-For example, some industrial control system networks are relatively static regarding IP address allocation, the number of open connections and the ports, protocols and settings they use.
-Therefore, features based on these attributes may provide good results in that particular environment \cite{mantere2013}.
-
-\subsection{Flow statistical features}
-
-A common approach is to group packets into flows, and then calculate statistical features based on these flows.
-Generally, a flow is a group of packets with the same source and destination IP addresses and ports, and the same protocol.
-However, a flow can also be terminated by a TCP FIN packet or a timeout \cite{Ids2018}.
-Statistical features of network flows include the following:
-\begin{itemize}
-	\item Flow duration \cite{Nb2015} \cite{Ids2018}
-	\item Time between two flows or packets \cite{Ids2018}
-	\item Time between SYN and SYN/ACK, and between SYN/ACK and ACK packets during TCP connection setup \cite{Nb2015}
-	\item Active time before a flow becomes idle and vice versa \cite{Ids2018}
-	\item Jitter \cite{Nb2015}
-	\item Number of packets in total, with at least 1 byte of payload, or with certain protocol flags \cite{Ids2018} \cite{waizumi2007} \cite{celik2015}
-	\item Number of fragmented packets \cite{waizumi2007}
-	\item Number of retransmitted or dropped packets \cite{Nb2015}
-	\item Size of packets, payload, headers or segments \cite{Ids2018} \cite{celik2015}
-	\item Number of bytes/packets transferred per second, in bulk, or in the initial TCP window \cite{Ids2018} \cite{celik2015}
-	\item Number of RTT samples \cite{celik2015}
-	\item Time to live value \cite{Nb2015}
-	\item TCP window advertisement value \cite{Nb2015}
-	\item TCP base sequence number \cite{Nb2015}
-	\item Ratio between download and upload \cite{Ids2018}
-	\item Ratio between maximum and minimum packet size \cite{celik2015}
-	\item Number of flows with the current flow's IP address or port number \cite{waizumi2007} \cite{Nb2015}
-\end{itemize}
-Most of these features can be derived by statistical calculations such as total, minimum, maximum, mean, median, standard deviation, variance, skewness and kurtosis \cite{Ids2018} \cite{celik2015, xu2015}.
-Also, the features can be extracted from bidirectional flows, or separately for forward and backward directions \cite{Ids2018}.
-
-The open-source Java software CICFlowMeter can be used to extract features of network flows (which end either by TCP FIN packet, or a configurable timeout) from pcap files.
-It allows to select from a list of available features, but can also be extended to support new features \cite{cicflowmeter}, and has therefore been used in various applications \cite{Ids2018} \cite{lashkari2017} \cite{sopuru2019}.
-
-\begin{figure}[htpb]
-\centerline{\includegraphics[scale=0.35]{hast-ids.png}}
-\caption{Feature learning process of HAST-IDS \cite{wang2018}}
-\label{hast-ids}
-\end{figure}
-
-The hierarchical spatial-temporal feature-based intrusion detection system (HAST-IDS), which is shown in fig.~\ref{hast-ids}, also extracts flow-based features.
-However, it first transforms the bytes of network flows (HAST-I) or packets (HAST-II) into images using one-hot encoding (OHE), and then uses deep convolutional neural networks (CNNs) and long short-term memory (LSTM) networks to learn features based on these images \cite{wang2018}.
-
-\subsection{Timeslot features}
-
-Timeslot features are extracted by counting the number of certain events in fixed time intervals.
-For the detection of DoS and probing traffic, the following timeslot features are proposed \cite{waizumi2007}:
-\begin{itemize}
-	\item Number of TCP, UDP and ICMP packets
-	\item Number of bytes sent and received through all TCP connections
-	\item Number of TCP ports
-	\item Number of occurences of TCP flags
-	\item Number of DNS packets
-	\item Number of fragmented packets
-	\item Number of values of the four fields of IP addresses
-\end{itemize}
-
-\subsection{Behavioral features}
-
-A different approach is to define features which measure specific behavioral differences between benign and malicious software.
-
-One system following that approach is called "traffic aggregation for malware detection".
-It defines three characteristics which can be combined to distinguish between benign and malware traffic \cite{yen2011}:
-\begin{itemize}
-	\item Common destinations (assuming that compromised systems contact different hosts than uncompromised systems)
-	\item Similar payload (assuming that command-and-control traffic uses common protocol syntax)
-	\item Common software platforms (assuming that malware usually relies upon a specific operating system and also specific applications, such as web browsers)
-\end{itemize}
-
-The UNSW-NB15 datasets contains mostly flow statistical features, but also a few features based on behavior \cite{Nb2015}:
-\begin{itemize}
-	\item A binary value denoting if the source and destination IP addresses and port numbers are equal
-	\item Service (HTTP, FTP, SMTP, SSH, DNS, or IRC)
-	\item Pipelined depth into the connection of HTTP request/response transaction
-	\item Actual uncompressed content size of the data transferred from the server's HTTP service
-	\item Number of flows containing HTTP GET and POST methods
-	\item A binary value denoting if a FTP session is accessed by user and password
-	\item Number of flows containing FTP commands
-\end{itemize}
-
-Reference \cite{li2013} lists the following features which can be used for detecting communication with known and unknown botnets:
-\begin{itemize}
-	\item Accessing the main and backup DNS server at the same time (benign software would only access the backup DNS server if the main DNS server is not available)
-	\item Resolving a specific domain name periodically
-	\item Accessing a fast-flux service network (FFSN)
-	\item Downloading malicious binary code (which can be checked using antivirus software)
-	\item Scanning for open ports
-	\item Periodically creating null TCP connections (zero TCP window or IRC PING/PONG connections which are used for contacting the command and control server)
-\end{itemize}
-
-Similarily, the character code distribution of payloads can be used to detect buffer overflow traffic (where some character codes have exceptionally high frequency) \cite{waizumi2007}.
-
-\subsection{Host-based features}
-
-The previous sections examined different features which can be extracted from network traffic, but there are also host-based data which can be used for intrusion detection.
-These include operating system event logs \cite{Ids2018} and other log files (e.g., firewall, mail and FTP logs) \cite{abad2003}.
-
-A different approach is the use of features derived from raw system call traces.
-The advantage of system call traces over log files is that they contain only raw information about the interaction between programs and the operating system kernel, whereas log files contain data interpreted by various programs, and potentially include large amounts of irrelevant data.
-There are multiple methods of analysing system calls.
-Some only look at the pattern of system calls, while others also look at the arguments passed to these system calls.
-Also, there is a distinction between syntactic features, which depend only on contiguous patterns of system calls, and semantic features, which additionally consider the meaning of system call arguments or groups of discontiguous system calls.
-The advantage of the semantic approach is that it is able to counter mimicry attacks, where the payload is modified by the attacker with the aim that the system calls it executes cannot be classified as being malicious \cite{creech2014}.
-
-\subsection{Label definition}
-
-In addition to the feature vectors, datasets to be used for supervised learning need to include labels.
-For the purpose of intrusion detection, each record (e.g., network flow) has to be labelled  binary as benign or malicious traffic, or nominally to distinguish between certain attack types.
-For example, the UNSW-NB15 dataset defines a binary label (0 for normal and 1 for attack) and a nominal label ("Normal" for benign traffic and "Fuzzers", "Analysis", "Backdoors", "DoS", "Exploits", "Generic", "Reconnaissance", "Shellcode" or "Worms" for the various attack types) \cite{Nb2015}.
-For some applications, there may also be a nominal label for benign traffic (e.g., browsing, chat, mail, P2P, streaming, VoIP) \cite{lashkari2017} \cite{bagui2017}.
-
-For data generated by simulating attacks, the labels can simply be based on the schedule of simulated attacks in combination with the IP addresses, ports and protocols involved \cite{Ids2018}.
-
 \section{Conclusion}
-
-The quality of a dataset is generally crucial for the performance of machine learning algorithms, and generating a high quality dataset for intrusion detection systems is not a simple task.
-While the generation of representative benign and malicious data in itself is fundamental, the goal of this paper was to give an overview about the equally important tasks of network simulation, recording and data extraction.
-For each of these tasks, a variety of different approaches has been examined.
-Because each approach has its own strengths and limitations, they have to be evaluated individually for each application, depending on its specific requirements.
-Also, the further development of the Internet and the ongoing introduction of new attack techniques has to be considered, which may affect the suitability of current methods in the future and therefore cause the necessity to alter them or to develop new ones.
+In summary, the performance of neural networks in terms of intrusion detection over different datasets is rather mixed.
+If these machine learning algorithms are tested with the same dataset with which they were trained, the result is promising.
+Using another dataset for testing results in a poor classification performance.
+However, there is lots of potential for future research with neuronal networks in terms of intrusion detection.
+It seems useful to analyse in detail why the mentioned difficulties occur in the classification and which factors are relevant for the neural network.
+An intensive examination of the finding of optimal hyperparameters would also be desirable in the course of future work.
+Furthermore, it would also make sense to simplify the way classification is done.
+Instead, a simple classification like network traffic is “benign” or “not benign” would be conceivable.
+Moreover, it should be checked whether a better performance can be achieved with other machine learning algorithms.
 
 \begin{thebibliography}{00}
-\bibitem{b1} Seifert, C.
-and Reißmann, S.
-and Rieger, S.
-and Pape, C.
-Evaluation von VIRL, GNS3 und Mininet als Virtual Network Testbeds in der Hochschullehre.
-11.
-DFN-Forum Kommunikationstechnologien.
-Bonn: Gesellschaft für Informatik e.V.., pp.
-103-112.
-\bibitem{b2} Network Computing.
-(2020).
-Cisco VIRL: More Than A Certification Study Lab.
-[online] Available at: \url{https://www.networkcomputing.com/networking/cisco-virl-more-certification-study-lab} [Accessed 22 Feb.
-2020].
-\bibitem{b3} Virl.cisco.com. (2020). VIRL Getting Started Tutorial. [online] Available at: \url{http://virl.cisco.com/virl.apis.php} [Accessed 22 Feb. 2020].
-\bibitem{b4} Gns3.com.
-(2020).
-GNS3 | The software that empowers network professionals.
-[online] Available at: \url{https://gns3.com/discussions/how-to-install-gns3-on-centos-7-} [Accessed 22 Feb.
-2020].
-\bibitem{b5} Gns3-server.readthedocs.io.
-(2020).
-Welcome to API documentation!
-— GNS3 2.1.22dev1-67e70c46 documentation.
-[online] Available at: \url{https://gns3-server.readthedocs.io/en/latest/} [Accessed 22 Feb.
-2020].
-\bibitem{b6} Eve-ng.net. (2020). Screenshots. [online] Available at: https://www.eve-ng.net/index.php/screenshots/ [Accessed 22 Feb. 2020].
-\bibitem{b7} Eve-ng.net. (2020). EVE-NG API. [online] Available at: \url{https://www.eve-ng.net/index.php/documentation/howtos/how-to-eve-ng-api/} [Accessed 22 Feb. 2020].
-\bibitem{b8} Bernstein, D. Containers and Cloud: From LXC to Docker to Kubernetes, in IEEE Cloud Computing, vol. 1, no. 3, pp. 81-84, Sept. 2014.
-\bibitem{b9} Sandobalín, J.
-and Insfran, E.
-and Abrahão, S.
-(2020).
-On the Effectiveness of Tools to Support Infrastructure as Code: Model-Driven versus Code-Centric.
-IEEE Access.
-10.1109/ACCESS.2020.2966597.
-\bibitem{b10} Spanaki, P. Cloud Computing: Security Issues and Establishing Virtual Cloud Environment via Vagrant to Secure Cloud Hosts, 2018, pp. 539-553
-\bibitem{b11} Ivanova, D. and Borovska, P. and Zahov, S. Development of PaaS using AWS and Terraform for medical imaging analytics, AIP Conference Proceedings, 2018
-\bibitem{winpcaporg} WinPcap: News, Riverbed Technology (2010), https://www.winpcap.org/misc/copyright.htm [Accessed 26.02.2020]
-\bibitem{tcpdumporg} Van Jacobson, C. Leres, S. McCanne,tcpdump.org (26.02.2020) https://www.tcpdump.org/manpages/tcpdump.1.html [Accessed 27.02.2020]
-\bibitem{goyal2017} P. Goyal, A. Goyal (2017), Comparative study of two most popular packet sniffing tools-Tcpdump and Wireshark
-\bibitem{asrodiapatel2012} P. Asrodia, H. Patel (2012). Analysis  of  Various  Packet  Sniffing  Tools  for Network  Monitoring  and  Analysis, ISSN  No.  (Online)  :  2277-2626
-\bibitem{suribatra2012} S.
-Suri, V.
-Batra, Comparative Study of Network Monitoring Tools, International Journal ofInnovative Technology and Exploring Engineering (IJITEE)ISSN: 2278-3075,Volume-1, Issue-3, August 2012
-\bibitem{firewallcx} http://www.firewall.cx/images/stories/review-capsa-enterprisev7-7.jpg [Accessed 26.02.2020]
-\bibitem{etherapewiki} https://upload.wikimedia.org/wikipedia/commons/a/ae/Etherape-gui.png [Accessed 25.02.2020]
-\bibitem{Nb2015} Moustafa, N.
-(2015).
-The UNSW-NB15 Dataset Description.
-[online] Available at: \url{https://www.unsw.adfa.edu.au/unsw-canberra-cyber/cybersecurity/ADFA-NB15-Datasets/} [Accessed 22 Feb.
-2020]
-\bibitem{Ids2018} Communications Security Establishment (CSE) and Canadian Institute for Cybersecurity (CIC) (2018).
-CSE-CIC-IDS2018 on AWS.
-[online] Available at: \url{https://www.unb.ca/cic/datasets/ids-2018.html} [Accessed 22 Feb.
-2020].
-\bibitem{vasquez2015} Vázquez Iglesias, F. and Zseby, T. (2015). Analysis of network traffic features for anomaly detection. Machine Learning, vol. 101, 2015.
-\bibitem{moustafa2015} Moustafa, N.
-and Slay, J.
-(2015).
-The significant features of the UNSW-NB15 and the KDD99 data sets for Network Intrusion Detection Systems.
-2015 4th International Workshop on Building Analysis Datasets and Gathering Experience Returns for Security (BADGERS), Kyoto, pp.
-25-31, 2015.
-\bibitem{dehghani2010} Dehghani, F.
-and Movahhedinia, N.
-and Khayyambashi, M.
-R.
-and Kianian, S.
-(2010).
-Real-time traffic classification based on statistical and payload content features.
-2010 2nd International Workshop on Intelligent Systems and Applications, Wuhan, pp.
-1-4, 2010.
-\bibitem{xu2015} Xu, M. and Zhu, W. and Xu, J. and Zheng, N. (2015). Towards selecting optimal features for flow statistical based network traffic classification. APNOMS 2015 - pp. 479–482, 08 2015.
-\bibitem{sopuru2019} Sopuru, J.
-and Sari, A.
-and Akkaya, M.
-(2019).
-Modeling a malware detection and categorization system based on seven network flow-based features.
-International Journal of Innovative Technology and Exploring Engineering (IJITEE), vol.
-8(7), pp.
-2982-2989, 05 2019
-\bibitem{celik2015} Celik, Z.
-Berkay and Walls, R.
-J.
-and McDaniel, P.
-and Swami, A.
-(2015).
-Malware traffic detection using tamper resistant features.
-MILCOM 2015 - 2015 IEEE Military Communications Conference, pp.
-330–335, Oct 2015.
-\bibitem{este2009} Este, A.
-and Gringoli, F.
-and Salgarelli, L.
-(2009).
-On the stability of the information carried by traffic flow features at the packet level.
-SIGCOMM Comput.
-Commun.
-Rev., vol.
-39, p.
-13–18, June 2009.
-\bibitem{mantere2013} Mantere, M.
-and Sailio, M.
-and Noponen, S.
-(2013).
-Network traffic features for anomaly detection in specific industrial control system network.
-Future Internet, vol.
-5, pp.
-460-473, 2013
-\bibitem{waizumi2007} Waizumi, Y.
-and Sato, Y.
-and Nemoto, Y.
-(2007).
-A network-based anomaly detection system using multiple network features.
-Proceedings of the Third International Conference on Web Information Systems and Technologies - Internet Technology, pp.
-410-413, 2007.
-\bibitem{cicflowmeter} Canadian Institute for Cybersecurity (CIC) (2016).
-Applications | Research | Canadian Institute for Cybersecurity | UNB.
-[online] Available at: \url{https://www.unb.ca/cic/research/applications.html#CICFlowMeter} [Accessed 24 Feb.
-2020].
-\bibitem{lashkari2017} Lashkari, A.
+\bibitem{max1} F. Ahrens, C. Cat and R. Freingruber, "A Neural-Network Approach for an Intrusion Detection System", FH Technikum, Vienna, 2019.
+\bibitem{max2} I.
+Sharafaldin, A.
 H.
-and Gil, G.
-D.
-and Mamun, M.
-S.
-I.
-and Ghorbani, A.
+Lashkari and A.
 A.
-(2017).
-Characterization of tor traffic using time based features.
-Proceedings of the 3rd International Conference on Information Systems Security and Privacy (ICISSP 2017), pp.
-253-262, 2017.
-\bibitem{wang2018} Wang, W.
-and Sheng, Y.
-and Wang, J.
-and and Zeng, X.
-and Ye, X.
-and Huang, Y.
-and Zhu, M.
-(2018).
-HAST-IDS: Learning hierarchical spatial-temporal features Using deep neural networks to improve intrusion detection.
-IEEE Access, vol.
-6, pp.
-1792-1806, 2018.
-\bibitem{yen2011} Yen, T. (2011). Detecting stealthy malware using behavioral features in network traffic. PhD thesis, Carnegie Mellon University, 2011.
-\bibitem{li2013} Li, W.
-M and Xie, S.
-L.
-and Luo, J.
-and Zhu, X.
-D.
-(2013).
-A detection method for botnet based on behavior features.
-Advanced Information and Computer Technology in Engineering and Manufacturing, Environmental Engineering, vol.
-765 of Advanced Materials Research, pp.
-1512–1517, Trans Tech Publications Ltd, 10 2013.
-\bibitem{abad2003} Abad, C.
-and Taylor, J.
-and Sengul, C.
-and Yurcik, W.
-and Zhou, Y.
-and Rowe, K.
-(2003).
-Log correlation for intrusion detection: a proof of concept.
-19th Annual Computer Security Applications Conference, 2003.
-Proceedings., pp.
-255–264, Dec 2003.
-\bibitem{creech2014} Creech, G.
-and Hu, J.
-(2014).
-A semantic approach to host-based intrusion detection systems using contiguous and discontiguous system call patterns.
-IEEE Transactions on Computers, vol.
-63, no.
-4, pp.
-807-819, April 2014.
-\bibitem{bagui2017} Bagui, S.
-and Fang, X.
-and Kalaimannan, E.
-and Bagui, S.
-C.
-and Sheehan, J.
-(2017).
-Comparison of machine-learning algorithms for classification of vpn network traffic flow using time-related features.
-Journal of Cyber Security Technology, vol.
-1, no.
-2, pp.
-108–126, 2017.
+Ghorbani, "Toward Generating a New Intrusion Detection Dataset and Intrusion Traffic Characterization", in 4th International Conference on Information Systems Security and Privacy (ICISSP), Portugal, 2018.
+\bibitem{max3} D. Meyer, "Performance Assessment \& Tuning - Presentation slides", FH Technikum, Vienna, 2018.
+\bibitem{max4} J.
+Brownlee, "How to Calculate Correlation Between Variables in Python", 3 June 2020.
+[Online].
+Available: https://machinelearningmastery.com/how-to-use-correlation-to-understand-the-relationship-between-variables/.
+[Accessed 27 June 2020].
+\bibitem{max5} S.
+Norena, "Finding Correlation Between Many Variables (Multidimensional Dataset) with Python", 26 April 2018.
+[Online].
+Available: https://medium.com/@sebastiannorena/finding-correlation-between-many-variables-multidimensional-dataset-with-python-5deb3f39ffb3.
+[Accessed 27 June 2020].
 \end{thebibliography}
 
-\vspace{12pt}
-\color{red}
 \end{document}

### 2020-06-28 paper1/paper1.tex
Handover conclusion to Damir
Add infrastructure chapter to paper1

@@
 \section{Introduction}
-@Damir
+@Damir [1p]
+
+\newpage
+test
+\newpage
 
 \section{Infrastructure}
-@Sebastian
+
+This chapter covers the design of the infrastructure to provide a tool for creating generic environments in which specific network scenarios can be simulated to record network packets and extract features with which neural networks inside intrusion prevention systems could be trained and evaluated.
+
+Therefore it is split into four parts.
+The first part designs the fundamental simulation architecture.
+The second part focuses on the creation of the hosts, followed by the configuration of the network in the third part.
+The last part covers the automation of the procedure with the help of provioning tools.
+
+\subsection{Designing the architecture}
+
+A simulation network consists of multiple scenarios based on network attacks and benign traffic.
+In this network generic hosts can be placed and connected to each other or the internet with the usage of hubs, switches and routers.
+This hosts reflect attackers and victims, or usual clients and servers, as well as recorders.
+To ensure recorded features can be labelled, recorders are placed on all positions where another kind of attack or network traffic should be recorded.
+Furthermore they are able to filter out selected traffic to ensure only the important traffic is recorded.
+This means network traffic is recorded within multiple files marked as bad or good which serve as input for the data extractor to extract the features and create the dataset to train and test the neural network.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[scale=0.26]{principle.png}}
+\caption{Design principle}
+\label{principle}
+\end{figure}
+
+Fig.~\ref{principle} shows the design principle.
+The recorders receive the data through network hubs placed between the attackers and victims, or clients and servers and record in pcap files to a central filesystem.
+The scenario names and filter settings are passed to the recorders as parameters.
+The recorded files - named like the associated scenario - are then collected and forwarded to the extractor which knows the names of the malicious scenarios and extracts the selected features to a csv file.
+The extractor is only required once and is not part of the simulation environment.
+
+\subsection{Creating the nodes}
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[scale=0.24]{docker.png}}
+\caption{Node creation}
+\label{node_creation}
+\end{figure}
+
+To offer users a very generic approach Docker containers are used for creating customized network nodes.
+The configuration of the nodes is based on code and can be easily created.
+It is defined within Dockerfiles which can be transformed to an image and uploaded to a registry through Docker client tools as shown in Fig.~\ref{node_creation}.
+When the images are placed on a public Docker registry, they can be pulled and executed from any place.
+This is the default approach in the current infrastructure design.
+
+The lines in a Dockerfile are describing layers of which the image consists.
+The first line is a statement to to specify the base image, for example Alpine Linux or Ubuntu.
+The new image inherites everything from this base layer.
+After that multiple commands can be added to install software, configure the system, copy files to the image or define environment variables.
+The last lines define the entrypoint and/or a command which the container runs when it is started.
+This could be for instance, a program to attack a host, or record the network traffic.
+Docker also supports mounting volumes to the containers when they are started as well as passing arguments to use the same image for similar nodes with different settings.
+More about creating Dockerfiles can be found on the offical documentation of Docker.\footnote{https://docs.docker.com/develop/develop-images/dockerfile\_best-practices}
+
+Once the images are built and uploaded to an accessible registry they can be further used in the simulation environment.
+The target hosts only need the Docker runtime environment installed and access to the image registry.
+In this case the official Docker Hub is used to build and store the images for the scenarios.
+This includes all clients and servers, the recorder and the extractor.
+
+The network recorder is based on a customized Docker image which runs tcpdump with additional arguments like the filename to specify the scenario name and the filter expression to sort out unwanted traffic.
+It records the packets passing the first interface.
+For more information about possible commands and configuration refer to the manpage of tcpdump.\footnote{https://www.tcpdump.org/manpages/tcpdump.1.html}
+
+The extractor more precisely the CI flowmeter is also placed into a Docker container mounting a volume containing all the recorded pcap files.
+It extracts only the features defined by a parameter and saves the csv file with the extracted features at the same place as the pcap files which can be later transferred to another host to evaluate the data.
+
+The runtime configuration and connection of the network nodes as well as the creation of other nodes like switches, hubs and routers is part of the simulation environment which is covered in the next section.
+
+\subsection{Connecting the nodes}
+
+The configuration of the network, more precisely the connection between the network nodes is done within GNS3, an open-source network simulator which can be easily controlled via a native graphical user interface or RESTful API.
+It consists of a server and a client application.
+In this setup the server application is installed on a virtual machine which offers 8-cores, 32 GiB RAM and 300 GiB hard disk space.
+It is possible to install multiple servers but in this case one was sufficient for simulating multiple scenarios including DoS attacks.
+More information about the installation of the server can be found on the official website of the GNS3 project.\footnote{https://docs.gns3.com/1c2Iyiczy6efnv-TS\_4Hc7p11gn03-ytz9ukgwFfckDk/index.html}
+
+Once the server runs a project can be created and the scenarios can be configured via the graphical interface of the GNS3 GUI.
+It is recommended to setup a VPN to the server to ensure the terminals of the network nodes are reachable remotely without configuring network tunnels.
+This means it is necessary to connect to the server via VPN before the GUI can reach the server applicaiton.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[scale=0.5]{network.png}}
+\caption{Network example}
+\label{network}
+\end{figure}
+
+Fig.~\ref{network} shows a simple network created in GNS3.
+The attacker, victim and recorder are created through Docker templates referring the created Docker images before.
+GNS3 supports the configuration of environment variables and volumes for Docker containers.
+All other components including switches, routers and the interface to the internet are basic network components of GNS3 and are already available in the device browser.
+Cisco routers need official IOS images before they can be created.
+
+The project is automatically saved to the filesystem and includes all components which are needed to run the network including configuration, logs and recorded data.
+The simulation can be started through the GUI or the API and produces the pcap files within the project directory.
+Once the simulation is stopped the pcap files can be copied to a temporary directory and send through the extractor container to produce the csv file including the extracted features.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[scale=0.33]{project.png}}
+\caption{Project structure}
+\label{project}
+\end{figure}
+
+The setup of the network is done through the GUI because the API does not support all features to create network nodes.
+The project folder can be checked in into version control and copied to the server when it is needed.
+As shown in Fig.~\ref{project} it consists of the gns3 file which contains all the network configuration in a JSON format.
+The project-files directory contains additional files which are part of the filesystems of network nodes for example the docker containers.
+The text files are the parameters which are passed to the extractor to determine the attack names and the features to extract.
+
+\subsection{Automating the procedure}
+
+The automated setup of the server and run of the simulation is done within Anisble playbooks.
+Therefore multiple roles are created, one for the setup of the GNS3 server, one for running the network simulation and one for extracting the features.
+As shown in Fig.~\ref{ansible} the server role is associated with the setup playbook, while the simulation and extraction roles are used by the run playbook.
+These playbooks are stored in YAMl files and can be executed with the ansible tools.\footnote{https://docs.ansible.com/ansible/latest/user\_guide/playbooks.html\#working-with-playbooks}
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[scale=0.33]{ansible.png}}
+\caption{Ansible roles}
+\label{ansible}
+\end{figure}
+
+The server role set ups the GNS3 server and OpenVPN, copies the device images and projects and configures the Docker runtime environment.
+It downloads the VPN configuration to the client which can be later used to connect to the server.
+Because the test setup restricted the incoming ports, VPN was changed to use TCP instead of UDP to enable the functionality of creating SSH tunnels.
+
+To run the simulation role a VPN connection to the server is required.
+This is accomplished by a simple connection script which opens a SSH tunnel and connects to the server via VPN.
+This is necessary because Ansible needs access to the GNS3 API and the network components.
+When the role is executed it opens the project, stops all network nodes and starts them again.
+The duration can be defined through a parameter.
+When the time is over the nodes are stopped.
+To control GNS3 via the API an additional Ansible collection is used.
+\footnote{https://github.com/davidban77/ansible-collection-gns3}
+
+The extraction role gathers the pcap files, creates a temporary directory and copies all files there.
+Then it determines all attack scenarios and features from a file in the project directory and runs the extraction container with the mounted pcap volume and the passed parameters.
+After that it downloads the csv file with the extracted features.
+If specified in a parameter the pcap files are also downloaded.
+
+The server can be setup by running
+\begin{verbatim}
+ansible-playbook -i hosts setup.yml 
+                 --ask-become-pass
+\end{verbatim}
+
+Once the GNS3 projects are checked in into the Ansible project folder the simulation can be started with
+\begin{verbatim}
+./connect.sh
+ansible-playbook -i hosts run.yml
+                 --ask-become-pass
+                 -e project=<project_name>
+                 -e duration=<minutes>
+                 [-e pcap=true]
+\end{verbatim}
+
+All Ansible commands have to be executed within the gns3 folder and need the sudo password from the server user. The hosts file contains the target hosts configuration.
+
+The next chapter focuses on the implementation and simulation of specific scenarios within a GNS3 project with the help of the introduced infrastructure to produce a new dataset for evaluating neural networks in IPS systems.
 
 \section{Simulation}
-@Bernhard
+@Bernhard [2-3p]
 
 \section{Conclusion}
-@Sebastian
+@Damir [0.5p]
 

### 2020-06-29 paper1/paper1.tex
Correct grammar in the infrastructure chapter of paper1

@@
 
-This chapter covers the design of the infrastructure to provide a tool for creating generic environments in which specific network scenarios can be simulated to record network packets and extract features with which neural networks inside intrusion prevention systems could be trained and evaluated.
+This chapter covers the design of an infrastructure which provides a tool for creating generic environments in which specific network scenarios can be simulated to record network packets and extract features with which neural networks inside intrusion prevention systems could be trained and evaluated.
 
-Therefore it is split into four parts.
+This work is split into four parts.
 The first part designs the fundamental simulation architecture.
-The second part focuses on the creation of the hosts, followed by the configuration of the network in the third part.
+The second focuses on the creation of network nodes, followed by the configuration of the network in the third part.
 The last part covers the automation of the procedure with the help of provioning tools.
@@
 A simulation network consists of multiple scenarios based on network attacks and benign traffic.
-In this network generic hosts can be placed and connected to each other or the internet with the usage of hubs, switches and routers.
-This hosts reflect attackers and victims, or usual clients and servers, as well as recorders.
-To ensure recorded features can be labelled, recorders are placed on all positions where another kind of attack or network traffic should be recorded.
-Furthermore they are able to filter out selected traffic to ensure only the important traffic is recorded.
+In this network, generic hosts can be placed and connected to each other or the internet with the usage of hubs, switches and routers.
+These hosts reflect attackers and victims or usual clients and servers, as well as recorders.
+To ensure recorded features can be labeled, recorders are placed on all positions where another kind of attack or network traffic should be recorded.
+Furthermore, they are able to filter out selected traffic to ensure only the important traffic is recorded.
 This means network traffic is recorded within multiple files marked as bad or good which serve as input for the data extractor to extract the features and create the dataset to train and test the neural network.
@@
 Fig.~\ref{principle} shows the design principle.
-The recorders receive the data through network hubs placed between the attackers and victims, or clients and servers and record in pcap files to a central filesystem.
+The recorders receive the data through network hubs placed between the attackers and victims or clients and servers and record in pcap files to a central filesystem.
 The scenario names and filter settings are passed to the recorders as parameters.
-The recorded files - named like the associated scenario - are then collected and forwarded to the extractor which knows the names of the malicious scenarios and extracts the selected features to a csv file.
+The recorded files - named after the associated scenario - are then collected and forwarded to the extractor which knows the names of the malicious scenarios and extracts the selected features to a csv file.
 The extractor is only required once and is not part of the simulation environment.
+
+\newpage
 
@@
 
-To offer users a very generic approach Docker containers are used for creating customized network nodes.
+Docker containers are used for creating customized network nodes to offer users a very generic approach.
 The configuration of the nodes is based on code and can be easily created.
 It is defined within Dockerfiles which can be transformed to an image and uploaded to a registry through Docker client tools as shown in Fig.~\ref{node_creation}.
-When the images are placed on a public Docker registry, they can be pulled and executed from any place.
+When the images are placed on a public Docker registry, they can be pulled and executed from anywhere.
 This is the default approach in the current infrastructure design.
 
-The lines in a Dockerfile are describing layers of which the image consists.
-The first line is a statement to to specify the base image, for example Alpine Linux or Ubuntu.
+The lines in a Dockerfile describe the layers of which the image consists.
+The first line is a statement to specify the base image, for example Alpine Linux or Ubuntu.
 The new image inherites everything from this base layer.
-After that multiple commands can be added to install software, configure the system, copy files to the image or define environment variables.
+After that, multiple commands can be added to install software, configure the system, copy files to the image or define environment variables.
 The last lines define the entrypoint and/or a command which the container runs when it is started.
-This could be for instance, a program to attack a host, or record the network traffic.
+This could be for instance a program to attack a host or record the network traffic.
 Docker also supports mounting volumes to the containers when they are started as well as passing arguments to use the same image for similar nodes with different settings.
@@
 
-Once the images are built and uploaded to an accessible registry they can be further used in the simulation environment.
+Once the images are built and uploaded to an accessible registry, they can be further used in the simulation environment.
 The target hosts only need the Docker runtime environment installed and access to the image registry.
 In this case the official Docker Hub is used to build and store the images for the scenarios.
-This includes all clients and servers, the recorder and the extractor.
+This includes all clients, servers, the recorder and the extractor.
 
@@
 
-The extractor more precisely the CI flowmeter is also placed into a Docker container mounting a volume containing all the recorded pcap files.
-It extracts only the features defined by a parameter and saves the csv file with the extracted features at the same place as the pcap files which can be later transferred to another host to evaluate the data.
+The extractor, or more precisely the CI flowmeter, is also placed into a Docker container mounting a volume containing all the recorded pcap files.
+It extracts only the features defined by a parameter and saves the csv file with the extracted features at the same place as the pcap files, which can be later transferred to another host to evaluate the data.
 
@@
 
-The configuration of the network, more precisely the connection between the network nodes is done within GNS3, an open-source network simulator which can be easily controlled via a native graphical user interface or RESTful API.
+The configuration of the network, or more precisely the connection between the network nodes, is done within GNS3, an open-source network simulator which can be easily controlled via a native graphical user interface or RESTful API.
 It consists of a server and a client application.
-In this setup the server application is installed on a virtual machine which offers 8-cores, 32 GiB RAM and 300 GiB hard disk space.
-It is possible to install multiple servers but in this case one was sufficient for simulating multiple scenarios including DoS attacks.
+In this setup, the server application is installed on a virtual machine which offers 8-cores, 32 GiB RAM and 300 GiB hard disk space.
+It is possible to install multiple servers, but in this case one was sufficient for simulating multiple scenarios including DoS attacks.
 More information about the installation of the server can be found on the official website of the GNS3 project.\footnote{https://docs.gns3.com/1c2Iyiczy6efnv-TS\_4Hc7p11gn03-ytz9ukgwFfckDk/index.html}
 
-Once the server runs a project can be created and the scenarios can be configured via the graphical interface of the GNS3 GUI.
+Once the server runs, a project can be created and the scenarios can be configured via the graphical interface of the GNS3 GUI.
 It is recommended to setup a VPN to the server to ensure the terminals of the network nodes are reachable remotely without configuring network tunnels.
@@
 Fig.~\ref{network} shows a simple network created in GNS3.
-The attacker, victim and recorder are created through Docker templates referring the created Docker images before.
+The attacker, victim and recorder are created through Docker templates referring to the prevously created Docker images.
 GNS3 supports the configuration of environment variables and volumes for Docker containers.
-All other components including switches, routers and the interface to the internet are basic network components of GNS3 and are already available in the device browser.
+All other components, including switches, routers and the interface to the internet, are basic network components of GNS3 and are already available in the device browser.
 Cisco routers need official IOS images before they can be created.
 
-The project is automatically saved to the filesystem and includes all components which are needed to run the network including configuration, logs and recorded data.
+The project is automatically saved to the filesystem and includes all components which are needed to run the network, including configuration, logs and recorded data.
 The simulation can be started through the GUI or the API and produces the pcap files within the project directory.
-Once the simulation is stopped the pcap files can be copied to a temporary directory and send through the extractor container to produce the csv file including the extracted features.
+Once the simulation is stopped, the pcap files can be copied to a temporary directory and sent through the extractor container to produce the csv file including the extracted features.
 
@@
 The setup of the network is done through the GUI because the API does not support all features to create network nodes.
-The project folder can be checked in into version control and copied to the server when it is needed.
-As shown in Fig.~\ref{project} it consists of the gns3 file which contains all the network configuration in a JSON format.
-The project-files directory contains additional files which are part of the filesystems of network nodes for example the docker containers.
-The text files are the parameters which are passed to the extractor to determine the attack names and the features to extract.
+The project folder can be checked into version control and copied to the server when it is needed.
+As shown in Fig.~\ref{project}, it consists of the gns3 file which contains all the network configuration in a JSON format.
+The project-files directory contains additional files which are part of the filesystems of the network nodes, for example the docker containers.
+The text files are the parameters, which are passed to the extractor to determine the attack names and the features to extract.
 
@@
 The automated setup of the server and run of the simulation is done within Anisble playbooks.
-Therefore multiple roles are created, one for the setup of the GNS3 server, one for running the network simulation and one for extracting the features.
-As shown in Fig.~\ref{ansible} the server role is associated with the setup playbook, while the simulation and extraction roles are used by the run playbook.
+Therefore, multiple roles are created, one for the setup of the GNS3 server, one for running the network simulation and one for extracting the features.
+As shown in Fig.~\ref{ansible}, the server role is associated with the setup playbook, while the simulation and extraction roles are used by the run playbook.
 These playbooks are stored in YAMl files and can be executed with the ansible tools.\footnote{https://docs.ansible.com/ansible/latest/user\_guide/playbooks.html\#working-with-playbooks}
@@
 
-The server role set ups the GNS3 server and OpenVPN, copies the device images and projects and configures the Docker runtime environment.
-It downloads the VPN configuration to the client which can be later used to connect to the server.
+The server role sets up the GNS3 server and OpenVPN, copies the device images and projects and configures the Docker runtime environment.
+It downloads the VPN configuration to the client, which can be later used to connect to the server.
 Because the test setup restricted the incoming ports, VPN was changed to use TCP instead of UDP to enable the functionality of creating SSH tunnels.
 
-To run the simulation role a VPN connection to the server is required.
-This is accomplished by a simple connection script which opens a SSH tunnel and connects to the server via VPN.
+To run the simulation role, a VPN connection to the server is required.
+This is accomplished by a simple connection script which opens an SSH tunnel and connects to the server via VPN.
 This is necessary because Ansible needs access to the GNS3 API and the network components.
-When the role is executed it opens the project, stops all network nodes and starts them again.
+When the role is executed, it opens the project and restarts all network nodes.
 The duration can be defined through a parameter.
-When the time is over the nodes are stopped.
-To control GNS3 via the API an additional Ansible collection is used.
+When the time is over, the nodes are stopped.
+To control GNS3 via the API, an additional Ansible collection is used.
 \footnote{https://github.com/davidban77/ansible-collection-gns3}
@@
 After that it downloads the csv file with the extracted features.
-If specified in a parameter the pcap files are also downloaded.
+If specified in a parameter, the pcap files are also downloaded.
 
-The server can be setup by running
+The server can be setup by running the following command:
 \begin{verbatim}
@@
 
-Once the GNS3 projects are checked in into the Ansible project folder the simulation can be started with
+Once the GNS3 projects are checked into the Ansible project folder, the simulation can be started with the following command:
 \begin{verbatim}
@@
 
-All Ansible commands have to be executed within the gns3 folder and need the sudo password from the server user. The hosts file contains the target hosts configuration.
+All Ansible commands must be executed within the gns3 folder and need the sudo password from the server user. The hosts file contains the target hosts configuration.
 

### 2020-06-30 paper2/paper2.tex
Added introduction and feature extraction

@@
 \section{Introduction}
-@Boris
+The main goal of this paper is the comparison of a new dataset for intrusion prevention systems, which was created using automatically generated network traffic \cite{paper1}, with the CSE-CIC-IDS-2018 dataset \cite{max2}, using both a newly trained neural network and a neural network previously trained using the CSE-CIC-IDS-2018 dataset \cite{max1}.
+Because raw traffic network data cannot be used for training and classification using a neural network, meaningful features have to be extracted first.
+This process is described in Section \ref{extraction}.
+The extracted features can then be used for training and classification using multiple neural networks.
+This task is discussed and its results are shown in Section \ref{nn}.
+Finally, a conclusion based on these results is given in Section \ref{conclusion}.
 
-\section{Feature Extraction}
-@Boris
+\section{Feature Extraction}\label{extraction}
+Before the neural network can be trained and used for classification, features have to be extracted from the raw network traffic data which was recorded in a previous step \cite{paper1}.
+For this task, the open-source Java software CICFlowMeter \cite{cicflowmeter} was chosen.
+The CICFlowMeter software was also used to create the CSE-CIC-IDS-2018 dataset \cite{max2}.
+This was a major reason for using it in this paper, as one of its goals is comparing the newly created dataset to the CSE-CIC-IDS2018 dataset.
+Some adaptations had to be made to the software, however, before it could be used to extract the desired features and also the labels from the recorded network traffic.
+Subsection \ref{extraction_using_cicflowmeter} describes how the CICFlowMeter in general extracts features from recorded network traffic.
+Subsection \ref{adaptation} then outlines the adaptations which had to be made.
 
-\section{Neural Network}
+\subsection{Extraction of flow-based features using CICFlowMeter}\label{extraction_using_cicflowmeter}
+A common approach for extracting features from network traffic is to group packets into flows, and then calculate statistical features based on these flows.
+Because the CICFlowMeter software\cite{cicflowmeter} has been used used to fulfill that task for the creation of various datasets \cite{max2,sopuru,lashkari} and also during the creation of this paper, the following explanation is describing the functionality of that particular software.
+However, there is also a number of applications using various other software, also based on the idea of network-flow generation \cite{wang,moustafa,celik,waizumi}.
+
+For each recorded packet read from a file in the \verb|pcap| format, the \verb|FlowGenerator| first checks if there is already a flow with that source IP address, destination IP address, source port, destination port and protocol (\verb|TCP| or \verb|UDP|).
+When this is the case, the packet information is added to that flow.
+Because CICFlowMeter is actually a bi-flow generator, the IP address and port of the source and destination may also be swapped, in which case the packet is regarded as a backwards packet.
+When no matching flow is found, a new flow is created (the first packet of each flow determines the forward and backward directions).
+A flow ends when a FIN packet is sent (TCP) or a configurable timeout occurs (UDP and TCP, 120 seconds by default), and the state of a flow changes from active to idle after a shorter timeout (default: five seconds).
+A flow is also divided into subflows whenever more than one second elapsed between two packets.
+The \verb|FlowGenerator| aggregates the information of each packet and calculates various statistical features.
+In addition to the features that define a flow (flow ID, IP address, port and protocol), CICFlowMeter supports a wide range of statistical features, including simple packet counts, totals, extrema (max, min), mean and standard deviation (Std).
+Some features, such as the inter-arrival-time (IAT, i.e.
+the time elapsed between two packets), are time-based, others depend on packet size, certain TCP flags and bulk (Blk) data flows.
+As the flows are bidirectional, the features can also distinguish between forward (Fwd) and backward (Bwd) directions.
+The full list of features supported by CICFlowMeter is given in Table \ref{table:cicflowmeter_features}, although the software is open-source and can therefore easily be extended to support currently disregarded features.
+
+\begin{table}[htbp]
+\caption{List of features supported by CICFlowMeter}
+\begin{tabular}{| l l l |}
+\hline
+Flow ID & Src IP & Src Port \\
+Dst IP & Dst Port & Protocol \\
+Timestamp & Flow Duration & Tot Fwd Pkts \\
+Tot Bwd Pkts & TotLen Fwd Pkts & TotLen Bwd Pkts \\
+Fwd Pkt Len Max & Fwd Pkt Len Min & Fwd Pkt Len Mean \\
+Fwd Pkt Len Std & Bwd Pkt Len Max & Bwd Pkt Len Mi \\
+Bwd Pkt Len Mean & Bwd Pkt Len Std & Flow Bytes/s \\
+Flow Pkts/s & Flow IAT Mean & Flow IAT Std \\
+Flow IAT Max & Flow IAT Min & Fwd IAT Tot \\
+Fwd IAT Mean & Fwd IAT Std & Fwd IAT Max \\
+Fwd IAT Min & Bwd IAT Tot & Bwd IAT Mean \\
+Bwd IAT Std & Bwd IAT Max & Bwd IAT Min \\
+Fwd PSH Flags & Bwd PSH Flags & Fwd URG Flags \\
+Bwd URG Flags & Fwd Header Len & Bwd Header Len \\
+Fwd Pkts/s & Bwd Pkts/s & Pkt Len Min \\
+Pkt Len Max & Pkt Len Mean & Pkt Len Std \\
+Pkt Len Var & FIN Flag Cnt & SYN Flag Cnt \\
+RST Flag Cnt & PSH Flag Cnt & ACK Flag Cnt \\
+URG Flag Cnt & CWE Flag Count & ECE Flag Cnt \\
+Down/Up Ratio & Pkt Size Avg & Fwd Seg Size Avg \\
+Bwd Seg Size Avg & Fwd Byts/b Avg & Fwd Pkts/b Avg \\
+Fwd Blk Rate Avg & Bwd Byts/b Avg & Bwd Pkts/b Avg \\
+Bwd Blk Rate Avg & Subflow Fwd Pkts & Subflow Fwd Byts \\
+Subflow Bwd Pkts & Subflow Bwd Byts & Init Fwd Win Byts \\
+Init Bwd Win Byts & Fwd Act Data Pkts & Fwd Seg Size Min \\
+Active Mean & Active Std & Active Max \\
+Active Min & Idle Mean & Idle Std \\
+Idle Max & Idle Min & \\
+\hline
+\end{tabular}
+\label{table:cicflowmeter_features}
+\end{table}
+
+The CICFlowMeter has a command-line interface (which can be accessed by executing the Java main class \verb|cic.cs.unb.ca.ifm.Cmd|) and a graphical user interface (GUI, \verb|cic.cs.unb.ca.ifm.CICFlowMeter|).
+The command-line interface takes two parameters: the name of the input file (in \verb|pcap| format) or directory (in that case, all pcap files in that directory are read) and the name of the output directory.
+For each pcap file which was read, a separate comma-separated values (csv) file containing the extracted features is written to the output directory.
+The GUI allows to adjust additional parameters, such as the flow timeouts (see Figure \ref{fig:cicflowmeter_offline}), and is able to extract features in realtime by directly listening to a network interface (see Figure \ref{fig:cicflowmeter_realtime}).
+
+\begin{figure}[htbp]  
+\centerline{\includegraphics[scale=0.45]{CICFlowMeterOffline.png}}
+\caption{CICFlowMeter GUI (offline mode)}
+\label{fig:cicflowmeter_offline}
+\end{figure}
+
+\begin{figure}[htbp]  
+\centerline{\includegraphics[scale=0.45]{CICFlowMeterRealtime.png}}
+\caption{CICFlowMeter GUI (realtime mode)}
+\label{fig:cicflowmeter_realtime}
+\end{figure}
+
+\subsection{Adaptation of the CICFlowMeter software}\label{adaptation}
+Because the CICFlowMeter is open-source software available at GitHub \cite{cicflowmeter_github} and written in Java, it could easily be extended for the requirements of this paper.
+
+The biggest change was necessary because of the way how the labelling was achieved.
+To ensure both accuracy and flexibility, the network traffic resulting from each host was recorded separately, resulting in multiple files in the pcap format \cite{paper1}.
+The label is defined by the file name: if the file name starts with an attack type label, that label is assigned to it - otherwise, its contents are labelled as benign network traffic.
+For this reason, the command-line interface of the CICFlowMeter was extended.
+It now retrieves a list of attack type labels, separated by a colon (e.g.
+\verb|ssh-bruteforce:bot:dos-goldeneye|) as an additional command-line parameter.
+It then reads the contents of the input directory specified in another command-line parameter and opens every file in it which name ends with \verb|.pcap|.
+So, instead of reading one file at a time as in the original implementation, multiple instances of the class \verb|cic.cs.unb.ca.jnetpcap.PacketReader| are created (one for each pcap file in the input directory), and assigned a label depending on the file name.
+Then, the information of the first valid packet is read from each \verb|PacketReader|.
+In the next step, the timestamps of these packets are compared, and the packet with the lowest timestamp (i.e.
+the earliest) is sent to a single \verb|FlowGenerator| instance.
+The only change made to the \verb|FlowGenerator| implementation is that it now takes the label as a parameter and forwards it to the \verb|BasicFlow| instance.
+The \verb|BasicFlow| keeps track of how many packets per label were found, and then sets the label of the flow accordingly: if all packets of that flow are benign, the label is set to \verb|BENIGN|, otherwise it is set to the attack type label with the highest packet count.
+After processing the packet data, the next packet from the same \verb|PacketReader| is read.
+Afterwards, the determination of the packet with the lowest timestamp is repeated, and its contents are also sent to the \verb|FlowGenerator|.
+When the end of file (EOF) has been reached for any pcap file, the corresponding \verb|PacketReader| is no longer considered, and the process continues with the remaining \verb|PacketReader| instances.
+This is repeated until all packets from all opened pcap files have been read.
+
+Another change was made in order to efficiently use the newly created dataset with the neural network from \cite{max1}.
+For that reason, and to achieve as much flexibility as possible, another command-line parameter was added to the CICFlowMeter.
+This parameter lists all features which should be extracted (as with the attack labels, separated by a colon).
+The function which writes the flow features to the resulting \verb|csv| file was extended and now retrieves the list of flow features, which is passed from the newly added command-line parameter.
+Instead of writing all features supported by the CICFlowMeter, it now writes only the desired features, in the exact order they were specified.
+Of course, this also has to be considered when writing the headers to the first line of the \verb|csv| file (these list the names of the features).
+The new parameter allows the data to be extracted in a way that does not unnecessarily complicate the preprocessing of the neural network.
+If the parameter is ommitted, all features are written to the \verb|csv| file, exactly as in the original version of the CICFlowMeter.
+
+For optimal integration into the infrastructure created in \cite{paper1}, a Docker image for feature extraction was created.
+For reasons of efficiency, the \verb|Dockerfile| used to create that image was written so that it employs a multi-stage build \cite{docker-multistage}.
+The first stage uses the image \verb|gradle:6.3.0-jdk11| (the Gradle and Java versions are set explicitly to ensure compatibility with the CICFlowMeter source code) from Docker Hub \cite{dockerhub}, copies the modified CICFlowMeter source code from the build context (in this case this is the directory that the \verb|Dockerfile| is contained in) into a temporary image and runs \verb|gradle| to build the CICFlowMeter distribution.
+The second stage is based on the image \verb|azul/zulu-openjdk:11| (again, the Java version is set explicitly).
+It runs \verb|jlink| \cite{jlink} to create a minimal Java 11 Runtime Environment containing only the modules required by the CICFlowMeter, which is written to another temporary image.
+Finally, the last stage builds the actual feature extraction Docker image.
+It is based on \verb|ubuntu:latest|, installs the \verb|libpcap| library required by the CICFlowMeter and copies the Java Runtime Environment and the CICFlowMeter distribution from the temporary images created in the previous stages.
+The \verb|ENTRYPOINT| of the docker image is set to the Java Virtual Machine binary, and all command-line parameters which are required to run the command-line interface of the CICFlowMeter are set as well.
+The data directory is set to \verb|/mnt/packet-data|, which means that the external directory containing all the pcap files has to be mounted to that directory of the Docker container.
+The \verb|csv| file with the extracted features is also written to that directory.
+Another dependency required by the Docker container is the environment variable \verb|ATTACKS|, which has to contain the attack labels separated by a colon.
+The list of features is read from the environment variable \verb|FEATURES|.
+When this environment variable is missing, all supported features are extracted.
+
+\section{Neural Network}\label{nn}
 The created dataset will be used in the further course of this paper to build a neural network.
@@
 
-\section{Conclusion}
+\section{Conclusion}\label{conclusion}
 In summary, the performance of neural networks in terms of intrusion detection over different datasets is rather mixed.
@@
 \begin{thebibliography}{00}
-\bibitem{max1} F. Ahrens, C. Cat and R. Freingruber, "A Neural-Network Approach for an Intrusion Detection System", FH Technikum, Vienna, 2019.
+\bibitem{paper1} B. Gally, S. Lipp and D. Marijanovic, "Automated Network Packet Generation for Evaluation of Neural Networks in Intrusion Prevention Systems", FH Technikum, Vienna, 30 June 2020.
 \bibitem{max2} I.
@@
 Ghorbani, "Toward Generating a New Intrusion Detection Dataset and Intrusion Traffic Characterization", in 4th International Conference on Information Systems Security and Privacy (ICISSP), Portugal, 2018.
+\bibitem{max1} F. Ahrens, C. Cat and R. Freingruber, "A Neural-Network Approach for an Intrusion Detection System", FH Technikum, Vienna, 2019.
+\bibitem{cicflowmeter} Canadian Institute for Cybersecurity (CIC), "Applications | Research | Canadian Institute for Cybersecurity | UNB", 2016.
+[Online].
+Available: https://www.unb.ca/cic/research/applications.html\#CICFlowMeter.
+[Accessed 29 June 2020].
+\bibitem{sopuru} J.
+Sopuru, A.
+Sari and M.
+Akkaya, "Modeling a malware detection and categorization system based on seven network flow-based features", in International Journal of Innovative Technology and Exploring Engineering (IJITEE), vol.
+8(7), pp.
+2982-2989, 05 2019
+\bibitem{lashkari} A.
+H.
+Lashkari, G.
+D.
+Gil, M.
+S.
+I.
+Mamun and A.
+A.
+Ghorbani, "Characterization of tor traffic using time based features", in Proceedings of the 3rd International Conference on Information Systems Security and Privacy (ICISSP 2017), pp.
+253-262, 2017.
+\bibitem{wang} W.
+Wang, Y.
+Sheng, J.
+Wang, X.
+Zeng, X.
+Ye, Y.
+Huang and M.
+Zhu, HAST-IDS: Learning hierarchical spatial-temporal features Using deep neural networks to improve intrusion detection.
+IEEE Access, vol.
+6, pp.
+1792-1806, 2018.
+\bibitem{moustafa} N.
+Moustafa and J.
+Slay, "UNSW-NB15: a comprehensive data set for network intrusion detection systems (UNSW-NB15 network data set)", in 2015 Military Communications and Information Systems Conference (MilCIS), Canberra, ACT, 2015, pp.
+1-6.
+\bibitem{celik} Z.
+Celik, Berkay, R.
+J.
+Walls, P.
+McDaniel and A.
+Swami, "Malware traffic detection using tamper resistant features", in MILCOM 2015 - 2015 IEEE Military Communications Conference, pp.
+330–335, Oct 2015.
+\bibitem{waizumi} Y.
+Waizumi, Y.
+Sato and Y.
+Nemoto, "A network-based anomaly detection system using multiple network features", in Proceedings of the Third International Conference on Web Information Systems and Technologies - Internet Technology, pp.
+410-413, 2007.
+\bibitem{cicflowmeter_github} GitHub - ahlashkari/CICFlowMeter: CICFlowmeter-V4.0 (formerly known as ISCXFlowMeter) is an Ethernet traffic Bi-flow generator and analyzer for anomaly detection that has been used in many Cybersecurity datsets such as Android Adware-General Malware dataset (CICAAGM2017), IPS/IDS dataset (CICIDS2017), Android Malware dataset (CICAndMal2017) and Distributed Denial of Service (CICDDoS2019).
+[Online].
+Available: https://github.com/ahlashkari/CICFlowMeter.
+[Accessed 29 June 2020].
+\bibitem{docker-multistage} Use multi-stage builds | Docker Documentation. [Online]. Available: https://docs.docker.com/develop/develop-images/multistage-build/. [Accessed 29 June 2020].
+\bibitem{dockerhub} Docker Hub. [Online]. Available: https://hub.docker.com/. [Accessed 29 June 2020].
+\bibitem{jlink} The Incredible Shrinking Java Platform - Azul Systems, Inc.. [Online]. Available: https://www.azul.com/blog/the-incredible-shrinking-java-platform/. [Accessed 29 June 2020].
 \bibitem{max3} D. Meyer, "Performance Assessment \& Tuning - Presentation slides", FH Technikum, Vienna, 2018.

### 2020-06-30 gns3/docker/extractor/CICFlowMeter/src/main/java/cic/cs/unb/ca/ifm/Cmd.java
Revert "Added IGNORE environment variable to ignore certain ethernet types from attacks"
This reverts commit 59b723ee07466cc0befad493b63635e8e2529aea.

@@ -68,33 +68,12 @@ public class Cmd {
             return;
         }
         Label attackTypes[] = Label.getAttackTypeLabels(args[2]);
-        if (args.length < 4) {
-            logger.info("Please supply the ignored ethernet types (separated by ':')!");
-            return;
-        }
-        Set<Integer> ignoredTypes = new HashSet<>();
-        for(String typeString : args[3].split(":")) {
-            if(!typeString.trim().isEmpty()) {
-                Integer type;
-                try {
-                    if (typeString.startsWith("0x")) {
-                        type = Integer.valueOf(typeString.substring(2), 16);
-                    } else {
-                        type = Integer.valueOf(typeString);
-                    }
-                } catch (NumberFormatException e) {
-                    logger.info(typeString + " is not a valid integer!");
-                    return;
-                }
-                ignoredTypes.add(type);
-            }
-        }
         List<FlowFeature> features;
-        if (args.length < 5 || args[4].isEmpty()) {
+        if (args.length < 4 || args[3].isEmpty()) {
             logger.info("Extracting all features, because the feature list has not been specified via command-line");
             features = Arrays.asList(FlowFeature.values());
         } else {
-            String[] featureNames = args[4].split(",");
+            String[] featureNames = args[3].split(",");
             features = new ArrayList<>(featureNames.length);
             for(String featureName : featureNames) {
                 FlowFeature feature = FlowFeature.getByName(featureName);
@@ -109,13 +88,12 @@ public class Cmd {
         logger.info("You select: {}",pcapPath);
         logger.info("Out folder: {}",outPath);
         logger.info("Attack Types: {}", Arrays.toString(attackTypes));
-        logger.info("Ignored Ethernet Types: {}", ignoredTypes);
         logger.info("Features: {}",features);
 
-        readPcapFiles(in, attackTypes, ignoredTypes, outPath, features, flowTimeout, activityTimeout);
+        readPcapFiles(in, attackTypes, outPath,features,flowTimeout,activityTimeout);
 }
 
-    private static void readPcapFiles(File inDir, Label[] attackTypes, Set<Integer> ignoredTypes, String outPath, List<FlowFeature> features, long flowTimeout, long activityTimeout) {
+    private static void readPcapFiles(File inDir, Label[] attackTypes, String outPath, List<FlowFeature> features, long flowTimeout, long activityTimeout) {
         if(inDir==null || attackTypes==null || outPath==null ) {
             return;
         }
[…]

### 2020-06-30 gns3/docker/extractor/CICFlowMeter/src/main/java/cic/cs/unb/ca/jnetpcap/PacketReader.java

@@ -1,6 +1,5 @@
 package cic.cs.unb.ca.jnetpcap;
 
-import com.google.common.collect.ImmutableSet;
 import org.jnetpcap.Pcap;
 import org.jnetpcap.PcapClosedException;
 import org.jnetpcap.PcapHeader;
@@ -17,9 +16,6 @@ import org.jnetpcap.protocol.tcpip.Udp;
 import org.jnetpcap.protocol.vpn.L2TP;
 import org.slf4j.Logger;
 import org.slf4j.LoggerFactory;
-
-import java.util.Collections;
-import java.util.Set;
 
 public class PacketReader {
 
@@ -40,14 +36,12 @@ public class PacketReader {
 	
 	private boolean readIP6;
 	private boolean readIP4;
-	private Set<Integer> ignoredTypes;
 	private String file;
 	
 	public PacketReader(String filename) {
 		super();	
 		this.readIP4 = true;
-		this.readIP6 = false;
-		ignoredTypes = Collections.emptySet();
+		this.readIP6 = false;		
 		this.config(filename);
 	}
 	
@@ -55,18 +49,9 @@ public class PacketReader {
 		super();	
 		this.readIP4 = readip4;
 		this.readIP6 = readip6;
-		ignoredTypes = Collections.emptySet();
 		this.config(filename);
-	}
-
-	public PacketReader(String filename, boolean readip4, boolean readip6, Set<Integer> ignoredTypes) {
-		super();
-		this.readIP4 = readip4;
-		this.readIP6 = readip6;
-		this.ignoredTypes = ImmutableSet.copyOf(ignoredTypes);
-		this.config(filename);
-	}
-
+	}	
+	
 	private void config(String filename){
         file = filename;
 		StringBuilder errbuf = new StringBuilder(); // For any error msgs
@@ -95,29 +80,25 @@ public class PacketReader {
 		 try{
 			 if(pcapReader.nextEx(hdr,buf) == Pcap.NEXT_EX_OK){
 				 packet = new PcapPacket(hdr, buf);
-				 packet.scan(Ethernet.ID);
-				 Ethernet ethernet = new Ethernet();
-				 packet.getHeader(ethernet);
-				 if(ethernet == null || ignoredTypes.contains(ethernet.type())) {
-					 packetInfo = null;
-				 } else {
-					 if(this.readIP4){
-						 packetInfo = getIpv4Info(packet);
-						 if (packetInfo == null && this.readIP6){
[…]

### 2020-06-30 paper2/paper2.tex
Added abstract and implemented feedback from Boris

@@
 
-\title{Title}
+\title{Performance Evaluation of Neural Networks for Intrusion Prevention Systems}
 
@@
 \begin{abstract}
-@TBA
+The detection of malicious behavior in network traffic through the use of machine learning algorithms in intrusion prevention systems is still a current topic of research.
+This paper focuses on neural networks and examines their performance across different datasets.
+Basically, the quality of the data with which machine learning algorithms are trained is a decisive factor for the classification performance.
+In addition to the publicly available CSE-CIC-IDS2018 dataset, this work also uses a dataset that was created with the help of feature extraction from automatically generated network traffic.
+Different scenarios are tested, but it turns out that the classification performance of neural networks in this context is only satisfactory if they are tested with data from the dataset with which they were trained.
+If another dataset is used for testing, the result of the classification is poor.
+This indicates that neural networks react sensitively to differences in the datasets and that the classification performance is influenced by this.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-artificial intelligence, neural networks, computer networks, network security
+artificial intelligence, neural networks, computer networks, network security, feature extraction
 \end{IEEEkeywords}
@@
 
-It is suspected that there are differences in the intrusion detection datasets because, for example, cyber attacks are not always executed or recorded in the same way or the network infrastructure is structured differently.
+It is expectable that there are differences in the intrusion detection datasets because, for example, cyber attacks are not always executed or recorded in the same way or the network infrastructure is structured differently.
 The aim is to determine how sensitive neural networks react to this and are thus dependent on the data with which they have been trained.
@@
 
-\newpage
-
 While the CSE-CIC-IDS2018 dataset \cite{max2} has a total of 10,182,638 data points (6.41GB distributed over multiple CSV files), the dataset created in this work consists of 7,654,545 data points (4.1GB, only one CSV file).
@@
 0001, Optimizer: adam, Weight initialization: normal, Activation function: relu, Number of neurons per layer: 69, Epochs: 850, Batch size: 256) are reused.
-In addition, "MinMaxScale" was used to normalize the data and a weight value of 500 was set for benign" traffic.
+In addition, "MinMaxScale" was used to normalize the data and a weight value of 500 was set for benign traffic.
 At this point it is worth mentioning that attempts have been made to adjust individual parameters such as the learn rate or the weight value of benign traffic in order to achieve a better result.
@@
 \centerline{\includegraphics[scale=0.65]{NeuesModellNeueDaten.png}}
-\caption{Confusion Matrix for Szenario 1}
+\caption{Confusion Matrix for Scenario 1}
 \label{fig:cm1}
@@
 \centerline{\includegraphics[scale=0.65]{NeuesModellAlteDaten.png}}
-\caption{Confusion Matrix for Szenario 2}
+\caption{Confusion Matrix for Scenario 2}
 \label{fig:cm2}
@@
 This continues the low classification success (see Figure \ref{fig:cm3}).
-In the class "benign", 99.8\% of all data are correctly predicted (recall), but the precision is noticeable low (0.
+In the class "benign", 99.8\% of all data are correctly predicted (recall), but the precision is noticeably low (0.
 008\%).
@@
 \centerline{\includegraphics[scale=0.65]{AltesModellNeueDaten.png}}
-\caption{Confusion Matrix for Szenario 3}
+\caption{Confusion Matrix for Scenario 3}
 \label{fig:cm3}
@@
 
+\newpage
+
 \section{Conclusion}\label{conclusion}
@@
 Using another dataset for testing results in a poor classification performance.
-However, there is lots of potential for future research with neuronal networks in terms of intrusion detection.
+However, there is lots of potential for future research with neural networks in terms of intrusion detection.
 It seems useful to analyse in detail why the mentioned difficulties occur in the classification and which factors are relevant for the neural network.

### 2020-06-30 paper1/paper1.tex
added simulation part

@@
 \usepackage{xcolor}
+\usepackage{listings}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 \section{Simulation}
-@Bernhard [2-3p]
+Benign and malicious traffic have to be generated for multiple scenarios in order to provide a diverse dataset that can be used to train the neural network.
+As previously mentioned these scenarios are divided into separate groups, each with their own recorder so that benign and malicious traffic do not mix.
+Each attack and its benign counterpart were implemented using Docker containers which can be found on Dockerhub.
+
+\subsection{Secure Shell (SSH)}
+SSH is one of the most common protocols used today.
+By utilizing it secure connections over the internet can be established to remote servers in order to execute commands or copy files from one machine to another.
+Simulating such a configuration therefore requires a server, a benign client and a malicious client
+
+\subsubsection{SSH server}
+To allow a client to connect to the SSH Server, the client has to exist on the remote host.
+We therefore added a new user with a known password to the server.
+Because the password can only be set in an interactive way via useradd, we had to use the chpasswd utility.
+The configuration of the server can be seen below:
+
+\begin{lstlisting}[basicstyle=\footnotesize]
+FROM ubuntu:latest
+
+RUN apt-get update && \ 
+    apt-get install -y openssh-server && \
+    mkdir -p /var/run/sshd && \
+    useradd technikum -m -d /home/technikum && \
+    echo technikum:technikum | chpasswd
+
+EXPOSE 22
+CMD ["/usr/sbin/sshd", "-D"]
+\end{lstlisting}
+
+\subsubsection{Benign client}
+The idea behind our implementation of benign SSH traffic is to first connect to an SSH server.
+Secondly, send commands that are executed on executed on the remote server which return a response message that generates traffic over the network.
+Because each command will only produce a minute amount of traffic multiple clients are needed to increase the data volume.
+Furthermore, since it is not possible to enter a password via the default ssh tool in a non-interactive way sshpass had to be installed.
+This utility allows us to provide the password used for the SSH login in a non-interactively.
+The exact implementation of the benign traffic container that was used in our simulation can be seen below:
+
+\begin{lstlisting}[basicstyle=\footnotesize]
+FROM ubuntu:latest
+
+ENV targetip=192.168.1.100
+ENV command="echo test; mkdir test; rmdir test;"
+
+RUN apt-get update && \ 
+    apt-get install -y sshpass
+
+ENTRYPOINT while true; do sshpass -p "technikum" \ 
+			ssh -o StrictHostKeyChecking=no \
+			"technikum@${targetip}" ${command}; \
+			sleep 5; done
+\end{lstlisting}
+
+\subsubsection{Malicious client}
+By making use of a password list an SSH brute-force attack can be executed quite easily on an SSH server.
+An attacker uses the list in combination with a given username to attempt a successful login on the victim machine.
+Luckily, some Docker images created for pen-testing come equipped with such password lists already.
+The utility that was used to attack the SSH server is called hydra, a parallelized password cracking tool which supports multiple protocols and the use of password lists.
+To further increase the generated data multiple containers were used.
+The configuration of the malicious ssh containers can be seen below:
+
+\begin{lstlisting}[basicstyle=\footnotesize]
+FROM c4nc/kalibase:latest
+
+RUN gunzip /usr/share/wordlists/rockyou.txt.gz
+
+ENTRYPOINT ["hydra", "-l", "root", \
+            "-P", "/usr/share/wordlists/rockyou.txt", \
+            "192.168.0.10", "-t", "4", "ssh"]
+\end{lstlisting}
+
+\subsection{Botnet}
+The second scenario that was used is a botnet.
+In a botnet a master or server controls multiple agents to perform some action.
+Screenshots, file downloads, uploads, installations and much more can be done once the server has control over the infected pc.
+
+\subsubsection{Server}
+The server controls the agents with a list of commands.
+The server send these commands out to the agents which then perform the requested operations.
+In our simulation we let all clients do the following operations repeatedly:
+
+\begin{lstlisting}[basicstyle=\footnotesize]
+apt -y install tor
+download http://ipv4.download.thinkbroadband.com/5MB.zip
+screenshot
+mkdir -p /tmp/zzz
+cp /etc/passwd /tmp/zzz/
+cp /etc/shadow /tmp/zzz/
+zip zzz.zip /tmp/zzz/
+upload zzz.zip
+rm -rf /tmp/zzz zzz.zip 5MB.zip
+apt -y remove tor
+download http://ipv4.download.thinkbroadband.com/100MB.zip
+apt -y install curl wget netcat
+
+\end{lstlisting}
+
+\subsubsection{Agent}
+Similarly to the benign traffic, a single host would not generate a noticeable amount of network traffic.
+Multiple agents were added to the scenario to increase the generated data.
+By using Ares the required configuration for the agents is the port and IP address of the control server:
+
+\begin{lstlisting}[basicstyle=\footnotesize]
+FROM ubuntu:18.04
+USER root
+WORKDIR /opt
+
+RUN apt-get update && \
+    apt-get install -y \
+      git \
+      python-pip && \
+    apt-get clean all
+
+RUN git clone https://github.com/sweetsoftware/Ares.git && \
+    cd Ares && \
+    pip install -r requirements.txt
+
+ARG SERVER=192.168.0.200
+ARG PORT=8080
+WORKDIR /opt/Ares/agent
+RUN sed -i "s/localhost/$SERVER/g" ./config.py
+RUN sed -i "s/8080/$PORT/g" ./config.py
+ENTRYPOINT ["./agent.py"]
+\end{lstlisting}
+
+\subsection{Goldeneye}
+Goldeneye is a DoS attack that can take down webservers.
+Similar to other DoS tools like Slowloris it occupies all available sessions of the webserver and therefore locks out anyone else trying to connect to it.
+Unlike Slowloris it uses Keep-Alive to keep its connected sessions alive.
+In order to be able to use this attack a webserver must be provided. Because of performance issues the attack could not be ran continuously and had to be timed out every few minutes.
+
+\subsection{Benign traffic}
+Network traffic consists of many different network protocols and connections.
+Furthermore, attacks generally speaking are only a small portion of the network traffic.
+Most of the traffic is benign traffic that also has to be simulated in some way in our environment.
+To overcome this problem multiple benign containers were started that generated harmless traffic.
+Downloading files and browsing the internet are the most common type of traffic in normal network.
+Although it is worth mentioning that the definition of normal traffic is highly dependant on what kind of servers are connected to the network.
+A server farm for example has a different definition of normal traffic than a home network regarding amount and protocols used.
+For our environment we considered file downloads and web browsing as normal traffic.
+Hence, Docker containers that downloaded 10MB, 100MB files were added to the network.
+A third container was added to browse the web.
+The containers do not store the downloaded files anywhere.
+They are redirected to /dev/null meaning the containers do not need any significant amount of storage space but still download every byte of the requested file before restarting.
+After the downloads are finished the container waits 5 seconds before it retries the same operation.
+The configurations of these containers can be seen below:
+
+\begin{lstlisting}[basicstyle=\footnotesize]
+FROM ubuntu:latest
+
+ENV website="google.at"
+
+RUN apt-get update && \
+    apt-get install -y wget
+
+ENTRYPOINT  while true; \
+            do wget -O /dev/null ${website}; \ 
+            sleep 5; done
+\end{lstlisting}
 

### 2020-07-12 paper2/paper2.tex

@@
 
-It is expectable that there are differences in the intrusion detection datasets because, for example, cyber attacks are not always executed or recorded in the same way or the network infrastructure is structured differently.
+It is expectable that there are differences in intrusion detection or -prevention datasets because, for example, cyber attacks are not always executed or recorded in the same way or the network infrastructure is structured differently.
 The aim is to determine how sensitive neural networks react to this and are thus dependent on the data with which they have been trained.
@@
 \section{Conclusion}\label{conclusion}
-In summary, the performance of neural networks in terms of intrusion detection over different datasets is rather mixed.
+In summary, the performance of neural networks in terms of intrusion prevention over different datasets is rather mixed.
 If these machine learning algorithms are tested with the same dataset with which they were trained, the result is promising.
 Using another dataset for testing results in a poor classification performance.
-However, there is lots of potential for future research with neural networks in terms of intrusion detection.
+However, there is lots of potential for future research with neural networks in this context.
 It seems useful to analyse in detail why the mentioned difficulties occur in the classification and which factors are relevant for the neural network.

### 2020-07-12 paper1/paper1.tex

@@
 
-ENTRYPOINT while true; do sshpass -p "technikum" \ 
+ENTRYPOINT  while true; \
+            do sshpass -p "technikum" \
             ssh -o StrictHostKeyChecking=no \ 
@@
 ENTRYPOINT ["hydra", "-l", "root", \
-            "-P", "/usr/share/wordlists/rockyou.txt", \
+            "-P", \
+            "/usr/share/wordlists/rockyou.txt", \
             "192.168.0.10", "-t", "4", "ssh"]
@@
 apt -y install tor
-download http://ipv4.download.thinkbroadband.com/5MB.zip
+download \
+  http://ipv4.download.thinkbroadband.com/5MB.zip
 screenshot
@@
 apt -y remove tor
-download http://ipv4.download.thinkbroadband.com/100MB.zip
+download \
+  http://ipv4.download.thinkbroadband.com/100MB.zip
 apt -y install curl wget netcat
@@
 
-RUN git clone https://github.com/sweetsoftware/Ares.git && \
+RUN git clone \ 
+    https://github.com/sweetsoftware/Ares.git && \
     cd Ares && \

### 2020-07-12 paper1/paper1.tex
Abstract, Intorduction and Conclusion added
Abstract, Intorduction and Conclusion added

--- 2020-06-29 paper1/paper1.tex
@@
 \begin{abstract}
-@Damir
+This Paper is a part of a project.
+It describes the part of automated network packet generation.
+The generated packets are used to evaluate neural networks in intrusion prevention systems.
+The challenges in this project include building the infrastructure and connecting the individual nodes in the network.
 \end{abstract}
@@
 \section{Introduction}
-@Damir [1p]
-
-\newpage
-test
-\newpage
+For a good, respectively successful detection of malicious network traffic using neural networks, the neural network has to be trained well.
+Therefore a large set of labelled network traffic is necessary.
+The more attack scenarios there were, the better the differentiation between attacks and normal traffic.
+Of course, not all possible scenarios can be dealt with here, which is why you have to concentrate on the attacks that occur most often in the real world.
+This subproject relates to the generation, labelling, distribution and recording of network data and packets.
+The network was deliberately kept simple and the different types of attacks were treated separately.
+The infrastructure was considered from scratch and implemented into a network simulation tool called \textit{GNS3}.
+To make the individual parts dynamic, the decision was made to use docker containers.
+This made it possible to evaluate the individual scenarios separately.
+The traffic was recorded using the \textit{tcpdump}, and the entire filtered traffic was saved in so-called pcap files.
+As far as the hardware is concerned, sufficient computing power was required, as well as sufficient storage space to be able to record the traffic.
 
@@
 \section{Conclusion}
-@Damir [0.5p]
+The result of this project can be used in the knowledge as well as in the structure.
+Due to the dynamic structure of the infrastructure, extensions can be added with little effort.
+During the implementation of the infrastructure, minor difficulties came up again and again, but these could usually be remedied with a relatively small change.
+Among other things, difficulties arose in the transfer of environment variables in the individual Docker containers.
+The other part of this project used this project as the basis, in which the recorded information and packages were evaluated.
+The successful evaluation showed that both the infrastructure and the individual nodes worked well.
 

### 2020-07-12 paper1/paper1.tex
Added Abstract, Intorduction and Conclusion

--- 2020-07-12 paper1/paper1.tex
@@
 \begin{abstract}
-@Damir
+This paper is a part of a project.
+It describes the part of automated network packet generation.
+The generated packets are used to evaluate neural networks in intrusion prevention systems.
+The challenges in this project include
+building the infrastructure and connecting individual nodes in the network.
 \end{abstract}
@@
 \section{Introduction}
-@Damir [1p]
+For a good, respectively successful detection of malicious network traffic using neural networks, the neural network has to be trained well.
+Therefore a large set of labelled network traffic is necessary.
+The more attack scenarios there are, the better the differentiation between attacks and normal traffic.
+Of course, not all possible scenarios can be dealt with here, which is why we have to concentrate on the attacks that occur most often in the real world.
+This subproject relates to the generation, labelling, distribution and recording of network data and packets.
 
-\newpage
-test
-\newpage
+The network has to be deliberately kept simple and the different types of attacks have to be treated separately.
+The infrastructure is considered from scratch and implemented into a network simulation tool called \textit{GNS3}.
+To make the individual parts dynamic, docker containers are used.
+This makes it possible to evaluate the individual scenarios separately.
+The traffic is recorded using the \textit{tcpdump}, and the entire filtered traffic is saved in so-called pcap files.
+As far as the hardware is concerned, sufficient computing power is required, as well as sufficient storage space to be able to record the traffic.
+
+The design of these parts are covered in the next two sections.
+The first part describes the infrastructure to create simulation environments.
+The second part focuses on the implementation of specific scenarios.
 
 \section{Infrastructure}
-
 This chapter covers the design of an infrastructure which provides a tool for creating generic environments in which specific network scenarios can be simulated to record network packets and extract features with which neural networks inside intrusion prevention systems could be trained and evaluated.
@@
 \subsection{Designing the architecture}
-
 A simulation network consists of multiple scenarios based on network attacks and benign traffic.
@@
 The extractor is only required once and is not part of the simulation environment.
-
-\newpage
 
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[scale=0.5]{network.png}}
+\centerline{\includegraphics[scale=0.4]{network.png}}
 \caption{Network example}
@@
 \subsection{Automating the procedure}
-
 The automated setup of the server and run of the simulation is done within Anisble playbooks.
@@
 \end{verbatim}
+
+\newpage
 
@@
 \section{Conclusion}
-@Damir [0.5p]
-
-\begin{thebibliography}{00}
-\bibitem{b1} zzz
-\end{thebibliography}
+The result of this project can be used in the knowledge as well as in the structure.
+Due to the dynamic structure of the infrastructure, extensions can be added with little effort.
+During the implementation of the infrastructure, minor difficulties came up again and again, but these could usually be remedied with a relatively small change.
+Among other things, difficulties arose in the transfer of environment variables in the individual Docker containers.
+The other part of this project used this project as the basis, in which the recorded information and packages were evaluated.
+The successful evaluation showed that both the infrastructure and the individual nodes worked well.
 
